---
title: "Tidying data"
author: "Rich Cottrell"
date: "15/02/2022"
output: html_document
---

```{r}
library(tidyverse)
library(parallel)
library(raster)
library(janitor)
library(here)
library(countrycode)
library(rredlist)
library(raster)

source(here("src/directories.R"))
source(here("src/spatial.R"))

```


This is the code to unzip and adjust the extents of the mapspam rasters (-180, 180, -90, 90) - DO NOT RUN IN PROJECT - here as a backup to scripts in the raw data files

```{r}


mapspam_zip_files <- list.files(path = mapspam_dir, pattern = "\\.zip$")


lapply(X=mapspam_zip_files, FUN = unzip)


#Adjust the extent of rasters to full lon lat coveragge

#test
banana <- raster(file.path(mapspam_dir, "spam2010V2r0_global_H_BANA_H.tif"))
banana

#this should be the new extent for all
new_extent <- c(-180, 180, -90, 90)

extent(banana) <- new_extent

#apply new extents to all files

tifs <- list.files(dir, pattern = "\\.tif")

adjust_extent <- function(this_file){
  message(paste("processing", which(tifs==this_file), "of", length(tifs)))
  this_raster <- raster(this_file)
  extent(this_raster) <- new_extent
  writeRaster(x=this_raster, filename = file.path("/mnt/rdsi/raw_data/MAPSPAM/new-extents", this_file), overwrite = TRUE)
}

#run over mutliple cores to speed things up
parallel::mclapply(X = tifs, FUN = adjust_extent, mc.cores = 6)


```



This is the code to unzip the gridded livestock of the world rasters - DO NOT RUN IN PROJECT - here as a backup to scripts in the raw data files

```{r}

dir <- "/mnt/rdsi/raw_data/gridded-livestock-of-the-world"

setwd(dir)

library(tidyverse)
library(parallel)
library(raster)

glw_zip_files <- list.files(dir, pattern = "\\.zip$")

unzip_raw_data <- function(this_file){
  this_file_name <- file.path(dir, this_file)
  unzip(this_file_name)
  
}


lapply(X = glw_zip_files, FUN = unzip_raw_data)



```


Tidy global aquaculture production data

```{r}

(aqua_prod_raw <- 
   read_csv(file.path(fishstat_dir, "aqua-prod-raw.csv")) %>% 
  clean_names() %>% 
  rename(country= country_name,
         species = asfis_species_name,
         area = fao_major_fishing_area_name,
         environment = environment_name) %>% 
   dplyr::select(-unit_name) %>% 
   filter(!country %in% c("Totals - Tonnes - live weight", "FAO. 2021. Fishery and Aquaculture Statistics. Global aquaculture production 1950-2019 (FishstatJ). In: FAO Fisheries Division [online]. Rome. Updated 2021. www.fao.org/fishery/statistics/software/fishstatj/en"))
)


flags <- aqua_prod_raw %>% 
  dplyr::select(-starts_with("x")) %>% 
  pivot_longer(names_to = "flag", values_to = "symbol", -c(country, species, area, environment, unit)) %>% 
  mutate(flag = case_when(symbol == "..." ~ "No data",
                           symbol == " " ~ "Data not separately available",
                           symbol == "-" ~ "Nil or zero",
                           symbol == "0 0" ~ "0<x<0.5",
                           symbol == "E" ~ "estimate",
                           is.na(symbol) ~ "Reported")) %>% 
  dplyr::select(-symbol)


#sorts country coding to deal with non-UTF characters that country code depends on
Encoding(aqua_prod_raw$country) <- "latin1" #deals with the non-UTF
aqua_prod_raw$country <- iconv(aqua_prod_raw$country, "latin1", "UTF-8",sub='')



aquaculture_prod <- 
  
  bind_cols(
    aqua_prod_raw %>%
      dplyr::select(-c(starts_with("s_"), s)) %>% 
      pivot_longer(names_to = "year", values_to = "value", cols = -c(country, species, area, environment, unit)) %>% 
      mutate(iso_3c = countrycode(country, origin = "country.name", destination = "iso3c", warn = TRUE)) %>%
      mutate(iso_3c = case_when(country == "Zanzibar" ~ "TZA",
                                TRUE ~ iso_3c)) %>%
      mutate(year = gsub("x", "", year) %>% 
               as.numeric),
    
    flags %>% dplyr::select(flag)
  ) %>% 
  drop_na(iso_3c)



saveRDS(object = aquaculture_prod, file = here("data", "tidy_data", "aquaculture_production_tidy.rds"))



```

Unzip FAOSTAT Production data
```{r}
unzip(list.files(path=file.path(rdsi_dir, "raw_data/fao/FAOSTAT_2022"), pattern = "Production_Crops", full.names = TRUE), exdir = file.path(rdsi_dir, "raw_data/fao/FAOSTAT_2022"))

```

Tidy FAOSTAT Production data
```{r}

production_raw <- read_csv(file.path(rdsi_dir, "raw_data/fao/FAOSTAT_2022/Production_Crops_Livestock_E_All_Data_(Normalized).csv"))

production_flags <- read_csv(file.path(rdsi_dir, "raw_data/fao/FAOSTAT_2022/Production_Crops_Livestock_E_Flags.csv"))

production_raw <- production_raw |> 
  left_join(production_flags) |> 
  mutate(Description = if_else(is.na(Description), true = "Official data", false = Description)) |> 
  mutate(Sector = case_when(grepl("Milk|milk|Meat|meat|Fat|fat|Chicken|Buffaloes|Cattle|cattle|Camel|ffal|Pigs|abbit|worm|odent|urkey|oghurt|Wool|Sheep|Goat|Egg|egg|Asses|Beeh|Bees|Ducks|Cream|Geese|Horses|Honey|Lard|Mules|Birds|Skins|Butter|Cheese|Snail", Item) ~ "Livestock",
                            TRUE ~ "Crops"))

Encoding(production_raw$Area) <- "latin1" #deals with the non-UTF
production_raw$Area <- iconv(production_raw$Area, "latin1", "UTF-8",sub='')


#separate by sector
production_sector_list <- 
  production_raw  |> 
  clean_names()  |>  
  mutate(iso3c = countrycode(area, origin = "country.name", destination = "iso3c", warn = TRUE)) |> 
  group_split(sector)

#save
map(.x = production_sector_list, .f = function(this_element){
  saveRDS(object = this_element, file = sprintf(here("data/tidy_data/production-data/%s_production_tidy.rds"), tolower(unique(this_element$sector))))
})
  



```
Create a base raster
```{r}

base_raster <- raster()
values(base_raster) <- 1:ncell(base_raster)
base_raster_ea <- projectRaster(base_raster, crs = equal_area_gp_proj)
res(base_raster_ea) <- 10000
values(base_raster_ea) <- 1:ncell(base_raster_ea) 

writeRaster(x = base_raster_ea, file = here("data/spatial/base_raster_gall_peters.tif"))




```

