---
title: "03_calculate_embodied_crop_materials_from_feed_demand"
output: html_document
---

Next step is convert crop ingredient demand into raw material demand which will vary by processing technique (e.g. SPC vs SBM) and source. So first, we need to know where the ingredients are largely sourced from. THis may change with information that we gain from Biomar but the idea is eventually to make this flexible enough to select any from the top producing countries. This requires knowing the raw material that supports the ingredient and the major places this raw material is grown.




```{r}
library(tidyverse)
library(here)
library(janitor)
library(countrycode)
library(data.table)
library(dtplyr)
library(raster)
library(terra)
library(sf)
library(jsonlite)
library(ggnewscale)

source(here("src/directories.R"))

select <- dplyr::select
values <- terra::values
```


#CROP INGREDIENTS

Have added codes and FAO_names for plant ingredients in teams so join these to production data. 

```{r}

ingredient_by_raw_material <- 
  read_csv(here("data/tidy_data/diets/plant_ingredient_codes.csv")) |> 
  select(-reference) |> 
  drop_na(ingredient) |> 
  add_row(ingredient = "pea starch", raw_material = "peas", FAOSTAT_name = "Peas, green", FAO_code = 187, map_spam_code = "opul") |> 
  mutate(FAOSTAT_name = if_else(FAOSTAT_name == "Peas, dry; Peas, green", true = "Peas, dry", false = FAOSTAT_name)) 
 

crop_production_raw <- readRDS(file = here("data/tidy_data/production-data/crops_production_tidy.rds"))

(crop_production <- 
  crop_production_raw |>
  filter(element == "Production"  & area!= "China" & item %in% ingredient_by_raw_material$FAOSTAT_name & year %in% seq(from = 2015, to = 2019)) |> 
  group_by(area, item) |> 
  summarise(mean_value = mean(value, na.rm = TRUE)) |> 
    ungroup() |> 
  mutate(area = case_when(grepl("Ivoire", area) ~ "Cote d'Ivoire",
                          TRUE ~ area),
         iso3c = countrycode(sourcevar = area, origin = "country.name", destination = "iso3c", warn = TRUE)) |> 
  filter(!is.na(iso3c)) |> 
    arrange(item,-mean_value)
)


#need to decide whether we should pick the top 10 producers for things like guar so countries like pakistan are included for pulses nes
top_10_crop_producers <- 
  crop_production |> 
  group_split(item) |> 
  map_df(\(.){ 
    #this takes the top ten producers - unless it is for Guar (Pulses nes) where I used the suggestion that most guar production comes from India, Pakistan and Africa (so subsequently took the top 8 African producers of pulses nes to make up the top 10)
    
    if(unique(.$item) == "Pulses nes"){
      . |> filter(area %in% c("India", "Pakistan", "Kenya", 
                              "Ethiopia", "United Republic of Tanzania", 
                              "Sudan", "Sierra Leone", "Nigeria", 
                              "Guinea",  "Central African Republic"))
      }else{
        . |>  slice(1:10)}
  }
)


write.csv(x = top_10_crop_producers, file = here("data/tidy_data/production-data/top_crop_producers.csv"))
saveRDS(object = top_10_crop_producers, file = here("data/tidy_data/production-data/top_crop_producers.rds"))


```


Now we have the top 10 producers for each crop that is used as a raw material in feed, we need to get conversion factors (yields) from ingredients to  each of these raw materials in different places and also the gross energy values of each ingredient and co-product so that energetic allocation can be achieved. This was done separately in the Sustainable Aquafeeds Project Teams interface. 

```{r}
#import codes that will be compatible with MAPSPAM layers

plant_ingredient_codes <- 
  read_csv(here("data/tidy_data/diets/plant_ingredient_codes.csv")) |> 
  select(-reference) |> 
  drop_na(ingredient) |> 
  #add_row(ingredient = "pea starch", raw_material = "peas", FAOSTAT_name = "Peas, green", FAO_code = 187, map_spam_code = "opul") |> this row is for if you want pea starch to come from green peas and dried peas - just using dried peas for simplicity
  mutate(FAOSTAT_name = if_else(FAOSTAT_name == "Peas, dry; Peas, green", true = "Peas, dry", false = FAOSTAT_name),
         ingredient = if_else(ingredient == "canola/camelina oil", true = "canola oil", false = ingredient))

#join codes to top 10 producers and conversion factors data
(top_10_crop_w_cf <- read_csv(here("data/tidy_data/production-data/top_crop_producers_conversions.csv")) |> 
    select(-X7, -mean_value) |> 
  left_join(plant_ingredient_codes |> select(-raw_material), by = c("ingredient", "item" = "FAOSTAT_name"))
)

#get yield data from main ingredients
cf_data <- read_csv(here("data/tidy_data/production-data/top_crop_producers_conversions.csv")) |> 
  mutate(ingredient = if_else(ingredient == "canola/camelina oil", true = "canola oil", false = ingredient)) 

# get coproduct yield data
cf_coproduct_data <- read_csv(here("data/raw_data/allocation/coproduct_conversions.csv"))

#join product and coproduct yield (conversion) data
all_cf_data <- 
  cf_data |> 
  select(-X7) |> 
  filter(!ingredient %in% c("tapioca starch", "lupins")) |> 
  left_join(cf_coproduct_data, by = c("ingredient", "iso3c"="country"))


#bring in the gross energy values
ge_values <- read_csv(here("data/raw_data/allocation/feed_coproduct_ge_allocation.csv")) |> 
  select(-c(unit,reference, notes)) |> 
  filter(!ingredient %in% c("fishmeal", "fish oil")) |> 
  mutate(coproduct = if_else(coproduct == "soybean meal" & ingredient == "soybean meal", true = "soy protein concentrate", false = coproduct ))

(allocation_raw <- 
  ge_values |> 
  left_join(all_cf_data, by = c("ingredient", "coproduct")) |> 
  mutate(energy_cf_product = gross_energy_product*cf,
         energy_cf_coproduct = gross_energy_coproduct*coproduct_cf) |> 
  ungroup() |> 
  group_by(ingredient, area) |> 
  nest() |> 
  mutate(allocation_factor = map(data, ~(.$gross_energy_product/ (sum(unique(.$energy_cf_product), sum(.$energy_cf_coproduct)))))) |> 
  unnest(c(data, allocation_factor)))

write_csv(x = allocation_raw, file = here("data/tidy_data/allocation/crop_ingredient_allocation_factors.csv"))

```


Join the ingredient demand and allocation data to generate crop raw material demand

```{r}

#get saved allocation data
(allocation_factors <- read_csv(here("data/tidy_data/allocation/crop_ingredient_allocation_factors.csv")) |> 
  select(item, ingredient, area, iso3c, mean_value, allocation_factor) |> 
  left_join(plant_ingredient_codes |> select(-raw_material), by = c("ingredient", "item" = "FAOSTAT_name")) |> 
  distinct()
)

#join the ingredient demand by diet to the top 10 producer data so for each ingredient there will be an raw material demand per source, and diet scenario. Create new raw_material demand column by dividing total ingredient demand by the conversion factor (i.e. the ingredient extraction rate from the raw material in the first place)

crop_raw_material_demand <- readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> 
  ungroup() |> 
  left_join(allocation_factors, by = c("ingredients" = "ingredient")) |> 
  drop_na(item) |> 
  rename(FAOSTAT_name = item,
         source_country = area,
         source_iso3c = iso3c) |> 
  mutate(total_crop_demand = total_ingredient_demand*allocation_factor) |> 
  mutate(ingredients = if_else(ingredients == "canola/camelina oil", true = "canola oil", false = ingredients)) #to prevent file_path issues later

#export total crop demand for a woring data source
saveRDS(object = crop_raw_material_demand, file = here("data/tidy_data/demand/total_crop_demand.rds"))


rm(list = ls(all.names = TRUE))
```
