---
title: "Calculate embodied fish demand in feeds"
output: html_document
---

```{r}
library(tidyverse)
library(here)
library(janitor)
library(countrycode)
library(data.table)
library(dtplyr)
library(raster)
library(terra)
library(sf)
library(jsonlite)
library(ggnewscale)

source(here("src/directories.R"))

select <- dplyr::select
values <- terra::values
```


#FORAGE FISH INGREDIENTS

Now we need to work out how the ingredient demand converts to forage fish biomass. This will depend on species which depends on geographic region. So we need different conversions from different species. First bring in the demand data and the data from Kok et al re embodied fisg depending on species. This also requires the species list for forage fish.
  

```{r}
#bring in total ingredient demand data and isolate fishmeal and oil demand

total_ingredient_demand <- readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds"))

fmfo_demand <- total_ingredient_demand |> filter(grepl("Marine", groups))

rm(total_ingredient_demand)


(embodied_fish_ratios <- read_csv(here("data/raw_data/allocation/embodied_fish_ge_allocation_tidy.csv")) |> 
  filter(common_name != "Gross energy") |> 
  mutate(ecosystem_combined = case_when(grepl("IS|NWS|NS|BS", ecosystem) ~ "North Atlantic",
                                        ecosystem == "HC" ~ "Humboldt",
                                        ecosystem == "GM" ~ "Gulf of Mexico", 
                                        ecosystem == "SO" ~ "Southern Ocean",
                                        ecosystem == "GA" ~ "Other")) |> 
  select(common_name, sci_name, ecosystem_combined, ingredient, value) |> 
  mutate(sci_name = case_when(sci_name == "Brevoorti patronus" ~ "Brevoortia patronus",
                              TRUE ~ sci_name)) |> distinct())



#global values for later
globalav_embodiedfish_fm <- embodied_fish_ratios |> filter(common_name == "Global average" & ingredient == "Fishmeal") |> pull(value)
globalav_embodiedfish_fo <- embodied_fish_ratios |> filter(common_name == "Global average" & ingredient == "Fish oil") |> pull(value)


forage_catch <- readRDS(here("data/raw_data/fisheries/spatial_forage_catch_2017.rds")) |> filter(!grepl("krill", CommonName))

(forage_catch_summary <- forage_catch |> 
  group_by(TaxonName, CommonName) |> 
  summarise(total_catch = sum(total_catch, na.rm=TRUE)) |> 
  arrange(-total_catch) |> 
  left_join(embodied_fish_ratios |> filter(ingredient == "Fishmeal") |> select(-c(ingredient, ecosystem_combined)) |> rename(fishmeal_cf =value), by = c("TaxonName" = "sci_name")) |> 
    left_join(embodied_fish_ratios |> filter(ingredient == "Fish oil") |> select(-c(ingredient, ecosystem_combined)) |> rename(fish_oil_cf =value), by = c("TaxonName" = "sci_name", "common_name"))
)



#plot forage catch to identify main regions for production. The areas I'm interested in is the Humboldt system, the caribeban coast off latin america, the gulf of mexico, the north atlantic, west africa, the western INdian Ocean (India, Red Sea) the east asia regions
forage_xyz <- forage_catch |> group_by(LonCentre, LatCentre) |> 
  summarise(total_catch = sum(total_catch)) |> 
  #select(LonCentre, LatCentre, total_catch) |> 
  rename(x=LonCentre, y= LatCentre, z = total_catch) 

total_forage_raster <- rast(forage_xyz, type="xyz")

plot(log10(total_forage_raster+1))

sum(values(total_forage_raster), na.rm = TRUE)

#CROP THE FISHING REGIONS BY FAO AREA

fao_areas <- c(77, 67, 87, 21, 31, 41, 27, 37, 34, 47, 51,57, 61, 71, 81) #these fao areas broadly correspond with the area of most interest above

#read in the fao area shapefiles from the FAO json
FAO_regions_shp <- read_sf("https://www.fao.org/fishery/geoserver/fifao/ows?service=WFS&request=GetFeature&version=1.0.0&typeName=fifao:FAO_AREAS_CWP_NOCOASTLINE&outputFormat=json") |> filter(F_LEVEL == "MAJOR") |> 
  mutate(F_AREA = as.numeric(F_AREA)) |> 
  mutate(source_code = case_when(NAME_EN == "Atlantic, Northeast" ~ "NAT",
                                 NAME_EN %in% c("Pacific, Northeast", "Pacific, Eastern Central", "Pacific, Southeast") ~ "EPC",
                                                      TRUE ~ "OTH"))

#check FAO shapefile appears correctly
ggplot()+
  geom_sf(data = FAO_regions_shp, 
          aes(fill = F_AREA))


#get the xyz file for each fao region sand create one data frame to join to the forage catch data by coordinates

fao_xyz_of_interest <- map_df(.x = fao_areas, .f = \(this_area){
  
  this_area_shp <- FAO_regions_shp |> filter(F_AREA == this_area)
  
  this_area_raster <- terra::rasterize(x = vect(this_area_shp), y = total_forage_raster, field = this_area_shp$F_AREA)
  
  this_area_xyz <- terra::as.data.frame(x = this_area_raster, xy = TRUE)  |> mutate(area_name = this_area_shp$NAME_EN)
  
  return(this_area_xyz)
  
})

#check the regions
fao_areas_of_interest <- unique(fao_xyz_of_interest$area_name)

#plot the FAO regions used for a supplementary figure
forage_regions_shp <- FAO_regions_shp |> filter(F_AREA %in% fao_areas)
world_shp <- rnaturalearth::ne_countries(scale = 110, returnclass = "sf")
catch_raster <- total_forage_raster |> terra::as.data.frame(xy=TRUE)


ggplot()+
  geom_sf(data = forage_regions_shp |> mutate(f_area = factor(F_AREA)), aes(fill = NAME_EN), colour=NA)+
  geom_sf(data = world_shp, size=NA, fill="grey60")+
  labs(fill = "")+
  scale_fill_manual(values = colorRampPalette(RColorBrewer::brewer.pal(7, name = "Blues"))(length(fao_areas)))+
   theme_void()+
  theme(
        panel.background = element_rect(fill= "white", colour = NA),
        plot.background = element_rect(fill= "white", colour = NA),
        axis.title = element_blank(),
        legend.text = element_text(size=6),
        legend.title = element_text(size=6),
        legend.key.size = unit(0.3, "cm"))+
  new_scale_fill()+
  geom_tile(data = catch_raster, aes(x = x, y = y, fill=log10(z+1)))+
  scale_fill_gradientn(colours = RColorBrewer::brewer.pal(9, name = "YlOrRd"))+
  labs(fill = bquote(log[10]~Catch))+
   theme_void()+
  theme(
        panel.background = element_rect(fill= "white", colour=NA),
        plot.background = element_rect(fill= "white", colour = NA),
        axis.title = element_blank(),
        legend.text = element_text(size=4),
        legend.title = element_text(size=4),
        legend.key.size = unit(0.2, "cm"))
  
  
ggsave(filename = here("figures/supplementary/fao_regions_catch.jpeg"), device = "jpg", dpi = 300, width = 13, height = 5, units = "cm")


#join FAO area xyz of interest to catch data and  embodied fish factors to get the potential fmfo provisioning by each area
fao_area_production_potential <- 
  forage_catch |> 
  left_join(fao_xyz_of_interest |> rename(fao_area_code = layer, fao_area_name = area_name), 
            by = c("LonCentre" = "x", "LatCentre" = "y")) |> 
  drop_na(fao_area_name) |>
  group_by(fao_area_name) |> 
  summarise(total_catch = sum(total_catch, na.rm=TRUE)) |> 
  mutate(all_catch = sum(total_catch)) |>  # from this Atlantic Northeast, Eastern Pacific (all regions), and "Other" seem like the best options
  mutate(source_code = case_when(fao_area_name == "Atlantic, Northeast" ~ "NAT",
                                 fao_area_name %in% c("Pacific, Northeast", "Pacific, Eastern Central", "Pacific, Southeast") ~ "EPC",
                                                      TRUE ~ "OTH")) |> 
  ungroup()


#the new umbrella groups - these catch levels are relatively even as they can be
source_production_potential <- 
  fao_area_production_potential |> 
  group_by(source_code, all_catch) |> 
  summarise(total_catch = sum(total_catch))



#now distribute fishmeal and oil demand by the proportion of catch the species make up in each source region
forage_catch_w_prop_supply <- 
  forage_catch |> 
  left_join(fao_xyz_of_interest |> rename(fao_area_code = layer, fao_area_name = area_name), 
            by = c("LonCentre" = "x", "LatCentre" = "y")) |> 
  drop_na(fao_area_name) |>
  mutate(source_code = case_when(fao_area_name == "Atlantic, Northeast" ~ "NAT",
                                 fao_area_name %in% c("Pacific, Northeast", "Pacific, Eastern Central", "Pacific, Southeast") ~ "EPC",
                                                      TRUE ~ "OTH")) |> 
  group_by(source_code) |> 
  nest() |> 
  mutate(prop_source_catch = map(data, ~(.$total_catch/sum(.$total_catch)))) |> 
  unnest(cols = c(data, prop_source_catch)) |> 
  ungroup() |> 
  left_join(embodied_fish_ratios |> filter(ingredient == "Fishmeal") |> select(-c(ingredient, ecosystem_combined)) |> rename(fishmeal_cf=value), by = c("TaxonName" = "sci_name")) |> 
    left_join(embodied_fish_ratios |> filter(ingredient == "Fish oil") |> select(-c(ingredient, ecosystem_combined)) |> rename(fish_oil_cf=value), by = c("TaxonName" = "sci_name", "common_name")) |> 
  mutate(fishmeal_cf = case_when(is.na(common_name) ~ globalav_embodiedfish_fm,
                                 TRUE ~ fishmeal_cf),
         fish_oil_cf = case_when(is.na(common_name) ~ globalav_embodiedfish_fo,
                                 TRUE ~ fish_oil_cf)) |> 
  select(-common_name) 
 


#use the proportions for each region to divide the demand up and work out how the spp_based demand converts to embodied fish

forage_supply_demand_by_diet <- 
  map_df(.x =  c("marine_diet", "plant_diet"), .f = \(this_diet){
    
    this_df <- forage_catch_w_prop_supply |> 
      mutate(diet = this_diet,
             fm_demand = fmfo_demand |> filter(diet == this_diet & ingredients == "fishmeal-wild") |> pull(total_ingredient_demand),
             fo_demand = fmfo_demand |> filter(diet == this_diet & ingredients == "fish oil-wild") |> pull(total_ingredient_demand),
             spp_fm_demand = fm_demand * prop_source_catch,
             spp_fo_demand = fo_demand * prop_source_catch,
             fm_embodied_fish = spp_fm_demand * fishmeal_cf,
             fo_embodied_fish = spp_fo_demand * fish_oil_cf,
             total_embodied_fish = fm_embodied_fish+fo_embodied_fish) |> 
      mutate(demand_vs_supply = total_embodied_fish/total_catch) |> 
      relocate(diet, .before = source_code)
    
  })
```


Now do some checks on if demand and supply match up per record i.e. by per cell-species-gear combinations. 

```{r}
#check the demand vs supply match up - are there cells where supply is outstripped by demand? 
forage_supply_demand_by_diet |> filter(demand_vs_supply>1) #lots of rows where demand exceeds supply - likely due to global average conversions, but by how much typically?
nrow(forage_supply_demand_by_diet |> filter(demand_vs_supply>1))/nrow(forage_supply_demand_by_diet) # in 6.7% of catch records demand exceed catch in 2017
forage_supply_demand_by_diet |> filter(demand_vs_supply>1) |> pull(demand_vs_supply) |> unique() #nothing exceeding 1.3 of catch. How much does this matter- i.e. what does that equate to in tonnes and is that within a range of variation for that cell?
nrow(forage_supply_demand_by_diet |> filter(demand_vs_supply>1.1))/nrow(forage_supply_demand_by_diet) # the same 2.8% of records because this is the only value

#look at what quantities and proportion demand exceeded supply by
exceedance_check <- forage_supply_demand_by_diet |> 
        mutate(exceedance = case_when(demand_vs_supply>1 ~ (demand_vs_supply-1)*total_embodied_fish,
                                      TRUE ~ total_embodied_fish-total_catch)) |>
   mutate(prop_exceedance = case_when(exceedance>0 ~ exceedance/total_catch,
                                      TRUE ~ 0))

#what is the maximum exceedance in tonnage?
max_exceedance <- exceedance_check |> pull(exceedance) |> max() #25193 tonnes - could be due to artisanal catch
#but what does the catch look like when max absolute exceedance occurs?
exceedance_check |> filter(exceedance == max_exceedance) # catch is 64000 tonnes when demand exceeds catch by as much as a lot
#what is the maximum proportional exceedance?
max_prop_exceedance <- exceedance_check |> pull(prop_exceedance) |> max() #38% this is not great
#what is the catch when proportional exceedance is this high?
exceedance_check |> filter(prop_exceedance == max_prop_exceedance) # catch is tiny when proportional exceedance is maximised 
 
#plot it out to chekc the pattern
exceedance_only_df <- 
  exceedance_check |> filter(demand_vs_supply>1)

#no  pattern to the relationshoip between size and prop of exceedance  so will have redistribute - seems the global value for FMFO embodied fish creates a constant exceedance because the values for the cell are so small and close together the proportional difference in the cell is accruate to several dp?
ggplot(data = exceedance_only_df)+
  aes(x=prop_exceedance, y = exceedance+1)+
  geom_point(alpha=0.3)+
  scale_y_continuous(trans = "log10")




```

Redistribute the demand in cells where it is greater than supply across other cells where there is available deficit. How to allocate can be scaled by the proportion that cell accounts for in terms of the total deficit.  Do this independently for each diet-source combo though, so need to split the data.

```{r}
#data list by source and diet
forage_supply_demand_list <- forage_supply_demand_by_diet |> 
  relocate(diet, .before = source_code) |> 
  group_split(diet, source_code)


forage_supply_demand_redistributed <- 
  map_df(.x = forage_supply_demand_list, .f = \(this_df){
    
    exceed_supply <- this_df |> filter(demand_vs_supply>1)
    below_supply <- this_df |> filter(demand_vs_supply<=1)
    
    this_diet_source <-  paste0(unique(this_df$diet),"_",unique(this_df$source_code))
    
    if(nrow(exceed_supply)>0){
      
      message(this_diet_source)
      
      exceed_supply2 <-  exceed_supply |>
        mutate(exceedance = (demand_vs_supply-1)*total_embodied_fish)
      
      total_exceedance <- sum(exceed_supply2$exceedance)
      total_deficit <- below_supply |> mutate(exceedance = total_embodied_fish-total_catch) |> pull(exceedance) |> sum()
      
      message("Total deficit is greater than total exceedance? ", abs(total_deficit)>total_exceedance)
      
      below_supply2 <- below_supply |> 
        mutate(exceedance = total_embodied_fish-total_catch) |> 
        mutate(prop_exceedance = exceedance/sum(exceedance)) |> 
        mutate(added_embodied_fish = prop_exceedance * total_exceedance) |> 
        mutate(adj_embodied_fish = total_embodied_fish+added_embodied_fish)
      
      message("Total biomass from exceedance cells is equal to that added to deficit cells = ", round(sum(below_supply2$added_embodied_fish), digits = 0 )==round(total_exceedance, digits = 0))
      
      exceed_supply3 <-  exceed_supply2 |> 
        mutate(prop_exceedance = NA) |> 
        mutate(added_embodied_fish = NA) |> 
        mutate(adj_embodied_fish = total_embodied_fish-exceedance)
      
      diff_after_adjustment <- sum(exceed_supply3$total_embodied_fish)-sum(exceed_supply3$adj_embodied_fish)
      
      message("Total biomass removed from exceedance cells is equal to the difference between the original and adjusted embodied fish = ", round(diff_after_adjustment, digits = 3)==round(total_exceedance, digits = 3))
      
      
      message("Returns same number of rows as original df? ",  (nrow(exceed_supply3)+nrow(below_supply2))==nrow(this_df))
      
      this_df_adj <- bind_rows(below_supply2, exceed_supply3)
      
      return(this_df_adj)
      
    } else{
      
        this_df <- this_df |> 
        mutate(exceedance = 0) |> 
        mutate(prop_exceedance = 0) |> 
        mutate(added_embodied_fish = 0) |> 
        mutate(adj_embodied_fish = total_embodied_fish)
      
    return(this_df)
      
    }

  })


fwrite(forage_supply_demand_redistributed, file = here("data/tidy_data/demand/embodied_fish_per_cell.csv"))


#check same number of rows as the input file and that totals are the same in combination and by diet

nrow(forage_supply_demand_by_diet)==nrow(forage_supply_demand_redistributed)

sum(forage_supply_demand_by_diet$total_embodied_fish) == sum(forage_supply_demand_redistributed$adj_embodied_fish, na.rm = TRUE)

#check marine diet - yep total embodied fish is a match between the original and redistributed sets
forage_supply_demand_by_diet |> filter(diet=="marine_diet") |> pull(total_embodied_fish) |> sum() == forage_supply_demand_redistributed |> filter(diet=="marine_diet") |> pull(adj_embodied_fish) |> sum()

#check plant diet - yep total embodied fish is a match between the original and redistributed sets
forage_supply_demand_by_diet |> filter(diet=="plant_diet") |> pull(total_embodied_fish) |> sum() == forage_supply_demand_redistributed |> filter(diet=="plant_diet") |> pull(adj_embodied_fish) |> sum()


embodied_fish_demand_by_diet_source <- forage_supply_demand_redistributed |> 
  group_split(diet, source_code) |> 
  map_df(\(this_df){
    
    return(tibble(diet = unique(this_df$diet), source_code = unique(this_df$source_code), embodied_fish_demand = sum(this_df$adj_embodied_fish)))
    
  })

saveRDS(object = embodied_fish_demand_by_diet_source, file = here("data/tidy_data/demand/embodied_fish_demand_by_source_diet.rds"))




#For the East Pacific, supply falls short of demand based on catch for the marine diet - look into this here

marine_EPC_checks <- 
  forage_supply_demand_redistributed |> filter(diet == "marine_diet" & source_code == "EPC") |> #78,466 rows 
  mutate(adj_demand_vs_supply = adj_embodied_fish/total_catch) |> 
  filter(adj_demand_vs_supply >1) |> #51300 rows have exceedance (most)
  mutate(actual_exceedance_adj = adj_embodied_fish-total_catch) |> 
  mutate(prop_exceedance_adj = actual_exceedance_adj/total_catch)

max_exceedance <- max(marine_EPC_checks$actual_exceedance_adj) #can exceed 2017 catch levels by 2400 tonnes
max_prop_exceedance_adj <- max(marine_EPC_checks$prop_exceedance_adj) #max exceedance is 24% levels

#Where it exceeds by the max value in tonnes, this is a 23% increase
marine_EPC_checks |> filter(actual_exceedance_adj == max_exceedance)
median(marine_EPC_checks$actual_exceedance_adj)
median(marine_EPC_checks$prop_exceedance_adj)

# 2 tonnes exceeded at the 90th percentile.

sequence <- c(seq(from = 0.25, 0.99, by =0.05), 0.96,0.97,0.98, 0.99, 0.999)
exceedance_list <- vector(length = length(sequence))

for(i in 1:length(sequence)){
  exceedance_list[[i]] <- quantile(marine_EPC_checks$actual_exceedance_adj,sequence[i])
}

#Given the variability of forage fish population biomass, for most catch (95% and 98% respectively) to be within 7 tonnes and 70 tonnes of 2017 catch and at no point exceeded 25% of that years catch, this seems reasonable.

ggplot(data = tibble(percentile = sequence, exceedance_tonnes = exceedance_list))+
  aes(x = percentile, y = exceedance_tonnes)+
  geom_point()+
  theme_bw()+
  labs(x = "Exceedance percentile", y= "Exceedance (tonnes)")+
  theme(text = element_text(size=8))


ggsave(filename = here("figures/supplementary/cell_catch_exceedance_for_marinediet_EPC.jpg"), dpi=600, width = 9, height = 6, units = "cm")


```
  
Create an FMFO embodied fish demand by source and ingredient

```{r}
embodied_fish_by_dietsource <- readRDS(file = here("data/tidy_data/demand/embodied_fish_demand_by_source_diet.rds"))

embodied_fish_by_dietsource |> 
  ggplot()+
  aes(x = diet, y = embodied_fish_demand/1000, fill = source_code)+
  geom_bar(stat= "identity", position = "dodge")+
  labs(fill = "Source", y = "Embodied fish demand (1000s T)")+
  scale_fill_manual(values = c("#edf8b1", "#7fcdbb", "#2c7fb8"))+
  theme_bw()+
  theme(legend.position = c(0.8, 0.8),
        legend.background = element_rect(fill="transparent"), 
        axis.title.x = element_blank(),
        panel.grid = element_blank(), 
        text = element_text(size=12),
        legend.text = element_text(size = 10))


ggsave(filename = here("figures/presos/embodied_forage_fish_demand.jpg"), device="jpg", dpi = 600, width = 10, height = 10, units="cm")

```

