---
title: "Calculate embodied fish demand in feeds"
output: html_document
---

```{r}
library(tidyverse)
library(here)
library(janitor)
library(countrycode)
library(data.table)
library(dtplyr)
library(raster)
library(terra)
library(sf)
library(jsonlite)
library(ggnewscale)

source(here("src/directories.R"))

select <- dplyr::select
values <- terra::values
```


#FORAGE FISH INGREDIENTS

Now we need to work out how the ingredient demand converts to forage fish biomass. This will depend on species which depends on geographic region. So we need different conversions from different species. First bring in the demand data and the data from Kok et al re embodied fisg depending on species. This also requires the species list for forage fish.
  

```{r}
#bring in total ingredient demand data and isolate fishmeal and oil demand

total_ingredient_demand <- readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds"))

fmfo_demand <- total_ingredient_demand |> filter(grepl("Marine", groups))

rm(total_ingredient_demand)

(embodied_fish_ratios <- read_csv(here("data/raw_data/allocation/embodied_fish_allocation_tidy.csv")) |> 
  filter(common_name != "Gross energy") |> 
  mutate(ecosystem_combined = case_when(grepl("IS|NWS|NS|BS", ecosystem) ~ "North Atlantic",
                                        ecosystem == "HC" ~ "Humboldt",
                                        ecosystem == "GM" ~ "Gulf of Mexico", 
                                        ecosystem == "SO" ~ "Southern Ocean",   
                                        ecosystem == "GA" ~ "Other")) |> 
  select(common_name, sci_name, ecosystem_combined, ingredient, ge_value, mass_value) |> 
  mutate(sci_name = case_when(sci_name == "Brevoorti patronus" ~ "Brevoortia patronus",
                              TRUE ~ sci_name)) |> distinct())



#global values for later
globalav_embodiedfish_fm_mass <- embodied_fish_ratios |> filter(common_name == "Global average" & ingredient == "Fishmeal") |> pull(mass_value)
globalav_embodiedfish_fm_ge <- embodied_fish_ratios |> filter(common_name == "Global average" & ingredient == "Fishmeal") |> pull(ge_value)

globalav_embodiedfish_fo_mass <- embodied_fish_ratios |> filter(common_name == "Global average" & ingredient == "Fish oil") |> pull(mass_value)
globalav_embodiedfish_fo_ge <- embodied_fish_ratios |> filter(common_name == "Global average" & ingredient == "Fish oil") |> pull(ge_value)
```

Isolate catch for trimmings first
```{r}

trimmings_spp_catch <- readRDS(here("data/large_data/trimmings_spp_catch.rds")) |> filter(IYear == 2017)

trimmings_xyz <- trimmings_spp_catch |> group_by(LonCentre, LatCentre) |> 
  summarise(total_catch = sum(ReportedIND)) |> 
  #select(LonCentre, LatCentre, total_catch) |> 
  rename(x=LonCentre, y= LatCentre, z = total_catch) 


total_trimmings_raster <- rast(trimmings_xyz, type="xyz")

plot(total_trimmings_raster)


#FAO areas of most prominence from Biomar sustainability report 2022 for both wild caught and trimmings https://www.biomar.com/globalassets/.global/sustainability-report/biomar-global-sustainability-report-2022.pdf

#bring in the FAO regions shape file

fao_areas_shp <- read_sf("https://www.fao.org/fishery/geoserver/fifao/ows?service=WFS&request=GetFeature&version=1.0.0&typeName=fifao:FAO_AREAS_CWP_NOCOASTLINE&outputFormat=json") |> filter(F_LEVEL == "MAJOR") |> 
  mutate(F_AREA = as.numeric(F_AREA)) |> 
  group_by(F_AREA) |> 
  group_split()

this_area <- fao_areas_shp[[1]]

#get xyz file for the FAO areas in 0.5 resolution (catch data resolution)

fao_xyz <- 
  
  map_df(.x = fao_areas_shp, .f = \(this_area){
  
  this_area_shp <- this_area
  
  this_area_raster <- terra::rasterize(x = vect(this_area_shp), y = rast(res = 0.5), field = this_area_shp$F_AREA)
  
  this_area_xyz <- terra::as.data.frame(x = this_area_raster, xy = TRUE)  |> mutate(fao_area_name = this_area_shp$NAME_EN)
  
  return(this_area_xyz)
  
}) |> rename(fao_area_code = layer,
             LonCentre = x,
             LatCentre = y)



#join catch and fao xyz to identify the catch for species used in trimmings across FAO regions

trimmings_spp_fao_area <- 
  trimmings_spp_catch |> 
  left_join(fao_xyz) |> 
  group_by(fao_area_name, fao_area_code) |> 
  summarise(trim_spp_catch = sum(ReportedIND, na.rm=TRUE)) |> 
  ungroup() |> 
  arrange(-trim_spp_catch) |> 
  mutate(prop_trimmings = trim_spp_catch/sum(trim_spp_catch, na.rm=TRUE))


```



```{r}
forage_catch <- readRDS(here("data/raw_data/fisheries/spatial_forage_catch_2017.rds")) |> filter(total_ind >0)

(forage_catch_summary <- forage_catch |> 
  group_by(TaxonName, CommonName) |> 
  summarise(total_catch = sum(total_catch, na.rm=TRUE)) |> 
  arrange(-total_catch) |> 
  left_join(embodied_fish_ratios |> filter(ingredient == "Fishmeal") |> select(-c(ingredient, ecosystem_combined)) |> rename(fishmeal_cf_mass = mass_value, fishmeal_cf_ge = ge_value), by = c("TaxonName" = "sci_name")) |> 
    left_join(embodied_fish_ratios |> filter(ingredient == "Fish oil") |> select(-c(ingredient, ecosystem_combined)) |> rename(fish_oil_cf_mass = mass_value, fish_oil_cf_ge = ge_value), by = c("TaxonName" = "sci_name", "common_name"))
)





#plot forage catch to identify main regions for production. The areas I'm interested in is the Humboldt system, the caribeban coast off latin america, the gulf of mexico, the north atlantic, west africa, the western INdian Ocean (India, Red Sea) the east asia regions
forage_xyz <- forage_catch |> group_by(LonCentre, LatCentre) |> 
  summarise(total_catch = sum(total_catch)) |> 
  #select(LonCentre, LatCentre, total_catch) |> 
  rename(x=LonCentre, y= LatCentre, z = total_catch) 

total_forage_raster <- rast(forage_xyz, type="xyz")

plot(log10(total_forage_raster+1))

sum(values(total_forage_raster), na.rm = TRUE)

#CROP THE FISHING REGIONS BY FAO AREA

fao_areas_shp <- read_sf("https://www.fao.org/fishery/geoserver/fifao/ows?service=WFS&request=GetFeature&version=1.0.0&typeName=fifao:FAO_AREAS_CWP_NOCOASTLINE&outputFormat=json") |> filter(F_LEVEL == "MAJOR") |> 
  mutate(F_AREA = as.numeric(F_AREA)) |> 
  group_by(F_AREA) |> 
  group_split()

#OLD approach 
# fao_areas <- c(77, 67, 87, 21, 31, 41, 27, 37, 34, 47, 51,57, 61, 71, 81) #these fao areas broadly correspond with the area of most interest above
# 
# #read in the fao area shapefiles from the FAO json
# FAO_regions_shp <- read_sf("https://www.fao.org/fishery/geoserver/fifao/ows?service=WFS&request=GetFeature&version=1.0.0&typeName=fifao:FAO_AREAS_CWP_NOCOASTLINE&outputFormat=json") |> filter(F_LEVEL == "MAJOR") |> 
#   mutate(F_AREA = as.numeric(F_AREA)) 
# |> mutate(source_code = case_when(NAME_EN == "Atlantic, Northeast" ~ "NAT",
#                                  NAME_EN %in% c("Pacific, Northeast", "Pacific, Eastern Central", "Pacific, Southeast") ~ "EPC",
#                                                       TRUE ~ "OTH"))

#check FAO shapefile appears correctly
ggplot()+
  geom_sf(data = FAO_regions_shp, 
          aes(fill = F_AREA))


this_area <- fao_areas_shp[[1]]

#get xyz file for the FAO areas in 0.5 resolution (catch data resolution)

fao_xyz <- 
  
  map_df(.x = fao_areas_shp, .f = \(this_area){
  
  this_area_shp <- this_area
  
  this_area_raster <- terra::rasterize(x = vect(this_area_shp), y = rast(res = 0.5), field = this_area_shp$F_AREA)
  
  this_area_xyz <- terra::as.data.frame(x = this_area_raster, xy = TRUE)  |> mutate(fao_area_name = this_area_shp$NAME_EN)
  
  return(this_area_xyz)
  
}) |> rename(fao_area_code = layer,
             LonCentre = x,
             LatCentre = y)




#check the regions
fao_areas_of_interest <- unique(fao_xyz_of_interest$area_name)

#plot the FAO regions used for a supplementary figure
forage_regions_shp <- FAO_regions_shp |> filter(F_AREA %in% fao_areas)
world_shp <- rnaturalearth::ne_countries(scale = 110, returnclass = "sf")
catch_raster <- total_forage_raster |> terra::as.data.frame(xy=TRUE)


ggplot()+
  geom_sf(data = forage_regions_shp |> mutate(f_area = factor(F_AREA)), aes(fill = NAME_EN), colour=NA)+
  geom_sf(data = world_shp, size=NA, fill="grey60")+
  labs(fill = "")+
  scale_fill_manual(values = colorRampPalette(RColorBrewer::brewer.pal(7, name = "Blues"))(length(fao_areas)))+
   theme_void()+
  theme(
        panel.background = element_rect(fill= "white", colour = NA),
        plot.background = element_rect(fill= "white", colour = NA),
        axis.title = element_blank(),
        legend.text = element_text(size=6),
        legend.title = element_text(size=6),
        legend.key.size = unit(0.3, "cm"))+
  new_scale_fill()+
  geom_tile(data = catch_raster, aes(x = x, y = y, fill=(z)))+
  scale_fill_gradientn(colours = RColorBrewer::brewer.pal(9, name = "YlOrRd"))+
  labs(fill = bquote(Catch~(tonnes)))+
   theme_void()+
  theme(
        panel.background = element_rect(fill= "white", colour=NA),
        plot.background = element_rect(fill= "white", colour = NA),
        axis.title = element_blank(),
        legend.text = element_text(size=4),
        legend.title = element_text(size=4),
        legend.key.size = unit(0.2, "cm"))
  
  
ggsave(filename = here("figures/supplementary/fao_regions_catch.jpeg"), device = "jpg", dpi = 300, width = 13, height = 5, units = "cm")


#join FAO area xyz of interest to catch data and  embodied fish factors to get the potential fmfo provisioning by each area
fao_area_production_potential <- 
  forage_catch |> 
  left_join(fao_xyz_of_interest |> rename(fao_area_code = layer, fao_area_name = area_name), 
            by = c("LonCentre" = "x", "LatCentre" = "y")) |> 
  drop_na(fao_area_name) |>
  group_by(fao_area_name) |> 
  summarise(total_catch = sum(total_catch, na.rm=TRUE)) |> 
  mutate(all_catch = sum(total_catch)) |>  # from this Atlantic Northeast, Eastern Pacific (all regions), and "Other" seem like the best options
  mutate(source_code = case_when(fao_area_name == "Atlantic, Northeast" ~ "NAT",
                                 fao_area_name %in% c("Pacific, Northeast", "Pacific, Eastern Central", "Pacific, Southeast") ~ "EPC",
                                                      TRUE ~ "OTH")) |> 
  ungroup()


saveRDS(object = fao_area_production_potential, file = here("data/tidy_data/production-data/fao_regions_forage_production_potential.rds"))



#the new umbrella groups - these catch levels are relatively even as they can be
source_production_potential <- 
  fao_area_production_potential |> 
  group_by(source_code, all_catch) |> 
  summarise(total_catch = sum(total_catch))



#now distribute fishmeal and oil demand by the proportion of catch the species make up in each source region
forage_catch_w_prop_fmfo_supply <- 
  forage_catch |> 
  left_join(fao_xyz_of_interest |> rename(fao_area_code = layer, fao_area_name = area_name), 
            by = c("LonCentre" = "x", "LatCentre" = "y")) |> 
  drop_na(fao_area_name) |>
  mutate(source_code = case_when(fao_area_name == "Atlantic, Northeast" ~ "NAT",
                                 fao_area_name %in% c("Pacific, Northeast", "Pacific, Eastern Central", "Pacific, Southeast") ~ "EPC",
                                                      TRUE ~ "OTH")) |> 
  left_join(embodied_fish_ratios |> filter(ingredient == "Fishmeal") |> select(-c(ingredient, ecosystem_combined)) |> rename(fishmeal_cf_mass = mass_value, fishmeal_cf_ge = ge_value), by = c("TaxonName" = "sci_name")) |> 
    left_join(embodied_fish_ratios |> filter(ingredient == "Fish oil") |> select(-c(ingredient, ecosystem_combined)) |>  rename(fish_oil_cf_mass = mass_value, fish_oil_cf_ge = ge_value), by = c("TaxonName" = "sci_name", "common_name")) |> 
  mutate(fishmeal_cf_mass = case_when(is.na(common_name) ~ globalav_embodiedfish_fm_mass,
                                 TRUE ~ fishmeal_cf_mass),
         fishmeal_cf_ge = case_when(is.na(common_name) ~ globalav_embodiedfish_fm_ge,
                                 TRUE ~ fishmeal_cf_ge),
         fish_oil_cf_mass = case_when(is.na(common_name) ~ globalav_embodiedfish_fo_mass,
                                 TRUE ~ fish_oil_cf_mass),
         fish_oil_cf_ge = case_when(is.na(common_name) ~ globalav_embodiedfish_fo_ge,
                                 TRUE ~ fish_oil_cf_ge),
         fm_available_ge = total_ind/fishmeal_cf_ge,
         fo_available_ge = total_ind/fish_oil_cf_ge,
         fm_available_mass = total_ind/fishmeal_cf_mass,
         fo_available_mass = total_ind/fish_oil_cf_mass) |> 
  select(-common_name) |> 
  group_by(source_code) |> 
  nest() |> 
  mutate(prop_available_fm_ge = map(data, ~(.$fm_available_ge/sum(.$fm_available_ge, na.rm=TRUE))),
         prop_available_fm_mass = map(data, ~(.$fm_available_mass/sum(.$fm_available_mass, na.rm=TRUE))),
         prop_available_fo_ge = map(data, ~(.$fo_available_ge/sum(.$fo_available_ge,  na.rm=TRUE))),
         prop_available_fo_mass = map(data, ~(.$fo_available_mass/sum(.$fo_available_mass, na.rm=TRUE)))) |> 
  unnest(cols = c(data, prop_available_fm_ge, prop_available_fm_mass, prop_available_fo_ge, prop_available_fo_mass)) |> 
  ungroup() |> 
  filter(TaxonName != "Euphausia superba")
  

#Find out how much fishmeal and oil is available globally
 
(total_fm_available_ge <- forage_catch_w_prop_fmfo_supply$fm_available_ge |> sum())
(total_fo_available_ge <- forage_catch_w_prop_fmfo_supply$fo_available_ge |> sum())

(total_fm_available_mass <- forage_catch_w_prop_fmfo_supply$fm_available_mass |> sum())
(total_fo_available_mass <- forage_catch_w_prop_fmfo_supply$fo_available_mass |> sum())


fmfo_demand

#use the proportions for each region to divide the demand up and work out how the spp_based demand converts to embodied fish

forage_supply_demand_by_diet <- 
  map_df(.x =  c("marine_diet", "plant_diet"), .f = \(this_diet){
    
    this_df <- forage_catch_w_prop_fmfo_supply |> 
      mutate(diet = this_diet,
             #extract total fmfo demand for each diet
             total_fm_needed_for_diet = fmfo_demand |> filter(diet == this_diet & ingredients == "fishmeal-wild") |> pull(total_ingredient_demand),
             total_fo_needed_for_diet = fmfo_demand |> filter(diet == this_diet & ingredients == "fish oil-wild") |> pull(total_ingredient_demand),
             #distribute fishmeal and oil demand according to its distribution
             applied_fm_demand_ge = total_fm_needed_for_diet * prop_available_fm_ge,
             applied_fo_demand_ge = total_fo_needed_for_diet * prop_available_fo_ge,
             applied_fm_demand_mass = total_fm_needed_for_diet * prop_available_fm_mass,
             applied_fo_demand_mass = total_fo_needed_for_diet * prop_available_fo_mass,
             #calculate embodied demand
             fm_embodied_fish_ge = applied_fm_demand_ge * fishmeal_cf_ge,
             fo_embodied_fish_ge = applied_fo_demand_ge * fish_oil_cf_ge,
             fm_embodied_fish_mass = applied_fm_demand_mass * fishmeal_cf_mass,
             fo_embodied_fish_mass = applied_fo_demand_mass * fish_oil_cf_mass,
             total_fmfo_demand_ge = applied_fm_demand_ge + applied_fo_demand_ge,
             total_fmfo_demand_mass = applied_fm_demand_mass + applied_fo_demand_mass,
             total_embodied_fish_ge = fm_embodied_fish_ge+fo_embodied_fish_ge,
             total_embodied_fish_mass = fm_embodied_fish_mass+fo_embodied_fish_mass
             ) |> 
      mutate(demand_vs_supply_mass = total_embodied_fish_mass/total_ind) |>
      mutate(demand_vs_supply_ge = total_embodied_fish_ge/total_ind) |>
      relocate(diet, .before = source_code)
    
  })



fwrite(forage_supply_demand_by_diet, file = here("data/tidy_data/demand/embodied_fish_per_cell.csv"))


#Create a total embodied fish demand by diet and source
embodied_fish_demand_by_diet_source <- 
  forage_supply_demand_by_diet |> 
  group_split(diet, source_code) |> 
  map_df(\(this_df){
    
    return(tibble(diet = unique(this_df$diet), 
                  source_code = unique(this_df$source_code), 
                  embodied_fish_demand_from_fm_ge = sum(this_df$fm_embodied_fish_ge),
                  embodied_fish_demand_from_fo_ge = sum(this_df$fo_embodied_fish_ge),
                  embodied_fish_demand_from_fm_mass = sum(this_df$fm_embodied_fish_mass),
                  embodied_fish_demand_from_fo_mass = sum(this_df$fo_embodied_fish_mass),
                  embodied_fish_demand_ge = sum(this_df$total_embodied_fish_ge),
                  embodied_fish_demand_mass = sum(this_df$total_embodied_fish_mass)))
    
  }) |> 
  pivot_longer(cols = -c(diet, source_code), names_to = "demand_from", values_to = "embodied_fish" ) |> 
  mutate(allocation = case_when(grepl("_ge", demand_from) ~ "Energetic allocation",
                                TRUE ~ "Mass allocation"
                                )) |> 
  mutate(demand_from = case_when(grepl("from_fm", demand_from) ~ "Fishmeal",
                                 grepl("from_fo", demand_from) ~ "Fish oil",
                                 TRUE ~ "Total")) |> arrange(diet, source_code, allocation)



saveRDS(object = embodied_fish_demand_by_diet_source, file = here("data/tidy_data/demand/embodied_fish_demand_by_source_diet.rds"))

source_production_potential
embodied_fish_demand_by_diet_source





```


Now do some checks on if demand and supply match up per record i.e. by per cell-species-gear combinations. 

```{r}
#check the demand vs supply match up - are there cells where supply is outstripped by demand? 
forage_supply_demand_by_diet |> filter(demand_vs_supply_mass>1)  #in no cell is demand higher than the supply




#clear the environment
rm(list = ls(all.names = TRUE))

```



