---
title: "11_crop_disturbance_pressures"
author: "Rich Cottrell"
date: "29/06/2022"
output: html_document
---

Here we use the MAPSPAM physical area parameter as our disturbance metric for crops following the assumptions of Halpern et al 2022 that cropland creates complete disturbance to natural systems where they exist.

```{r}
library(tidyverse)
library(terra)
library(tidyterra)
library(here)
library(countrycode)

source(here("src/directories.R"))

select <- dplyr::select

```

Import the crop demand data to get the FAO-MAPSPAM look up names. Also import the area harvested data to use as a proxy to understand how physical area may have changed based on the trend in area harvested. 

```{r}

mapspam_fao_lookup <- 
  readRDS(here("data/tidy_data/demand/total_crop_demand.rds")) |> 
  select(FAOSTAT_name, map_spam_code, source_country) |> 
  distinct() |> 
  mutate(map_spam_code = toupper(map_spam_code),
         iso3c = countrycode(sourcevar = source_country, origin = "country.name", destination = "iso3c", warn=TRUE))



crop_area_raw <- 
  readRDS(here("data/tidy_data/production-data/crops_production_tidy.rds"))

(crop_area <- crop_area_raw |>  
  filter(element == "Area harvested") |> 
  filter(item %in% unique(mapspam_fao_lookup$FAOSTAT_name)) |> 
  filter(year %in% c(2009, 2010, 2011, 2016, 2017, 2018)) |> 
  mutate(value_km2 = value*0.01) |> 
  select(area, iso3c, item, element, year, value_km2) |> 
  drop_na(value_km2) |> 
  mutate(era = if_else(condition = year %in% c(2009, 2010, 2011), true = "early", false = "late")) |> 
  group_by(area, iso3c, item, element, era) |> 
  summarise(mean_value_km2 = mean(value_km2, na.rm = TRUE)) |> 
  drop_na(iso3c) |> 
  pivot_wider(id_cols = c(area, iso3c, item, element), names_from = era, values_from = mean_value_km2) |> 
  mutate(rel_change = case_when(early == 0 ~ 1,
                              is.na(early) | is.na(late) ~ 1,
                              TRUE ~ late/early)) |> 
  filter(area != "China")
  )
  

max(crop_area$rel_change, na.rm = TRUE)

```


Check the area harvested from FAOSTAT with physical area from MAPSPAM to identify any discrepencies see if it plausible to fix by adjustment factors based on the greatest area harvested in either 2010 or 2017.

```{r}

crop_area_files <- list.files(here("data/spatial/crop-layers-reprojected"), pattern = "_A_", full.names = TRUE)

country_rast_files <- list.files(here("data/spatial/country-rasters"), full.names = TRUE)

countries_of_interest <- 
  country_rast_files |> map(\(this_element){
  this_file <- tools::file_path_sans_ext(this_element)
  this_country <- substr(this_file, start = nchar(this_file)-2, stop = nchar(this_file))
  return(this_country)
}) |> unlist()

crop_area_data_countries <- crop_area |> 
  filter(iso3c %in% countries_of_interest) 

#Some crazy changes in area harvested which doubtfully reflect the physical change in area - s
hist(crop_area_data_countries$rel_change) |> 
  abline(v= quantile(crop_area_data_countries$rel_change, c(0.75,0.9, 0.95, 0.99)), lty = 2)

quantile(crop_area_data_countries$rel_change, c(0.9, 0.95, 0.99))
summary(crop_area_data_countries$rel_change) # highly skewed so those values greater than the mean get the mean value (mean is greater than the 90th percentile)
area_adj_value <- mean(crop_area_data_countries$rel_change)

crop_area_data <- crop_area_data_countries |> 
  mutate(rel_change = case_when(rel_change > area_adj_value ~ area_adj_value,
                                TRUE ~ rel_change)) |> 
  group_by(item, iso3c, area) |> 
  group_split()

this_country_crop <- crop_area_data[[1]]

mapspam_rel_change <- 
  
  map_df(.x = crop_area_data, .f = \(this_country_crop){
  
  this_iso3 <- this_country_crop$iso3c
  this_item <- this_country_crop$item
  this_mapsam_code <- mapspam_fao_lookup |> filter(FAOSTAT_name == this_item) |> pull(map_spam_code) |> unique()
  
  message("Calculating area for ", "'", this_mapsam_code, "'", " in ", this_iso3)
  
  this_country_rast <- rast(country_rast_files[grep(this_iso3, country_rast_files)])
  vals <- terra::values(this_country_rast)
  
  (this_global_crop_rast <- rast(crop_area_files[grep(this_mapsam_code, crop_area_files)]))
  
  (this_country_crop_rast <- this_global_crop_rast*this_country_rast)
  
  mapspam_value <- sum(terra::values(this_country_crop_rast), na.rm=TRUE)
  
  return(this_country_crop |> mutate(mapspam_value = mapspam_value))

}) 
```


Adjust area rasters by the demand_production ratio and save in the int folder for each diet.

```{r}

crop_area_files <- list.files(here("data/spatial/crop-layers-reprojected"), pattern = "_A_", full.names = TRUE)

country_rast_files <- list.files(here("data/spatial/country-rasters"), full.names = TRUE)

demand_production_compare_w_sources <- readRDS(here("data/tidy_data/demand/sourcing_countries_crops.rds"))

scalar_source_list <- 
  demand_production_compare_w_sources |> 
  left_join( y = mapspam_rel_change, by = c("source_country" = "area", "source_iso3c" = "iso3c", "FAOSTAT_name" = "item")) |> 
  select(-c(ge_allocation_factor, mass_allocation_factor, total_crop_demand)) |> 
  rename(ge_allocation = demand_production_ratio_ge, mass_allocation = demand_production_ratio_ms) |> 
  filter(allocation_method == "ge_allocation") |> 
  select(-allocation_method) |> 
  pivot_longer(cols = c(ge_allocation, mass_allocation), names_to = "allocation_method", values_to = "demand_production_ratio") |> 
  group_split(ingredients, diet, FAOSTAT_name, source_iso3c, allocation_method) 
  
this_file <- scalar_source_list[[1]]


#rescale area by the demand-production ratio and save the rasters for each diet, ingredient, raw material
map(.x = scalar_source_list, \(this_file){
  
  this_ingredient <- this_file$ingredients 
  
  this_mapspam_code <- toupper(this_file$map_spam_code)
 
  this_diet <- this_file$diet
  
  this_country <- this_file$source_iso3c
  
  this_allocation_method <- this_file$allocation_method
  
  this_demand_scalar <- this_file$demand_production_ratio
  
  this_raster_scalar <- this_file$rel_change
  
  saveName <- sprintf(here("data/spatial/%s/int/%s_disturbance_km2_%s_%s.tif"), this_diet, this_ingredient,  this_allocation_method, this_country)
  
   message("Processing ", this_diet, ", ", basename(tools::file_path_sans_ext(saveName)))
  
  if(!file.exists(saveName)){
 
    #raw area rasters are already in km2 so no need to convert
    this_raw_rast <- rast(crop_area_files[grepl(this_mapspam_code, crop_area_files)]) 
  
    this_country_rast <- rast(country_rast_files[grepl(this_country, country_rast_files)])
    
    this_country_crop_rast <- this_raw_rast*this_country_rast
    
    this_country_crop_adj_rast <- this_country_crop_rast*this_raster_scalar #increases the area associated with production
    
    this_country_crop_demand_rast <- this_country_crop_adj_rast*this_demand_scalar
    
    this_corrected_area_raster <- lapp(c(this_country_crop_demand_rast,cellSize(this_country_crop_demand_rast)), \(x,y){
      
      if_else(condition = x<y, true = x, false = y) #scales the area produced by the demand:production ratio but limits to cellSize
    })
  
    
    writeRaster(x= this_corrected_area_raster, filename = saveName, overwrite=TRUE)

  }
})


```

