---
title: "Rescaling environmental pressures"
author: "Rich Cottrell"
date: "30/06/2022"
output: html_document
---

This file is used to rescale all rasters' values for each pressure so that values among feeds may be compared.

```{r}

library(tidyverse)
library(terra)
library(here)
library(parallel)
library(doParallel)
library(foreach)
library(future)
library(furrr)

source(here("src/functions.R"))


select <- dplyr::select
values <- terra::values



```

Create equal area projection for rescaling
```{r}

gall_peters <- "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
#raster template 
base_rast <- rast(res=0.5)
ext(base_rast) <- c(-180, 180, -90, 90)

eq_area_rast <- project(base_rast, gall_peters) 
values(eq_area_rast) <- 1:ncell(eq_area_rast)

cellSize(eq_area_rast, unit="km", transform = FALSE)

```



## Disturbance

Bring in  forage fish and crop layers for disturbance for both diets. Reproject onto the equal area gall-peters projection and calcute the sum of disturbance for each diet.

```{r}

#ID files for disturbance for marine and plant diets
marine_diet_disturbance <- list.files(here("data/spatial/marine_diet/int"), pattern = "disturbance_km2", full.names = TRUE)
marine_diet_disturbance <- marine_diet_disturbance[!grepl("forage_fish|_gp", marine_diet_disturbance) & grepl("allocation", marine_diet_disturbance)]

#REPROJECT PER KM2 DATA TO GALL PETERS EQUAL AREA

this_file <- marine_diet_disturbance[[25]]

#marine diets
md_reprojected_rasters <- 
  
  map(.x = marine_diet_disturbance, .f = \(this_file){
    
    base <- tools::file_path_sans_ext(basename(this_file))  
    
    this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
    
    this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
    
    this_ingredient <- str_extract(base, "[^_]+")
    
    saveName <- sprintf(here("data/spatial/marine_diet/int/%s_disturbance_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
    
    if(file.exists(saveName)){
      
      message("processing ", basename(saveName))
      
      this_rast <- rast(this_file)
      
      # normalise by cell size to get proportion of cell before reprojecting otherwise reprojection screws up to be larger than cell size.

      this_prop_r <- this_rast/cellSize(this_rast, unit = "km")

      this_reprojected_rast <-  project(this_prop_r, eq_area_rast)
      
      this_adj_reprojected_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km", transform = FALSE)
      names(this_adj_reprojected_rast) <- "km2"
      
      writeRaster(x = this_adj_reprojected_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reprojected_rast)
    }
    
  })




#plant diets
plant_diet_disturbance <- list.files(here("data/spatial/plant_diet/int"), pattern = "disturbance_km2", full.names = TRUE)
plant_diet_disturbance <- plant_diet_disturbance[!grepl("forage_fish|_gp", plant_diet_disturbance) & grepl("allocation", plant_diet_disturbance)]


this_file <- plant_diet_disturbance[[50]]

pd_reprojected_rasters <- 
  
  map(.x = plant_diet_disturbance, .f = \(this_file){
    
    base <- tools::file_path_sans_ext(basename(this_file))  
    
    this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
    
    this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
    
    this_ingredient <- str_extract(base, "[^_]+")
    
    saveName <- sprintf(here("data/spatial/plant_diet/int/%s_disturbance_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
    
   if(file.exists(saveName)){
      
      message("processing ", basename(saveName))
      
      this_rast <- rast(this_file)
      
      # normalise by cell size to get proportion of cell before reprojecting otherwise reprojection screws up to be larger than cell size.

      this_prop_r <- this_rast/cellSize(this_rast, unit = "km")

      this_reprojected_rast <-  project(this_prop_r, eq_area_rast)
      
      this_adj_reprojected_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km", transform = FALSE)
      names(this_adj_reprojected_rast) <- "km2"
      
      writeRaster(x = this_adj_reprojected_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reprojected_rast)
    }
    
  })


#NOW FIND THE MAXIMUM TOTAL SCORES FOR DISTURBANCE INCLUDING ALL INGREDIENTS FOR EACH DIET AND ALLOCATION METHOD - THIS WILL BE USED TO GET A GIVEN CELLS' PROPORTIONAL CONTRIBUTION

#First need to find maximum scores each ingredient (i.e. which origin yields the greatest)

md_reprojected_rasters <- list.files(path = here("data/spatial/marine_diet/int"), pattern = "disturbance_gp", full=TRUE)

md_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> filter(diet=="marine_diet" & ingredient_demand_tonnes>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()

this_df <- md_ingredients_list[[1]]

maximum_disturbance_scores_md <-  
  
  map_df(md_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
  these_ingredient_rasters_ge <- md_reprojected_rasters[grepl(this_ingredient, md_reprojected_rasters) & grepl("ge_allocation", md_reprojected_rasters)]
  
  these_ingredient_rasters_mass <- md_reprojected_rasters[grepl(this_ingredient, md_reprojected_rasters) & grepl("mass_allocation", md_reprojected_rasters)]
  
  max_disturbance_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  max_disturbance_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  this_adj_df <- bind_cols(this_df, max_disturbance_ge = max_disturbance_ge, max_disturbance_mass= max_disturbance_mass)
  
  
})

#get maximum disturbance scores for each allocation method
max_md_disturbance_ge <- maximum_disturbance_scores_md |> pull(max_disturbance_ge) |> sum()
max_md_disturbance_mass <- maximum_disturbance_scores_md |> pull(max_disturbance_mass) |> sum()



pd_reprojected_rasters <- list.files(path = here("data/spatial/plant_diet/int"), pattern = "disturbance_gp", full=TRUE)


pd_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> filter(diet=="plant_diet" & ingredient_demand_tonnes>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()

this_df <- md_ingredients_list[[1]]

maximum_disturbance_scores_pd <-  
  
  map_df(pd_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
  these_ingredient_rasters_ge <- pd_reprojected_rasters[grepl(this_ingredient, pd_reprojected_rasters) & grepl("ge_allocation", pd_reprojected_rasters)]
  
  these_ingredient_rasters_mass <- pd_reprojected_rasters[grepl(this_ingredient, pd_reprojected_rasters) & grepl("mass_allocation", pd_reprojected_rasters)]
  
  max_disturbance_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  max_disturbance_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  this_adj_df <- bind_cols(this_df, max_disturbance_ge = max_disturbance_ge, max_disturbance_mass= max_disturbance_mass)
  
  
})




#get maximum disturbance scores for each allocation method
max_pd_disturbance_ge <- maximum_disturbance_scores_pd |> pull(max_disturbance_ge) |> sum()
max_pd_disturbance_mass <- maximum_disturbance_scores_pd |> pull(max_disturbance_mass) |> sum()


max_total_disturbance_ge <- max(max_md_disturbance_ge,max_pd_disturbance_ge)
max_total_disturbance_mass <- max(max_md_disturbance_mass,max_pd_disturbance_mass)



# #which diet has the highest disturbance values
# marine_disturbance_vals_ge <- md_reprojected_rasters[grepl("ge_allocation", md_reprojected_rasters)] |> map(rast) |> map(no_na_values) |> unlist()
# plant_disturbance_vals_ge <-  pd_reprojected_rasters[grepl("ge_allocation", pd_reprojected_rasters)] |> map(rast) |> map(no_na_values) |> unlist()
# 
# marine_disturbance_vals_mass <- md_reprojected_rasters[grepl("mass_allocation", md_reprojected_rasters)] |> map(rast) |> map(no_na_values) |> unlist()
# plant_disturbance_vals_mass <-  pd_reprojected_rasters[grepl("mass_allocation", pd_reprojected_rasters)] |> map(rast) |> map(no_na_values) |> unlist()
# 
# 
# marine_diet_total_ge <- sum(marine_disturbance_vals_ge, na.rm = TRUE)
# marine_diet_total_mass <- sum(marine_disturbance_vals_mass, na.rm=TRUE)
# 
# plant_diet_total_ge <- sum(plant_disturbance_vals_ge, na.rm = TRUE)
# plant_diet_total_mass <- sum(plant_disturbance_vals_mass, na.rm=TRUE)
# 
# 
# 
# #check marine is appearing more intense (given forage fish disturbance values are much higher)
# summary(marine_disturbance_vals_ge)
# summary(plant_disturbance_vals_ge)
# 
# summary(marine_disturbance_vals_mass)
# summary(plant_disturbance_vals_mass)
# 
# all_disturbance_vals_ge <- c(marine_disturbance_vals_ge, plant_disturbance_vals_ge) |> unlist()
# all_disturbance_vals_mass <- c(marine_disturbance_vals_mass, plant_disturbance_vals_mass) |> unlist()
# 
# disturbance_energy_skewness <-  3*(mean(all_disturbance_vals_ge)-median(all_disturbance_vals_ge))/sd(all_disturbance_vals_ge)
# disturbance_mass_skewness <- 3*(mean(all_disturbance_vals_mass)-median(all_disturbance_vals_mass))/sd(all_disturbance_vals_mass)
# 
# #determine the quantile for rescaling - given the extreme skew - 99.9% seems 
# hist((all_disturbance_vals_ge), main = "", xlab = "Disturbance values") |> 
# abline(v = quantile(all_disturbance_vals_ge, 0.999), lty=2)
# 
# hist((all_disturbance_vals_mass), main = "", xlab = "Disturbance values") |> 
# abline(v = quantile(all_disturbance_vals_mass, 0.999), lty=2)
# 
# disturbance_rescale_val_ge <- quantile(all_disturbance_vals_ge, 0.999) |> as.numeric()
# disturbance_rescale_val_mass <- quantile(all_disturbance_vals_mass, 0.999) |> as.numeric()


#rescale all rasters by the proportion each cell represents of total possible disturbance for each diet globally (for both energy and mass allocation, this is marine_diet)

#marine diets

this_file <- md_reprojected_rasters[[11]]


rescaled_marine_disturbance_rasters <- 
  
  map(.x= md_reprojected_rasters, .f = \(this_file){
    
    this_rast <- rast(this_file)
    
    this_basename <- str_replace(basename(this_file), pattern = "_km2_gp", replacement = "")
    
    this_allocation_method <- str_extract(this_file, "ge_allocation|mass_allocation")
    
    saveName <- sprintf(here("data/spatial/marine_diet/pressures/%s"), this_basename)
    
    if(file.exists(saveName)){
      
      message("Processing ", "'", this_basename, "'") 
      
      if(this_allocation_method=="ge_allocation"){
        
        this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_disturbance_ge, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
        
        
      } else { 
        
         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_disturbance_mass, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
        
        
        
        }
    }
  })


this_file <- pd_reprojected_rasters[[50]]


#plant diets

rescaled_plant_disturbance_rasters <- 
  
  map(.x= pd_reprojected_rasters, .f = \(this_file){
  
  this_rast <- rast(this_file)
  
  this_basename <- str_replace(basename(this_file), pattern = "_km2_gp", replacement = "")
  
  this_allocation_method <- str_extract(this_file, "ge_allocation|mass_allocation")
  
  saveName <- sprintf(here("data/spatial/plant_diet/pressures/%s"), this_basename)
  
  if(file.exists(saveName)){
    
   message("Processing ", "'", this_basename, "'")
    
    if(this_allocation_method=="ge_allocation"){
        
        this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_disturbance_ge, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
        
        
      } else { 
        
         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_disturbance_mass, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
        
        }
  }
})



```

# Greenhouse gas pressures

```{r}

marine_diet_ghg <- list.files(here("data/spatial/marine_diet/int"), pattern = "ghg", full.names = TRUE)
marine_diet_ghg <- marine_diet_ghg[!grepl("forage_fish|_gp", marine_diet_ghg)]

plant_diet_ghg <- list.files(here("data/spatial/plant_diet/int"), pattern = "ghg", full.names = TRUE)
plant_diet_ghg <- plant_diet_ghg[!grepl("forage_fish|_gp", plant_diet_ghg)]

# marine_ghg_remove <- marine_diet_ghg[!grepl("allocation", marine_diet_ghg)]
# plant_ghg_remove <- plant_diet_ghg[!grepl("allocation", plant_diet_ghg)]

#REPROJECT PER KM2 DATA TO GALL PETERS EQUAL AREA FOR TOTAL PRESSURES

this_file <- marine_diet_ghg[[1]]


# Marine diets


map(.x = marine_diet_ghg, .f =  \(this_file){
    
    base <- tools::file_path_sans_ext(basename(this_file))  
    
    this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
    
    this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
    
    this_ingredient <- str_extract(base, "[^_]+")
    
    saveName <- sprintf(here("data/spatial/marine_diet/int/%s_ghg_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
    
    if(file.exists(saveName)){
      
      message("Processing ", basename(saveName))
      
      this_rast <- rast(this_file)
      
      this_reprojected_rast <- project(this_rast, eq_area_rast)
      
      this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit="km", transform = FALSE) #multiply up the footprints from per km2 to total based on 100km2
      names(this_adj_reproj_rast) <- "kgCO2eq"

      writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reproj_rast)
    }
})


#Plant diets

this_file <- plant_diet_ghg[[1]]

  
map(.x = plant_diet_ghg, .f = \(this_file){
    
    base <- tools::file_path_sans_ext(basename(this_file))  
    
    this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
    
    this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
    
    this_ingredient <- str_extract(base, "[^_]+")
    
    saveName <- sprintf(here("data/spatial/plant_diet/int/%s_ghg_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
    
    if(file.exists(saveName)){
      
      message("processing ", basename(saveName))
      
      this_rast <- rast(this_file)
      
      this_reprojected_rast <- project(this_rast, eq_area_rast)
      
      this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit="km", transform = FALSE) #multiply up the footprints from per km2 to total based on 100km2
      
      names(this_adj_reproj_rast) <- "kgCO2eq"
      
      writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reproj_rast)
    }
  })
  


#RESCALE GHG DATA BY PROP OF MAXIMUM GLOBAL DISTURBANCE ACROSS DIETS


#NOW FIND THE MAXIMUM TOTAL SCORES FOR DISTURBANCE INCLUDING ALL INGREDIENTS FOR EACH DIET AND ALLOCATION METHOD - THIS WILL BE USED TO GET A GIVEN CELLS' PROPORTIONAL CONTRIBUTION

# First need to find maximum scores each ingredient (i.e. which origin yields the greatest)
# Then need to rescale using the maximum plausible disturbance for each diet as the denominator



## Marine diets


md_reprojected_rasters_ghg <- list.files(path = here("data/spatial/marine_diet/int"), pattern = "ghg_gp", full=TRUE)

md_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> filter(diet=="marine_diet" & ingredient_demand_tonnes>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()

this_df <- md_ingredients_list[[1]]



maximum_ghg_scores_md <-  
  
  map_df(md_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
  these_ingredient_rasters_ge <- md_reprojected_rasters_ghg[grepl(this_ingredient, md_reprojected_rasters_ghg) & grepl("ge_allocation", md_reprojected_rasters_ghg)]
  
  these_ingredient_rasters_mass <- md_reprojected_rasters_ghg[grepl(this_ingredient, md_reprojected_rasters_ghg) & grepl("mass_allocation", md_reprojected_rasters_ghg)]
  
  max_ghg_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  max_ghg_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  this_adj_df <- bind_cols(this_df, max_ghg_ge = max_ghg_ge, max_ghg_mass= max_ghg_mass)
  
  
})

#get maximum marine diet disturbance scores for each allocation method
max_md_ghg_ge <- maximum_ghg_scores_md |> pull(max_ghg_ge) |> sum()
max_md_ghg_mass <- maximum_ghg_scores_md |> pull(max_ghg_mass) |> sum()




#Plant diets

pd_reprojected_rasters_ghg <- list.files(path = here("data/spatial/plant_diet/int"), pattern = "ghg_gp", full=TRUE)


pd_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> filter(diet=="plant_diet" & ingredient_demand_tonnes>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()

this_df <- md_ingredients_list[[1]]

maximum_ghg_scores_pd <-  
  
  map_df(pd_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
  these_ingredient_rasters_ge <- pd_reprojected_rasters_ghg[grepl(this_ingredient, pd_reprojected_rasters_ghg) & grepl("ge_allocation", pd_reprojected_rasters_ghg)]
  
  these_ingredient_rasters_mass <- pd_reprojected_rasters_ghg[grepl(this_ingredient, pd_reprojected_rasters_ghg) & grepl("mass_allocation", pd_reprojected_rasters_ghg)]
  
  max_ghg_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  max_ghg_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  this_adj_df <- bind_cols(this_df, max_ghg_ge = max_ghg_ge, max_ghg_mass= max_ghg_mass)
  
  
})


#compare teh scores by diet
maximum_ghg_scores_md
maximum_ghg_scores_pd

#get maximum ghg scores for each allocation method
max_pd_ghg_ge <- maximum_ghg_scores_pd |> pull(max_ghg_ge) |> sum()
max_pd_ghg_mass <- maximum_ghg_scores_pd |> pull(max_ghg_mass) |> sum()


max_total_ghg_ge <- max(max_md_ghg_ge,max_pd_ghg_ge)
max_total_ghg_mass <- max(max_md_ghg_mass,max_pd_ghg_mass)





#RESCALE THE TOTAL GHG DATA

#test file
this_file <- md_reprojected_rasters_ghg[[1]]

#Rescale all ghg rasters to the 99.9th percentile

#MARINE DIETS 

rescaled_marine_ghg_rasters <- 
  
  map(.x= md_reprojected_rasters_ghg, .f = \(this_file){
  
  this_rast <- rast(this_file)
  
  this_basename <- str_replace(basename(this_file), pattern = "_gp", replacement = "")
  
  this_allocation_method <- str_extract(this_file, "ge_allocation|mass_allocation")
  
  saveName <- sprintf(here("data/spatial/marine_diet/pressures/%s"), this_basename)
  
  if(file.exists(saveName)){
      
      message("Processing ", "'", this_basename, "'") 
      
      if(this_allocation_method=="ge_allocation"){
        
        this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_ghg_ge, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
        
        
      } else { 
        
         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_ghg_mass, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
        
        }
  }
  
})



#PLANT DIETS

#test file
this_file <- pd_reprojected_rasters_ghg[[1]]


rescaled_plant_ghg_rasters <- 
  
  map(.x= pd_reprojected_rasters_ghg, .f = \(this_file){
  
  this_rast <- rast(this_file)
  
  this_basename <- str_replace(basename(this_file), pattern = "_gp", replacement = "")
  
  this_allocation_method <- str_extract(this_file, "ge_allocation|mass_allocation")
  
  saveName <- sprintf(here("data/spatial/plant_diet/pressures/%s"), this_basename)
  
  if(file.exists(saveName)){
    
    message("Processing ", "'", this_basename, "'")
    
   if(this_allocation_method=="ge_allocation"){
        
        this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_ghg_ge, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
        
        
      } else { 
        
         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_ghg_mass, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
        
        }
    
  }
  
})



```

# Nutrient pressures

Do first for aggregated nutrients
```{r}
marine_diet_nutrients <- list.files(here("data/spatial/marine_diet/int"), pattern = "nutrient_km2", full.names = TRUE)

plant_diet_nutrients <- list.files(here("data/spatial/plant_diet/int"), pattern = "nutrient_km2", full.names = TRUE)


#REPROJECT PER KM2 DATA TO GALL PETERS EQUAL AREA FOR TOTAL PRESSURES

this_file <- marine_diet_nutrients[[1]]


# Marine diets
  
map(.x = marine_diet_nutrients, .f = \(this_file){
    
  base <- tools::file_path_sans_ext(basename(this_file))  
  
  this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
  
  this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
  
  this_ingredient <- str_extract(base, "[^_]+")
  
  saveName <- sprintf(here("data/spatial/marine_diet/int/%s_nutrient_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
  
    if(file.exists(saveName)){
      
      message("Processing ", basename(saveName))
      
      this_rast <- rast(this_file)
      
      this_reprojected_rast <- project(this_rast, gall_peters)
      
      this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km", transform=FALSE) #multiply up the footprints from per km2 to total based on cellsize
      names(this_adj_reproj_rast) <- "summed MT N/P eq."

      writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reproj_rast)
    }
  })


#Plant diets 

this_file <- plant_diet_nutrients[[1]]

map(.x = plant_diet_nutrients, .f = \(this_file){
    
    base <- tools::file_path_sans_ext(basename(this_file))  
    
    this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
    
    this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
    this_ingredient <- str_extract(base, "[^_]+")
    
    saveName <- sprintf(here("data/spatial/plant_diet/int/%s_nutrient_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
    
    if(file.exists(saveName)){
      
      message("processing ", basename(saveName))
      
      this_rast <- rast(this_file)
      
      this_reprojected_rast <- project(this_rast, gall_peters)
      
      this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km", transform=FALSE) #multiply up the footprints from per km2 to total based on cellsize
      
      names(this_adj_reproj_rast) <- "summed MT N/P eq."
      
      writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reproj_rast)
    }
  })




#RESCALE NUTRIENT DATA BY PROP OF MAXIMUM GLOBAL NUTRIENTS ACROSS DIETS


#NOW FIND THE MAXIMUM TOTAL SCORES FOR NUTRIENTS INCLUDING ALL INGREDIENTS FOR EACH DIET AND ALLOCATION METHOD - THIS WILL BE USED TO GET A GIVEN CELLS' PROPORTIONAL CONTRIBUTION

# First need to find maximum scores each ingredient (i.e. which origin yields the greatest)
# Then need to rescale using the maximum plausible nutrient for each diet as the denominator



## Marine diets

md_reprojected_rasters_nutrient <- list.files(path = here("data/spatial/marine_diet/int"), pattern = "nutrient_gp", full=TRUE)

md_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> filter(diet=="marine_diet" & ingredient_demand_tonnes>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()


#test function
this_df <- md_ingredients_list[[1]]


maximum_nutrient_scores_md <-  
  
  map_df(md_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
  these_ingredient_rasters_ge <- md_reprojected_rasters_nutrient[grepl(this_ingredient, md_reprojected_rasters_nutrient) & grepl("ge_allocation", md_reprojected_rasters_nutrient)]
  
  these_ingredient_rasters_mass <- md_reprojected_rasters_nutrient[grepl(this_ingredient, md_reprojected_rasters_nutrient) & grepl("mass_allocation", md_reprojected_rasters_nutrient)]
  
  max_nutrient_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  max_nutrient_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  this_adj_df <- bind_cols(this_df, max_nutrient_ge = max_nutrient_ge, max_nutrient_mass= max_nutrient_mass)
  
  
})

#get maximum marine diet nutrient scores for each allocation method
max_md_nutrient_ge <- maximum_nutrient_scores_md |> pull(max_nutrient_ge) |> sum()
max_md_nutrient_mass <- maximum_nutrient_scores_md |> pull(max_nutrient_mass) |> sum()





#Plant diets

pd_reprojected_rasters_nutrient <- list.files(path = here("data/spatial/plant_diet/int"), pattern = "nutrient_gp", full=TRUE)


pd_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> filter(diet=="plant_diet" & ingredient_demand_tonnes>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()

this_df <- pd_ingredients_list[[5]]

maximum_nutrient_scores_pd <-  
  
  map_df(pd_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
  these_ingredient_rasters_ge <- pd_reprojected_rasters_nutrient[grepl(this_ingredient, pd_reprojected_rasters_nutrient) & grepl("ge_allocation", pd_reprojected_rasters_nutrient)]
  
  these_ingredient_rasters_mass <- pd_reprojected_rasters_nutrient[grepl(this_ingredient, pd_reprojected_rasters_nutrient) & grepl("mass_allocation", pd_reprojected_rasters_nutrient)]
  
  max_nutrient_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  max_nutrient_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  this_adj_df <- bind_cols(this_df, max_nutrient_ge = max_nutrient_ge, max_nutrient_mass= max_nutrient_mass)
  
  
})

#compare nutrient scores by ingredient by diet
maximum_nutrient_scores_md
maximum_nutrient_scores_pd

#get maximum nutrient scores for each allocation method
max_pd_nutrient_ge <- maximum_nutrient_scores_pd |> pull(max_nutrient_ge) |> sum()
max_pd_nutrient_mass <- maximum_nutrient_scores_pd |> pull(max_nutrient_mass) |> sum()


#get maximum total values for nutrients across diets 
max_total_nutrient_ge <- max(max_md_nutrient_ge,max_pd_nutrient_ge)
max_total_nutrient_mass <- max(max_md_nutrient_mass,max_pd_nutrient_mass)



##rescale all rasters by the proportion each cell represents of total possible nitrogen pollution for each diet globally (for both energy and mass allocation, this is marine_diet)

#marine diets

this_file <- md_reprojected_rasters_nutrient[[1]]

map(.x= md_reprojected_rasters_nutrient, .f = \(this_file){

  this_rast <- rast(this_file)

  this_basename <- str_replace(basename(this_file), pattern = "_gp", replacement = "")

  this_allocation_method <- str_extract(string = this_basename, pattern = "ge_allocation|mass_allocation")

  saveName <- sprintf(here("data/spatial/marine_diet/pressures/%s"), this_basename)

  if(file.exists(saveName)){

     message("Processing", "'", this_basename, "'")

    if(this_allocation_method=="ge_allocation"){

        this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_nutrient_ge, false = 0)})

        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)

        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)

        return(this_rescaled_rast)


      } else {

         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_nutrient_mass, false = 0)})

        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)

        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)

        return(this_rescaled_rast)
    }
  }

})



#plant diets 

map(.x= pd_reprojected_rasters_nutrient, .f = \(this_file){

  this_rast <- rast(this_file)

  this_basename <- str_replace(basename(this_file), pattern = "_gp", replacement = "")

  this_allocation_method <- str_extract(string = this_basename, pattern = "ge_allocation|mass_allocation")

  saveName <- sprintf(here("data/spatial/plant_diet/pressures/%s"), this_basename)

  if(file.exists(saveName)){

    message("Processing", "'", this_basename, "'")

   if(this_allocation_method=="ge_allocation"){

        this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_nutrient_ge, false = 0)})

        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)

        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)

        return(this_rescaled_rast)


      } else {

         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_nutrient_mass, false = 0)})

        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)

        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)

        return(this_rescaled_rast)

    }

  }

})


```

These two sections are for individual nutrients

First for Nitrogen
```{r}
# marine_diet_N <- list.files(here("data/spatial/marine_diet/int"), pattern = "_N_", full.names = TRUE)
# marine_diet_N <- marine_diet_N[!grepl("forage_fish|_gp", marine_diet_N)]
# 
# plant_diet_N <- list.files(here("data/spatial/plant_diet/int"), pattern = "_N_", full.names = TRUE)
# plant_diet_N <- plant_diet_N[!grepl("forage_fish|_gp", plant_diet_N)]
# 
# 
# 
# #REPROJECT PER KM2 DATA TO GALL PETERS EQUAL AREA FOR TOTAL PRESSURES
# 
# this_file <- marine_diet_N[[1]]
# 
# 
# # Marine diets
#   
# map(.x = marine_diet_N, .f = \(this_file){
#     
#     base <- tools::file_path_sans_ext(basename(this_file))  
#     
#      this_country <- if_else(grepl("EPC|NAT|OTH", base), true = str_extract(base, "EPC|NAT|OTH"), false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
#     
#     this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
#     
#     this_ingredient <- str_extract(base, "[^_]+")
#     
#     saveName <- sprintf(here("data/spatial/marine_diet/int/%s_N_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
#     
#     if(file.exists(saveName)){
#       
#       message("Processing ", basename(saveName))
#       
#       this_rast <- rast(this_file)
#       
#       this_reprojected_rast <- project(this_rast, gall_peters)
#       
#       this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km") #multiply up the footprints from per km2 to total based on cellsize
#       names(this_adj_reproj_rast) <- "N_eq"
# 
#       writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
#       
#       return(this_adj_reproj_rast)
#     }
#   })
# 
# 
# 
# #Plant diets
# 
# this_file <- plant_diet_N[[1]]
# 
#   
# map(.x = plant_diet_N, .f = \(this_file){
#     
#     base <- tools::file_path_sans_ext(basename(this_file))  
#     
#    this_country <- if_else(grepl("EPC|NAT|OTH", base), true = str_extract(base, "EPC|NAT|OTH"), false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
#     
#     this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
#     this_ingredient <- str_extract(base, "[^_]+")
#     
#     saveName <- sprintf(here("data/spatial/plant_diet/int/%s_N_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
#     
#     if(file.exists(saveName)){
#       
#       message("processing ", basename(saveName))
#       
#       this_rast <- rast(this_file)
#       
#       this_reprojected_rast <- project(this_rast, gall_peters)
#       
#       this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km") #multiply up the footprints from per km2 to total based on cellsize
#       
#       names(this_adj_reproj_rast) <- "N_eq"
#       
#       writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
#       
#       return(this_adj_reproj_rast)
#     }
#   })
# 
# 
# #RESCALE NITROGEN DATA BY PROP OF MAXIMUM GLOBAL NITROGEN ACROSS DIETS
# 
# 
# #NOW FIND THE MAXIMUM TOTAL SCORES FOR NITOGREN INCLUDING ALL INGREDIENTS FOR EACH DIET AND ALLOCATION METHOD - THIS WILL BE USED TO GET A GIVEN CELLS' PROPORTIONAL CONTRIBUTION
# 
# # First need to find maximum scores each ingredient (i.e. which origin yields the greatest)
# # Then need to rescale using the maximum plausible nitrogen for each diet as the denominator
# 
# 
# 
# ## Marine diets
# 
# 
# md_reprojected_rasters_N <- list.files(path = here("data/spatial/marine_diet/int"), pattern = "N_gp", full=TRUE)
# 
# md_ingredients_list <- 
#   readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet=="marine_diet" & total_ingredient_demand>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()
# 
# 
# #test function
# this_df <- md_ingredients_list[[1]]
# 
# 
# 
# maximum_N_scores_md <-  
#   
#   map_df(md_ingredients_list, .f = \(this_df){
#   
#   this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
#   
#   these_ingredient_rasters_ge <- md_reprojected_rasters_N[grepl(this_ingredient, md_reprojected_rasters_N) & grepl("ge_allocation", md_reprojected_rasters_N)]
#   
#   these_ingredient_rasters_mass <- md_reprojected_rasters_N[grepl(this_ingredient, md_reprojected_rasters_N) & grepl("mass_allocation", md_reprojected_rasters_N)]
#   
#   max_N_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   max_N_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   
#   this_adj_df <- bind_cols(this_df, max_N_ge = max_N_ge, max_N_mass= max_N_mass)
#   
#   
# })
# 
# #get maximum marine diet disturbance scores for each allocation method
# max_md_N_ge <- maximum_N_scores_md |> pull(max_N_ge) |> sum()
# max_md_N_mass <- maximum_N_scores_md |> pull(max_N_mass) |> sum()
# 
# 
# 
# 
# #Plant diets
# 
# pd_reprojected_rasters_N <- list.files(path = here("data/spatial/plant_diet/int"), pattern = "N_gp", full=TRUE)
# 
# 
# pd_ingredients_list <- 
#   readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet=="plant_diet" & total_ingredient_demand>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()
# 
# this_df <- pd_ingredients_list[[1]]
# 
# maximum_N_scores_pd <-  
#   
#   map_df(pd_ingredients_list, .f = \(this_df){
#   
#   this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
#   
#   these_ingredient_rasters_ge <- pd_reprojected_rasters_N[grepl(this_ingredient, pd_reprojected_rasters_N) & grepl("ge_allocation", pd_reprojected_rasters_N)]
#   
#   these_ingredient_rasters_mass <- pd_reprojected_rasters_N[grepl(this_ingredient, pd_reprojected_rasters_N) & grepl("mass_allocation", pd_reprojected_rasters_N)]
#   
#   max_N_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   max_N_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   
#   this_adj_df <- bind_cols(this_df, max_N_ge = max_N_ge, max_N_mass= max_N_mass)
#   
#   
# })
# 
# #get maximum N scores for each allocation method
# max_pd_N_ge <- maximum_N_scores_pd |> pull(max_N_ge) |> sum()
# max_pd_N_mass <- maximum_N_scores_pd |> pull(max_N_mass) |> sum()
# 
# 
# max_total_N_ge <- max(max_md_N_ge,max_pd_N_ge)
# max_total_N_mass <- max(max_md_N_mass,max_pd_N_mass)
# 
# 
# 
# 
# ##rescale all rasters by the proportion each cell represents of total possible nitrogen pollution for each diet globally (for both energy and mass allocation, this is marine_diet)
# 
# #marine diets
# 
# this_file <- md_reprojected_rasters_N[[1]]
# 
# map(.x= md_reprojected_rasters_N, .f = \(this_file){
#   
#   this_rast <- rast(this_file)
#   
#   this_basename <- str_replace(basename(this_file), pattern = "_gp", replacement = "")
#   
#   this_allocation_method <- str_extract(string = this_basename, pattern = "ge_allocation|mass_allocation")
#   
#   saveName <- sprintf(here("data/spatial/marine_diet/pressures/%s"), this_basename)
#   
#   if(file.exists(saveName)){
#     
#      message("Processing", "'", this_basename, "'")
#     
#     if(this_allocation_method=="ge_allocation"){
#         
#         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_N_ge, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#         
#         
#       } else { 
#         
#          this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_N_mass, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#     }
#   }
#   
# })
# 
# 
# 
# #plant diets _ Nitrogen
# 
# map(.x= pd_reprojected_rasters_N, .f = \(this_file){
#   
#   this_rast <- rast(this_file)
#   
#   this_basename <- str_replace(basename(this_file), pattern = "_gp", replacement = "")
#   
#   this_allocation_method <- str_extract(string = this_basename, pattern = "ge_allocation|mass_allocation")
#   
#   saveName <- sprintf(here("data/spatial/plant_diet/pressures/%s"), this_basename)
#   
#   if(file.exists(saveName)){
#     
#     message("Processing", "'", this_basename, "'")
#     
#    if(this_allocation_method=="ge_allocation"){
#         
#         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_N_ge, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#         
#         
#       } else { 
#         
#          this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_N_mass, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#       
#     }
#     
#   }
#   
# })
```


Repeat for phosphorus

```{r}
# Phosphorus
# marine_diet_P <- list.files(here("data/spatial/marine_diet/int"), pattern = "_P_", full.names = TRUE)
# marine_diet_P <- marine_diet_P[!grepl("forage_fish|_gp", marine_diet_P)]
# 
# plant_diet_P <- list.files(here("data/spatial/plant_diet/int"), pattern = "_P_", full.names = TRUE)
# plant_diet_P <- plant_diet_P[!grepl("forage_fish|_gp", plant_diet_P)]
# 
# 
# 
# #REPROJECT PER KM2 DATA TO GALL PETERS EQUAL AREA FOR TOTAL PRESSURES
# 
# this_file <- marine_diet_P[[1]]
# 
# 
# # Marine diets
#   
# map(.x = marine_diet_P, .f = \(this_file){
#     
#     base <- tools::file_path_sans_ext(basename(this_file))  
#     
#     this_country <- if_else(grepl("EPC|NAT|OTH", base), true = str_extract(base, "EPC|NAT|OTH"), false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
#     
#     this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
#     
#     this_ingredient <- str_extract(base, "[^_]+")
#     
#     saveName <- sprintf(here("data/spatial/marine_diet/int/%s_P_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
#     
#     if(file.exists(saveName)){
#       
#       message("Processing ", basename(saveName))
#       
#       this_rast <- rast(this_file)
#       
#       this_reprojected_rast <- project(this_rast, gall_peters)
#       
#       this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km") #multiply up the footprints from per km2 to total based on cellsize
#       names(this_adj_reproj_rast) <- "P_eq"
# 
#       writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
#       
#       return(this_adj_reproj_rast)
#     }
#   })
# 
# 
# 
# #Plant diets
# 
# this_file <- plant_diet_P[[1]]
# 
#   
# map(.x = plant_diet_P, .f = \(this_file){
#     
#     base <- tools::file_path_sans_ext(basename(this_file))  
#     
#     this_country <- if_else(grepl("EPC|NAT|OTH", base), true = str_extract(base, "EPC|NAT|OTH"), false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
#     
#     this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
#     
#     this_ingredient <- str_extract(base, "[^_]+")
#     
#     saveName <- sprintf(here("data/spatial/plant_diet/int/%s_P_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
#     
#     if(file.exists(saveName)){
#       
#       message("processing ", basename(saveName))
#       
#       this_rast <- rast(this_file)
#       
#       this_reprojected_rast <- project(this_rast, gall_peters)
#       
#       this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km") #multiply up the footprints from per km2 to total based on cellsize
#       
#       names(this_adj_reproj_rast) <- "P_eq"
#       
#       writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
#       
#       return(this_adj_reproj_rast)
#     }
#   })
# 
# 
# 
# 
# #RESCALE PHOSPHORUS DATA BY PROP OF MAXIMUM GLOBAL PHOSPHOROUS ACROSS DIETS
# 
# 
# #NOW FIND THE MAXIMUM TOTAL SCORES FOR PHOSPHOROUS INCLUDING ALL INGREDIENTS FOR EACH DIET AND ALLOCATION METHOD - THIS WILL BE USED TO GET A GIVEN CELLS' PROPORTIONAL CONTRIBUTION
# 
# # First need to find maximum scores each ingredient (i.e. which origin yields the greatest)
# # Then need to rescale using the maximum plausible disturbance for each diet as the denominator
# 
# 
# 
# ## Marine diets
# 
# 
# md_reprojected_rasters_P <- list.files(path = here("data/spatial/marine_diet/int"), pattern = "P_gp", full=TRUE)
# 
# md_ingredients_list <- 
#   readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet=="marine_diet" & total_ingredient_demand>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()
# 
# this_df <- md_ingredients_list[[1]]
# 
# 
# 
# maximum_P_scores_md <-  
#   
#   map_df(md_ingredients_list, .f = \(this_df){
#   
#   this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
#   
#   these_ingredient_rasters_ge <- md_reprojected_rasters_P[grepl(this_ingredient, md_reprojected_rasters_P) & grepl("ge_allocation", md_reprojected_rasters_P)]
#   
#   these_ingredient_rasters_mass <- md_reprojected_rasters_P[grepl(this_ingredient, md_reprojected_rasters_P) & grepl("mass_allocation", md_reprojected_rasters_P)]
#   
#   max_P_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   max_P_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   
#   this_adj_df <- bind_cols(this_df, max_P_ge = max_P_ge, max_P_mass= max_P_mass)
#   
#   
# })
# 
# #get maximum marine diet disturbance scores for each allocation method
# max_md_P_ge <- maximum_P_scores_md |> pull(max_P_ge) |> sum()
# max_md_P_mass <- maximum_P_scores_md |> pull(max_P_mass) |> sum()
# 
# 
# 
# 
# #Plant diets
# 
# pd_reprojected_rasters_P <- list.files(path = here("data/spatial/plant_diet/int"), pattern = "P_gp", full=TRUE)
# 
# 
# pd_ingredients_list <- 
#   readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet=="plant_diet" & total_ingredient_demand>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()
# 
# this_df <- pd_ingredients_list[[1]]
# 
# maximum_P_scores_pd <-  
#   
#   map_df(pd_ingredients_list, .f = \(this_df){
#   
#   this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
#   
#   these_ingredient_rasters_ge <- pd_reprojected_rasters_P[grepl(this_ingredient, pd_reprojected_rasters_P) & grepl("ge_allocation", pd_reprojected_rasters_P)]
#   
#   these_ingredient_rasters_mass <- pd_reprojected_rasters_P[grepl(this_ingredient, pd_reprojected_rasters_P) & grepl("mass_allocation", pd_reprojected_rasters_P)]
#   
#   max_P_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   max_P_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   
#   this_adj_df <- bind_cols(this_df, max_P_ge = max_P_ge, max_P_mass= max_P_mass)
#   
#   
# })
# 
# #get maximum N scores for each allocation method
# max_pd_P_ge <- maximum_P_scores_pd |> pull(max_P_ge) |> sum()
# max_pd_P_mass <- maximum_P_scores_pd |> pull(max_P_mass) |> sum()
# 
# 
# max_total_P_ge <- max(max_md_P_ge,max_pd_P_ge)
# max_total_P_mass <- max(max_md_P_mass,max_pd_P_mass)
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# # Now rescale pressure by the total plausible phosphorus pollution value 
# 
# #marine diets - Phosphorus
# 
# #this_file <- marine_diet_P[[1]]
# 
# map(.x= md_reprojected_rasters_P, .f = \(this_file){
#   
#   this_rast <- rast(this_file)
#   
#   this_basename <- str_replace(basename(this_file), pattern = "_gp", replacement = "")
#   
#   this_allocation_method <- str_extract(this_basename, "ge_allocation|mass_allocation")
#   
#   saveName <- sprintf(here("data/spatial/marine_diet/pressures/%s"), this_basename)
#   
#   if(file.exists(saveName)){
#     
#      message("Processing", "'", this_basename, "'")
#     
#     if(this_allocation_method=="ge_allocation"){
#         
#         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_P_ge, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#         
#         
#       } else { 
#         
#          this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_P_mass, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#       
#     }
#     
#     
#    
#   }
#   
# })
# 
# 
# 
# #plant diets - Phosphorus
# 
# map(.x= pd_reprojected_rasters_P, .f = \(this_file){
#   
#   this_rast <- rast(this_file)
#   
#   this_basename <- str_replace(basename(this_file), pattern = "_gp", replacement = "")
#   
#   this_allocation_method <- str_extract(this_basename, "ge_allocation|mass_allocation")
#   
#   saveName <- sprintf(here("data/spatial/plant_diet/pressures/%s"), this_basename)
#   
#   if(file.exists(saveName)){
#     
#      message("Processing", "'", this_basename, "'")
#     
#     if(this_allocation_method=="ge_allocation"){
#         
#         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_P_ge, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#         
#         
#       } else { 
#         
#          this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_P_mass, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#       
#     }
#     
#   }
#   
# })
# 




```


# Water extraction
```{r}

marine_diet_h2o <- list.files(here("data/spatial/marine_diet/int"), pattern = "_water_km2", full.names = TRUE)

plant_diet_h2o <- list.files(here("data/spatial/plant_diet/int"), pattern = "_water_km2", full.names = TRUE)



#REPROJECT PER KM2 DATA TO GALL PETERS EQUAL AREA FOR TOTAL PRESSURES

this_file <- marine_diet_h2o[[1]]


# Marine diets
  
map(.x = marine_diet_h2o, .f = \(this_file){
    
    base <- tools::file_path_sans_ext(basename(this_file))  
    
     this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
    
    this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
    
    this_ingredient <- str_extract(base, "[^_]+")
    
    saveName <- sprintf(here("data/spatial/marine_diet/int/%s_water_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
    
    if(file.exists(saveName)){
      
      message("Processing ", basename(saveName))
      
      this_rast <- rast(this_file)
      
      this_reprojected_rast <- project(this_rast, gall_peters)
      
      this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km", transform=FALSE) #multiply up the footprints from per km2 to total based on cellsize
      names(this_adj_reproj_rast) <- "m3"

      writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reproj_rast)
    }
  })



#Plant diets

this_file <- plant_diet_h2o[[1]]

  
map(.x = plant_diet_h2o, .f = \(this_file){
    
    base <- tools::file_path_sans_ext(basename(this_file))  
    
     this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
    
    this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
    
    this_ingredient <- str_extract(base, "[^_]+")
    
    saveName <- sprintf(here("data/spatial/plant_diet/int/%s_water_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
    
    if(file.exists(saveName)){
      
      message("processing ", basename(saveName))
      
      this_rast <- rast(this_file)
      
      this_reprojected_rast <- project(this_rast, gall_peters)
      
      this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km", transform=FALSE) #multiply up the footprints from per km2 to total based on cellsize
      
      names(this_adj_reproj_rast) <- "m3"
      
      writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reproj_rast)
    }
  })




#RESCALE WATER DATA BY PROP OF MAXIMUM GLOBAL WATER USE ACROSS DIETS


#NOW FIND THE MAXIMUM TOTAL SCORES FOR WATER INCLUDING ALL INGREDIENTS FOR EACH DIET AND ALLOCATION METHOD - THIS WILL BE USED TO GET A GIVEN CELLS' PROPORTIONAL CONTRIBUTION

# First need to find maximum scores each ingredient (i.e. which origin yields the greatest)
# Then need to rescale using the maximum plausible water consumption for each diet as the denominator



## Marine diets


md_reprojected_rasters_h2o <- list.files(path = here("data/spatial/marine_diet/int"), pattern = "water_gp", full=TRUE)

md_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> filter(diet=="marine_diet" & ingredient_demand_tonnes>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()

this_df <- md_ingredients_list[[3]]


maximum_h2o_scores_md <-  
  
  map_df(md_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
  these_ingredient_rasters_ge <- md_reprojected_rasters_h2o[grepl(this_ingredient, md_reprojected_rasters_h2o) & grepl("ge_allocation", md_reprojected_rasters_h2o)]
  
  these_ingredient_rasters_mass <- md_reprojected_rasters_h2o[grepl(this_ingredient, md_reprojected_rasters_h2o) & grepl("mass_allocation", md_reprojected_rasters_h2o)]
  
  max_h2o_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  max_h2o_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  this_adj_df <- bind_cols(this_df, max_h2o_ge = max_h2o_ge, max_h2o_mass= max_h2o_mass)
  
  
})

#get maximum marine diet disturbance scores for each allocation method
max_md_h2o_ge <- maximum_h2o_scores_md |> pull(max_h2o_ge) |> sum()
max_md_h2o_mass <- maximum_h2o_scores_md |> pull(max_h2o_mass) |> sum()




#Plant diets

pd_reprojected_rasters_h2o <- list.files(path = here("data/spatial/plant_diet/int"), pattern = "water_gp", full=TRUE)



pd_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> filter(diet=="plant_diet" & ingredient_demand_tonnes>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()

this_df <- pd_ingredients_list[[5]]

maximum_h2o_scores_pd <-  
  
  map_df(pd_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
  these_ingredient_rasters_ge <- pd_reprojected_rasters_h2o[grepl(this_ingredient, pd_reprojected_rasters_h2o) & grepl("ge_allocation", pd_reprojected_rasters_h2o)]
  
  these_ingredient_rasters_mass <- pd_reprojected_rasters_h2o[grepl(this_ingredient, pd_reprojected_rasters_h2o) & grepl("mass_allocation", pd_reprojected_rasters_h2o)]
  
  max_h2o_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  max_h2o_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  this_adj_df <- bind_cols(this_df, max_h2o_ge = max_h2o_ge, max_h2o_mass= max_h2o_mass)
  
  
})

#get maximum N scores for each allocation method
max_pd_h2o_ge <- maximum_h2o_scores_pd |> pull(max_h2o_ge) |> sum()
max_pd_h2o_mass <- maximum_h2o_scores_pd |> pull(max_h2o_mass) |> sum()


max_total_h2o_ge <- max(max_md_h2o_ge,max_pd_h2o_ge)
max_total_h2o_mass <- max(max_md_h2o_mass,max_pd_h2o_mass)

#Compare water consumption by diet
maximum_h2o_scores_md
maximum_h2o_scores_pd



#Rescale each diet by the max disturbance denomintor for each diet and allocation method

#marine diets - Water

map(.x= md_reprojected_rasters_h2o, .f = \(this_file){
  
  this_rast <- rast(this_file)
  
  this_basename <- str_replace(basename(this_file), pattern = "_gp", replacement = "")
  
  this_allocation_method <- str_extract(this_basename, "ge_allocation|mass_allocation")
  
  saveName <- sprintf(here("data/spatial/marine_diet/pressures/%s"), this_basename)
  
  if(file.exists(saveName)){
    
     message("Processing", "'", this_basename, "'")
    
     if(this_allocation_method=="ge_allocation"){
        
        this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_h2o_ge, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
        
        
      } else { 
        
         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_h2o_mass, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
    }
  }
  
})



#plant diets - water

map(.x= pd_reprojected_rasters_h2o, .f = \(this_file){
  
  this_rast <- rast(this_file)
  
  this_basename <- str_replace(basename(this_file), pattern = "_gp", replacement = "")
  
  this_allocation_method <- str_extract(this_basename, "ge_allocation|mass_allocation")
  
  saveName <- sprintf(here("data/spatial/plant_diet/pressures/%s"), this_basename)
  
  if(file.exists(saveName)){
    
     message("Processing", "'", this_basename, "'")
    
      if(this_allocation_method=="ge_allocation"){
        
        this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_h2o_ge, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
        
        
      } else { 
        
         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_h2o_mass, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
    } 
  }
  
})


#clear the environment
rm(list = ls(all.names = TRUE))
```







