---
title: "16 Conceptual plot of diets and approach"
author: "Rich Cottrell"
date: "14/02/2023"
output: html_document
---

Set up
```{r}
library(tidyverse)
library(ggpubr)
library(patchwork)
library(here)
library(sf)
library(terra)
library(rnaturalearth)
#library(rnaturalearthdata)
library(RColorBrewer)
library(gameofthrones)
library(parallel)
library(ggdist)
library(data.table)
library(dtplyr)
library(Cairo)


# taking cumulative impoact palette from Halpern et al 
red <- "#B90000"
d_brown <- "#515256"
m_brown <- "#B27B54"
l_brown <- "#BC995F"
green <- "#A8B072"
yellow<- "#EFCE71"
l_yellow <- "#F7F6C1"
light_gray <- "#F8F9FA"
discrete_pal <- c(l_yellow, yellow, green, l_brown, m_brown, d_brown, "firebrick")
continuous_pal <-  colorRampPalette(discrete_pal, space="Lab", bias = 3.5)(10000)
final_palette <- c(light_gray, continuous_pal, red)

#source files
source(here("src/fxns.R"))

#spatial products
moll_crs <- "ESRI:53009"
gp_crs <- "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"


countries <-    ne_countries(scale = "medium", returnclass = "sf") |> st_transform(crs = moll_crs)
bbox <- ne_download(scale = 50, type = "wgs84_bounding_box", category = "physical", returnclass = "sf") |> st_transform(crs = moll_crs)
coastline <- ne_coastline(scale = 50, returnclass = "sf") |> st_transform(crs = moll_crs)
select <- dplyr::select
values <- terra::values
```


Explanatory plot of the two diets

```{r}
marine_diet <- readRDS(here("data/tidy_data/diet-scenarios/marine_diet.rds")) |> 
  filter(prop >0) |> 
  add_row(groups = "Microingredients", ingredients = "microingredients", diet = "marine_diet", prop = 0.02) |> 
  mutate(ingredients = case_when(ingredients == "canola/camelina oil" ~ "Canola oil",
                                  TRUE ~ str_to_sentence(ingredients)),
         ingredients = gsub(pattern = "-wild", replacement = "", ingredients)) |> 
  mutate(ingredients = factor(ingredients, levels = c("Fishmeal", "Fish oil", "Faba beans", "Soybean meal", "Corn gluten meal", "Wheat gluten", "Wheat", "Microingredients"))) |> 
  mutate(cum_prop = cumsum(prop)) |> 
  mutate(label_y_pos = if_else(condition = prop != prop[1], true = prop/2+lag(cum_prop), false = cum_prop/2)) |> 
  mutate(label_y_pos = 1-label_y_pos) 


plant_diet <- readRDS(here("data/tidy_data/diet-scenarios/plant_diet.rds")) |> 
  filter(prop >0) |> 
  add_row(groups = "Microingredients", ingredients = "microingredients", diet = "plant_diet", prop = 0.04) |> 
  mutate(ingredients = case_when(ingredients == "canola/camelina oil" ~ "Canola oil",
                                  TRUE ~ str_to_sentence(ingredients)),
         ingredients = gsub(pattern = "-wild", replacement = "", ingredients)) |> 
  mutate(ingredients = factor(ingredients, levels = c("Fishmeal", "Fish oil", "Soy protein concentrate", "Corn gluten meal","Faba beans", "Pea protein concentrate",  "Sunflower meal", "Wheat gluten", "Canola oil", "Linseed oil", "Wheat", "Pea starch", "Corn starch", "Microingredients"))) |> 
  mutate(cum_prop = cumsum(prop)) |> 
  mutate(label_y_pos = if_else(condition = prop != prop[1], true = prop/2+lag(cum_prop), false = cum_prop/2))|> 
   mutate(label_y_pos = 1-label_y_pos) 

#check totals 
marine_diet |> pull(prop) |> sum() #adds to 1
plant_diet |> pull(prop) |> sum() #adds to 1

ingredient_levels <- c( "Fishmeal",  "Fish oil" , "Soybean meal","Corn gluten meal" , "Faba beans" , "Wheat gluten" ,"Wheat" , "Soy protein concentrate",   "Pea protein concentrate","Sunflower meal","Canola oil","Linseed oil","Pea starch","Corn starch", "Microingredients")

#get a plot and legend for each diet
marine_diet_plot <- ggplot(data = marine_diet,
       aes(x=diet, y = prop, fill = ingredients))+
  geom_col()+
  scale_fill_manual(values =  rev(colorRampPalette(colors = rcartocolor::carto_pal(n = 9, name = "Earth"))(16))[c(1:7,16)])
marine_legend <- get_legend(marine_diet_plot)
plant_diet_plot <- ggplot(data = plant_diet,
       aes(x=diet, y = prop, fill = ingredients))+
  geom_col()+
  scale_fill_manual(values =  rev(colorRampPalette(colors = rcartocolor::carto_pal(n = 9, name = "Earth"))(16))[c(1,2,5:16)])
plant_legend <- get_legend(plant_diet_plot)


#combine both diets
both_diets <- bind_rows(marine_diet, plant_diet) |> 
  
  mutate(label_x_pos = if_else(diet =="marine_diet", true =0.5, false = 2.5)) |> 
  mutate(diet = if_else(condition = diet == "marine_diet", true = "Marine-\ndominant", false = "Plant-\ndominant")) |> 
  mutate(hjust = if_else(condition = diet == "marine_diet", true = 1, false = 0)) |> 
  mutate(ingredients = factor(ingredients, levels = rev(ingredient_levels))) |> 
  mutate(diet = factor(diet, levels = rev(c("Plant-\ndominant", "Marine-\ndominant"))))

both_diets_p <- ggplot(data = both_diets,
       aes(x = diet , y = prop, fill = ingredients))+
  geom_col()+
  scale_fill_manual(values =  (colorRampPalette(colors = rcartocolor::carto_pal(n = 7, name = "Earth"))(16)))+
  theme_pubr()+
  theme(legend.title = element_blank(),
        legend.text = element_text(size=7),
        text = element_text(size=8),
        legend.position = "right",
        #legend.box.spacing = unit(-0.1, "cm"),
        panel.grid = element_blank(),
        legend.key.size = unit(0.25, "cm")
        )+
  labs(x = "Feed scenario", y = "Proportion")+
  guides(fill = guide_legend(byrow=TRUE, reverse = FALSE, ncol = 1))

ggsave(filename = here("figures/feed_composition.jpg"), dpi = 600, width = 10, height = 9, units = "cm")
```


Let's simplify that plot to just have "marine ingredients" (blue) or "plant ingredients" (green), and make it horizontal. 

```{r}

simplified_df <- both_diets %>%
  mutate(ingredients_2 = ifelse(ingredients %in% c("Fishmeal", "Fish oil"), "Marine ingredients", "Plant ingredients"))


pl <- ggplot(data = simplified_df,
       aes(x = diet , y = prop, fill = ingredients_2))+
  geom_col()+
  scale_fill_manual(values =  c("#93AAD8", "#7EAA55"))+
  theme_pubr()+
  theme(legend.title = element_blank(),
        legend.text = element_text(size=12),
        text = element_text(size=12),
        legend.position = "bottom",
        #legend.box.spacing = unit(-0.1, "cm"),
        panel.grid = element_blank(),
        legend.key.size = unit(0.25, "cm")
        ) +
  labs(x = "Feed scenario", y = "Proportion")+
  guides(fill = guide_legend(byrow=TRUE, reverse = FALSE, ncol = 1)) + 
  coord_flip()


CairoPNG(here("figures/feed_composition_3.jpg"))
pl
dev.off()

# ggsave(here("figures/feed_composition_3.jpg"), width = 7.5, height = 5, units = "in")


```

Ok, now lets make a plot like this: Feed profile; Ingredients along the top, two rows for fish-dominant or plant-dominant feed scenarios, alpha is the proportional contribution of each ingredient

```{r}

ggplot(data = simplified_df,
       aes(x = ingredients , y = diet, fill = ingredients_2, alpha = prop)) +
  geom_raster()+
  scale_fill_manual(values =  c("#93AAD8", "#7EAA55")) +
  theme_pubr() +
  theme(legend.title = element_blank(),
        legend.text = element_text(size=7),
        text = element_text(size=8),
        legend.position = "right",
        panel.grid = element_blank(),
        legend.key.size = unit(0.25, "cm"),
        axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)
        ) +
  labs(y = "Feed scenario", x = "") +
  guides(fill = "none") + 
  coord_flip()

ggsave(here("figures/feed_composition_2.jpg"), width = 7.65, height = 11.88, units = "in")
```



#clear the environment

```{r}
rm(list = ls(all.names = TRUE))

```


