---
title: "Rescaling the pressure data to create the CPI"
output: html_document
date: "2023-07-11"
---
This file is used to rescale all rasters' values for each pressure so that values among feeds may be compared.

```{r}

message("STEP 16 - RESCALING ENVIRONMENTAL PRESSURES")

library(tidyverse)
library(terra)
library(here)
library(parallel)
library(doParallel)
library(foreach)
library(future)
library(furrr)

source(here("src/functions.R"))


select <- dplyr::select
values <- terra::values

#allocation method (un-comment preference)
this_allocation_method <- "econ_allocation"
# this_allocation_method <- "ge_allocation"
# this_allocation_method <- "mass_allocation"

```

Create equal area projection for rescaling
```{r}

# Base raster wgs84
base_rast <- rast(res=0.5)
ext(base_rast) <- c(-180, 180, -90, 90)


#Gall peters projection

gall_peters <- "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
#raster template 
eq_area_rast <- project(base_rast, gall_peters) 
values(eq_area_rast) <- 1:ncell(eq_area_rast)
cellSize(eq_area_rast, unit="km", transform = FALSE)


#Mollweide projection
moll_rast <- project(base_rast, "ESRI:54009")
values(moll_rast) <- 1:ncell(moll_rast)
cellSize(moll_rast, unit="km", transform = FALSE)



# Check the reprojection doesn't appreciably change the values when density is multiplied by cell area

# marine_diet_disturbance <- list.files(here("data/spatial/marine_diet/int"), pattern = "wheat_disturbance_km2_econ_allocation_UKR", full.names = TRUE)
# 
# test <- rast(marine_diet_disturbance)
# (test1 <- test*cellSize(test, unit="km"))
# 
# test_rp <- project(test, moll_rast)
# (test_rp1 <- test_rp*cellSize(test_rp, unit="km"))
# 
# test_rs <- terra::resample(test, moll_rast)
# (test_rs1 <- test_rs*cellSize(test_rs, unit="km"))
# 
# test1 |> values() |> sum(na.rm = TRUE)
# test_rs1 |> values()  |> sum(na.rm = TRUE)
# test_rp1 |> values() |> sum(na.rm = TRUE)


```



## Disturbance

Bring in  forage fish and crop layers for disturbance for both diets. Reproject onto the equal area mollweide projection and calculate the sum of disturbance for each diet.

```{r}

#ID files for disturbance for marine and plant diets
marine_diet_disturbance <- list.files(here("data/spatial/marine_diet/int"), pattern = paste0("disturbance_km2_", this_allocation_method), full.names = TRUE)


#REPROJECT PER KM2 DATA TO GALL PETERS EQUAL AREA

this_file <- marine_diet_disturbance[[25]]

#marine diets
md_reprojected_rasters <- 
  
  map(.x = marine_diet_disturbance, .f = \(this_file){
    
    base <- tools::file_path_sans_ext(basename(this_file))  
    
    this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
    
    this_ingredient <- str_extract(base, "[^_]+")
    
    saveName <- gsub("km2", "moll",this_file)
    
    if(!file.exists(saveName)){
      
      message("processing ", basename(saveName))
      
      this_rast <- rast(this_file)
      
      this_reprojected_rast <-  project(this_rast, moll_rast)
      
      this_adj_reprojected_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km", transform = FALSE)
      names(this_adj_reprojected_rast) <- "km2"
      
      writeRaster(x = this_adj_reprojected_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reprojected_rast)
    }
    
  })






#plant diets
plant_diet_disturbance <- list.files(here("data/spatial/plant_diet/int"), pattern = paste0("disturbance_km2_", this_allocation_method), full.names = TRUE)


this_file <- plant_diet_disturbance[[50]]

pd_reprojected_rasters <- 
  
  map(.x = plant_diet_disturbance, .f = \(this_file){
    
    base <- tools::file_path_sans_ext(basename(this_file))  
    
    this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
    
    this_ingredient <- str_extract(base, "[^_]+")
    
    saveName <- gsub("km2", "moll",this_file)
    
   if(!file.exists(saveName)){
      
      message("processing ", basename(saveName))
      
      this_rast <- rast(this_file)
    
      this_reprojected_rast <-  project(this_rast, moll_rast)
      
      this_adj_reprojected_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km", transform = FALSE)
      names(this_adj_reprojected_rast) <- "km2"
      
      writeRaster(x = this_adj_reprojected_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reprojected_rast)
    }
    
  })


#NOW FIND THE MAXIMUM TOTAL SCORES FOR DISTURBANCE INCLUDING ALL INGREDIENTS FOR EACH DIET AND ALLOCATION METHOD - THIS WILL BE USED TO GET A GIVEN CELLS' PROPORTIONAL CONTRIBUTION

#First need to find maximum scores each ingredient (i.e. which origin yields the greatest)

md_reprojected_rasters <- list.files(path = here("data/spatial/marine_diet/int"), pattern = paste0("disturbance_moll_", this_allocation_method), full.names =TRUE)

md_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> filter(diet=="marine_diet" & ingredient_demand_tonnes>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()

this_df <- md_ingredients_list[[5]]

maximum_disturbance_scores_md <-  
  
  map_df(md_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
  these_ingredient_production_rasters <- md_reprojected_rasters[grepl(this_ingredient, md_reprojected_rasters) & grepl(this_allocation_method, md_reprojected_rasters) & !grepl("processing", md_reprojected_rasters)] |> map(rast)
  
  these_ingredient_processing_rasters <- md_reprojected_rasters[grepl(this_ingredient, md_reprojected_rasters) & grepl(this_allocation_method, md_reprojected_rasters) & grepl("processing", md_reprojected_rasters)] |> map(rast)
  
 these_combined_rasters <- map2(.x=these_ingredient_production_rasters, .y= these_ingredient_processing_rasters, .f= \(this_production_rast, this_processing_rast){
    return(app(c(this_production_rast, this_processing_rast), fun = sum, na.rm=TRUE))
  })
 
 max_disturbance <- these_combined_rasters |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  this_adj_df <- bind_cols(this_df, max_disturbance = max_disturbance)
  
  
})

#get maximum disturbance scores 
max_md_disturbance <- maximum_disturbance_scores_md |> pull(max_disturbance) |> sum()


#PLANT DIETS 
pd_reprojected_rasters <- list.files(path = here("data/spatial/plant_diet/int"), pattern = paste0("disturbance_moll_", this_allocation_method), full=TRUE)


pd_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> filter(diet=="plant_diet" & ingredient_demand_tonnes>0) |> 
  arrange(ingredients) |> 
  group_by(ingredients) |> 
  group_split()

this_df <- md_ingredients_list[[1]]

maximum_disturbance_scores_pd <-  
  
  map_df(pd_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
  these_ingredient_production_rasters <- pd_reprojected_rasters[grepl(this_ingredient, pd_reprojected_rasters) & grepl(this_allocation_method, pd_reprojected_rasters) & !grepl("processing", pd_reprojected_rasters)] |> map(rast)
  
  these_ingredient_processing_rasters <- pd_reprojected_rasters[grepl(this_ingredient, pd_reprojected_rasters) & grepl(this_allocation_method, pd_reprojected_rasters) & grepl("processing", pd_reprojected_rasters)] |> map(rast)
  
 these_combined_rasters <- map2(.x=these_ingredient_production_rasters, .y= these_ingredient_processing_rasters, .f= \(this_production_rast, this_processing_rast){
    return(app(c(this_production_rast, this_processing_rast), fun = sum, na.rm=TRUE))
  })
 
 max_disturbance <- these_combined_rasters |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  this_adj_df <- bind_cols(this_df, max_disturbance = max_disturbance)
  
  
})




#get maximum disturbance scores for each allocation method
max_pd_disturbance <- maximum_disturbance_scores_pd |> pull(max_disturbance) |> sum()


max_total_disturbance <- max(max_md_disturbance,max_pd_disturbance)



# #which diet has the highest disturbance values
# marine_disturbance_vals_ge <- md_reprojected_rasters[grepl("ge_allocation", md_reprojected_rasters)] |> map(rast) |> map(no_na_values) |> unlist()
# plant_disturbance_vals_ge <-  pd_reprojected_rasters[grepl("ge_allocation", pd_reprojected_rasters)] |> map(rast) |> map(no_na_values) |> unlist()
# 
# marine_disturbance_vals_mass <- md_reprojected_rasters[grepl("mass_allocation", md_reprojected_rasters)] |> map(rast) |> map(no_na_values) |> unlist()
# plant_disturbance_vals_mass <-  pd_reprojected_rasters[grepl("mass_allocation", pd_reprojected_rasters)] |> map(rast) |> map(no_na_values) |> unlist()
# 
# 
# marine_diet_total_ge <- sum(marine_disturbance_vals_ge, na.rm = TRUE)
# marine_diet_total_mass <- sum(marine_disturbance_vals_mass, na.rm=TRUE)
# 
# plant_diet_total_ge <- sum(plant_disturbance_vals_ge, na.rm = TRUE)
# plant_diet_total_mass <- sum(plant_disturbance_vals_mass, na.rm=TRUE)
# 
# 
# 
# #check marine is appearing more intense (given forage fish disturbance values are much higher)
# summary(marine_disturbance_vals_ge)
# summary(plant_disturbance_vals_ge)
# 
# summary(marine_disturbance_vals_mass)
# summary(plant_disturbance_vals_mass)
# 
# all_disturbance_vals_ge <- c(marine_disturbance_vals_ge, plant_disturbance_vals_ge) |> unlist()
# all_disturbance_vals_mass <- c(marine_disturbance_vals_mass, plant_disturbance_vals_mass) |> unlist()
# 
# disturbance_energy_skewness <-  3*(mean(all_disturbance_vals_ge)-median(all_disturbance_vals_ge))/sd(all_disturbance_vals_ge)
# disturbance_mass_skewness <- 3*(mean(all_disturbance_vals_mass)-median(all_disturbance_vals_mass))/sd(all_disturbance_vals_mass)
# 
# #determine the quantile for rescaling - given the extreme skew - 99.9% seems 
# hist((all_disturbance_vals_ge), main = "", xlab = "Disturbance values") |> 
# abline(v = quantile(all_disturbance_vals_ge, 0.999), lty=2)
# 
# hist((all_disturbance_vals_mass), main = "", xlab = "Disturbance values") |> 
# abline(v = quantile(all_disturbance_vals_mass, 0.999), lty=2)
# 
# disturbance_rescale_val_ge <- quantile(all_disturbance_vals_ge, 0.999) |> as.numeric()
# disturbance_rescale_val_mass <- quantile(all_disturbance_vals_mass, 0.999) |> as.numeric()


#rescale all rasters by the proportion each cell represents of total possible disturbance for each diet globally (for both energy and mass allocation, this is marine_diet)

#marine diets

this_file <- md_reprojected_rasters[[11]]


rescaled_marine_disturbance_rasters <- 
  
  map(.x= md_reprojected_rasters, .f = \(this_file){
    
    this_rast <- rast(this_file)
    
    this_basename <- str_replace(basename(this_file), pattern = "_moll", replacement = "")
    
    saveName <- sprintf(here("data/spatial/marine_diet/pressures/%s"), this_basename)
    
    if(!file.exists(saveName)){
      
      message("Processing ", "'", this_basename, "'") 
      
        this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_disturbance, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
        }
  })


this_file <- pd_reprojected_rasters[[50]]


#plant diets

rescaled_plant_disturbance_rasters <- 
  
  map(.x= pd_reprojected_rasters, .f = \(this_file){
  
  this_rast <- rast(this_file)
  
  this_basename <- str_replace(basename(this_file), pattern = "_moll", replacement = "")
  
  saveName <- sprintf(here("data/spatial/plant_diet/pressures/%s"), this_basename)
  
  if(!file.exists(saveName)){
    
   message("Processing ", "'", this_basename, "'")
        
        this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_disturbance, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
     
  }
})


```

# Greenhouse gas pressures

```{r}

marine_diet_ghg <- list.files(here("data/spatial/marine_diet/int"), pattern = paste0("ghg_km2_", this_allocation_method), full.names = TRUE)

plant_diet_ghg <- list.files(here("data/spatial/plant_diet/int"), pattern = paste0("ghg_km2_", this_allocation_method), full.names = TRUE)


#REPROJECT PER KM2 DATA TO GALL PETERS EQUAL AREA FOR TOTAL PRESSURES

this_file <- marine_diet_ghg[[1]]


# Marine diets


map(.x = marine_diet_ghg, .f =  \(this_file){
    
    base <- tools::file_path_sans_ext(basename(this_file))  
    
    this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
    
    this_ingredient <- str_extract(base, "[^_]+")
    
    saveName <- gsub("km2", "moll",this_file)
    
    if(!file.exists(saveName)){
      
      message("Processing ", basename(saveName))
      
      this_rast <- rast(this_file)
      
      this_reprojected_rast <- project(this_rast, moll_rast)
      
      this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit="km", transform = FALSE) #multiply up the footprints from per km2 to total based on cellsize
      
      names(this_adj_reproj_rast) <- "kgCO2eq"

      writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reproj_rast)
    }
})


#Plant diets

this_file <- plant_diet_ghg[[1]]

  
map(.x = plant_diet_ghg, .f = \(this_file){
    
    base <- tools::file_path_sans_ext(basename(this_file))  
    
    this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
    
    
    this_ingredient <- str_extract(base, "[^_]+")
    
    saveName <- gsub("km2", "moll",this_file)
    
    if(!file.exists(saveName)){
      
      message("processing ", basename(saveName))
      
      this_rast <- rast(this_file)
      
      this_reprojected_rast <- project(this_rast, moll_rast)
      
      this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit="km", transform = FALSE) #multiply up the footprints from per km2 to total based on cell size
      
      names(this_adj_reproj_rast) <- "kgCO2eq"
      
      writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reproj_rast)
    }
  })
  


#RESCALE GHG DATA BY PROP OF MAXIMUM GLOBAL DISTURBANCE ACROSS DIETS


#NOW FIND THE MAXIMUM TOTAL SCORES FOR DISTURBANCE INCLUDING ALL INGREDIENTS FOR EACH DIET AND ALLOCATION METHOD - THIS WILL BE USED TO GET A GIVEN CELLS' PROPORTIONAL CONTRIBUTION

# First need to find maximum scores each ingredient (i.e. which origin yields the greatest)
# Then need to rescale using the maximum plausible disturbance for each diet as the denominator



## Marine diets


md_reprojected_rasters_ghg <- list.files(path = here("data/spatial/marine_diet/int"), pattern = paste0("ghg_moll_", this_allocation_method), full=TRUE)

md_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> filter(diet=="marine_diet" & ingredient_demand_tonnes>0) |> arrange(ingredients) |> group_by(ingredients) |> group_split()

this_df <- md_ingredients_list[[4]]



maximum_ghg_scores_md <-  
  
  map_df(md_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
  these_ingredient_production_rasters <- md_reprojected_rasters_ghg[grepl(this_ingredient, md_reprojected_rasters_ghg) & grepl(this_allocation_method, md_reprojected_rasters_ghg) & !grepl("processing", md_reprojected_rasters_ghg)] |> map(rast)
  
  these_ingredient_processing_rasters <- md_reprojected_rasters_ghg[grepl(this_ingredient, md_reprojected_rasters_ghg) & grepl(this_allocation_method, md_reprojected_rasters_ghg) & grepl("processing", md_reprojected_rasters_ghg)] |> map(rast)
  
 these_combined_rasters <- map2(.x=these_ingredient_production_rasters, .y= these_ingredient_processing_rasters, .f= \(this_production_rast, this_processing_rast){
    return(app(c(this_production_rast, this_processing_rast), fun = sum, na.rm=TRUE))
  })
 
 max_ghg <- these_combined_rasters |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  this_adj_df <- bind_cols(this_df, max_ghg = max_ghg)
  
  
})

#get maximum marine diet disturbance scores
max_md_ghg<- maximum_ghg_scores_md |> pull(max_ghg) |> sum()




#Plant diets

pd_reprojected_rasters_ghg <- list.files(path = here("data/spatial/plant_diet/int"), pattern = paste0("ghg_moll_", this_allocation_method), full=TRUE)


pd_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> filter(diet=="plant_diet" & ingredient_demand_tonnes>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()

this_df <- md_ingredients_list[[1]]



maximum_ghg_scores_pd <-  
  
  map_df(pd_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
  these_ingredient_production_rasters <- pd_reprojected_rasters_ghg[grepl(this_ingredient, pd_reprojected_rasters_ghg) & grepl(this_allocation_method, pd_reprojected_rasters_ghg) & !grepl("processing", pd_reprojected_rasters_ghg)] |> map(rast)
  
  these_ingredient_processing_rasters <- pd_reprojected_rasters_ghg[grepl(this_ingredient, pd_reprojected_rasters_ghg) & grepl(this_allocation_method, pd_reprojected_rasters_ghg) & grepl("processing", pd_reprojected_rasters_ghg)] |> map(rast)
  
 these_combined_rasters <- map2(.x=these_ingredient_production_rasters, .y= these_ingredient_processing_rasters, .f= \(this_production_rast, this_processing_rast){
    return(app(c(this_production_rast, this_processing_rast), fun = sum, na.rm=TRUE))
  })
 
 max_ghg <- these_combined_rasters |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  
  this_adj_df <- bind_cols(this_df, max_ghg = max_ghg)
  
  
})


#compare teh scores by diet
maximum_ghg_scores_md
maximum_ghg_scores_pd

#get maximum ghg scores for each allocation method
max_pd_ghg<- maximum_ghg_scores_pd |> pull(max_ghg) |> sum()


max_total_ghg <- max(max_md_ghg,max_pd_ghg)



#RESCALE THE TOTAL GHG DATA

#test file
this_file <- md_reprojected_rasters_ghg[[1]]

#Rescale all ghg rasters to the 99.9th percentile

#MARINE DIETS 

rescaled_marine_ghg_rasters <- 
  
  map(.x= md_reprojected_rasters_ghg, .f = \(this_file){
  
  this_rast <- rast(this_file)
  
  this_basename <- str_replace(basename(this_file), pattern = "_moll", replacement = "")
  
  saveName <- sprintf(here("data/spatial/marine_diet/pressures/%s"), this_basename)
  
  if(!file.exists(saveName)){
      
    message("Processing ", "'", this_basename, "'") 
    
    this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
      if_else(this_cell>0, true = this_cell/max_total_ghg, false = 0)})
    
    names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
    
    writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
    
    return(this_rescaled_rast)
      } 
  })



#PLANT DIETS

#test file
this_file <- pd_reprojected_rasters_ghg[[1]]


rescaled_plant_ghg_rasters <- 
  
  map(.x= pd_reprojected_rasters_ghg, .f = \(this_file){
    
    this_rast <- rast(this_file)
    
    this_basename <- str_replace(basename(this_file), pattern = "_moll", replacement = "")
    
    saveName <- sprintf(here("data/spatial/plant_diet/pressures/%s"), this_basename)
    
    if(!file.exists(saveName)){
      
      message("Processing ", "'", this_basename, "'")
      
      this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
        if_else(this_cell>0, true = this_cell/max_total_ghg, false = 0)})
      
      names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
      writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
      return(this_rescaled_rast)
      
    }
    
  }
)



```

# Nutrient pressures

Do first for aggregated nutrients
```{r}
marine_diet_nutrients <- list.files(here("data/spatial/marine_diet/int"), pattern = paste0("nutrient_km2_", this_allocation_method), full.names = TRUE)
plant_diet_nutrients <- list.files(here("data/spatial/plant_diet/int"), pattern = paste0("nutrient_km2_", this_allocation_method), full.names = TRUE)


#REPROJECT PER KM2 DATA TO MOLLWEIDE EQUAL AREA FOR TOTAL PRESSURES

this_file <- marine_diet_nutrients[[1]]


# Marine diets
  
map(.x = marine_diet_nutrients, .f = \(this_file){
    
  base <- tools::file_path_sans_ext(basename(this_file))  
  
  this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
  
  this_ingredient <- str_extract(base, "[^_]+")
  
  saveName <- gsub("km2", "moll",this_file)
  
    if(!file.exists(saveName)){
      
      message("Processing ", basename(saveName))
      
      this_rast <- rast(this_file)
      
      this_reprojected_rast <- project(this_rast, moll_rast)
      
      this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km", transform=FALSE) #multiply up the footprints from per km2 to total based on cellsize
      names(this_adj_reproj_rast) <- "summed MT N/P eq."

      writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reproj_rast)
    }
  })


#Plant diets 

this_file <- plant_diet_nutrients[[1]]

map(.x = plant_diet_nutrients, .f = \(this_file){
    
    base <- tools::file_path_sans_ext(basename(this_file))  
    
    this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
    
    this_ingredient <- str_extract(base, "[^_]+")
    
    saveName <- gsub("km2", "moll",this_file)
    
    if(!file.exists(saveName)){
      
      message("processing ", basename(saveName))
      
      this_rast <- rast(this_file)
      
      this_reprojected_rast <- project(this_rast, gall_peters)
      
      this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km", transform=FALSE) #multiply up the footprints from per km2 to total based on cellsize
      
      names(this_adj_reproj_rast) <- "summed MT N/P eq."
      
      writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reproj_rast)
    }
  })




#RESCALE NUTRIENT DATA BY PROP OF MAXIMUM GLOBAL NUTRIENTS ACROSS DIETS


#NOW FIND THE MAXIMUM TOTAL SCORES FOR NUTRIENTS INCLUDING ALL INGREDIENTS FOR EACH DIET AND ALLOCATION METHOD - THIS WILL BE USED TO GET A GIVEN CELLS' PROPORTIONAL CONTRIBUTION

# First need to find maximum scores each ingredient (i.e. which origin yields the greatest)
# Then need to rescale using the maximum plausible nutrient for each diet as the denominator



## Marine diets

md_reprojected_rasters_nutrient <- list.files(path = here("data/spatial/marine_diet/int"), pattern = paste0("nutrient_moll_", this_allocation_method), full=TRUE)

md_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> filter(diet=="marine_diet" & ingredient_demand_tonnes>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()


#test function
this_df <- md_ingredients_list[[1]]


maximum_nutrient_scores_md <-  
  
  map_df(md_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
  these_ingredient_production_rasters <- md_reprojected_rasters_nutrient[grepl(this_ingredient, md_reprojected_rasters_nutrient) & grepl(this_allocation_method, md_reprojected_rasters_nutrient) & !grepl("processing", md_reprojected_rasters_nutrient)] |> map(rast)
  
  these_ingredient_processing_rasters <- md_reprojected_rasters_nutrient[grepl(this_ingredient, md_reprojected_rasters_nutrient) & grepl(this_allocation_method, md_reprojected_rasters_nutrient) & grepl("processing", md_reprojected_rasters_nutrient)] |> map(rast)
  
 these_combined_rasters <- map2(.x=these_ingredient_production_rasters, .y= these_ingredient_processing_rasters, .f= \(this_production_rast, this_processing_rast){
    return(app(c(this_production_rast, this_processing_rast), fun = sum, na.rm=TRUE))
  })
 
 max_nutrient <- these_combined_rasters |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  this_adj_df <- bind_cols(this_df, max_nutrient = max_nutrient)
  
  
})

#get maximum marine diet nutrient scores for each allocation method
max_md_nutrient <- maximum_nutrient_scores_md |> pull(max_nutrient) |> sum()






#Plant diets

pd_reprojected_rasters_nutrient <- list.files(path = here("data/spatial/plant_diet/int"), pattern = paste0("nutrient_moll_", this_allocation_method), full=TRUE)


pd_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> filter(diet=="plant_diet" & ingredient_demand_tonnes>0) |> arrange(ingredients) |> group_by(ingredients) |> group_split()

this_df <- pd_ingredients_list[[5]]

maximum_nutrient_scores_pd <-  
  
  map_df(pd_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
  these_ingredient_production_rasters <- pd_reprojected_rasters_nutrient[grepl(this_ingredient, pd_reprojected_rasters_nutrient) & grepl(this_allocation_method, pd_reprojected_rasters_nutrient) & !grepl("processing", pd_reprojected_rasters_nutrient)] |> map(rast)
  
  these_ingredient_processing_rasters <- pd_reprojected_rasters_nutrient[grepl(this_ingredient, pd_reprojected_rasters_nutrient) & grepl(this_allocation_method, pd_reprojected_rasters_nutrient) & grepl("processing", pd_reprojected_rasters_nutrient)] |> map(rast)
  
 these_combined_rasters <- map2(.x=these_ingredient_production_rasters, .y= these_ingredient_processing_rasters, .f= \(this_production_rast, this_processing_rast){
    return(app(c(this_production_rast, this_processing_rast), fun = sum, na.rm=TRUE))
  })
 
 max_nutrient <- these_combined_rasters |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  this_adj_df <- bind_cols(this_df, max_nutrient = max_nutrient)
  
  
})

#compare nutrient scores by ingredient by diet
maximum_nutrient_scores_md
maximum_nutrient_scores_pd

#get maximum nutrient scores for each allocation method
max_pd_nutrient <- maximum_nutrient_scores_pd |> pull(max_nutrient) |> sum()

#get maximum total values for nutrients across diets 
max_total_nutrient <- max(max_md_nutrient,max_pd_nutrient)


##rescale all rasters by the proportion each cell represents of total possible nitrogen pollution for each diet globally (for both energy and mass allocation, this is marine_diet)

#marine diets

this_file <- md_reprojected_rasters_nutrient[[1]]

map(.x= md_reprojected_rasters_nutrient, .f = \(this_file){

  this_rast <- rast(this_file)

  this_basename <- str_replace(basename(this_file), pattern = "_moll", replacement = "")

  saveName <- sprintf(here("data/spatial/marine_diet/pressures/%s"), this_basename)

  if(!file.exists(saveName)){

     message("Processing", "'", this_basename, "'")

        this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_nutrient, false = 0)})

        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)

        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)

        return(this_rescaled_rast)
      }

})



#plant diets 

map(.x= pd_reprojected_rasters_nutrient, .f = \(this_file){

  this_rast <- rast(this_file)

  this_basename <- str_replace(basename(this_file), pattern = "_moll", replacement = "")

  saveName <- sprintf(here("data/spatial/plant_diet/pressures/%s"), this_basename)

  if(!file.exists(saveName)){

    message("Processing", "'", this_basename, "'")

        this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_nutrient, false = 0)})

        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)

        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)

        return(this_rescaled_rast) }

})


```

These two sections are for individual nutrients

First for Nitrogen
```{r}
# marine_diet_N <- list.files(here("data/spatial/marine_diet/int"), pattern = "_N_", full.names = TRUE)
# marine_diet_N <- marine_diet_N[!grepl("forage_fish|_gp", marine_diet_N)]
# 
# plant_diet_N <- list.files(here("data/spatial/plant_diet/int"), pattern = "_N_", full.names = TRUE)
# plant_diet_N <- plant_diet_N[!grepl("forage_fish|_gp", plant_diet_N)]
# 
# 
# 
# #REPROJECT PER KM2 DATA TO GALL PETERS EQUAL AREA FOR TOTAL PRESSURES
# 
# this_file <- marine_diet_N[[1]]
# 
# 
# # Marine diets
#   
# map(.x = marine_diet_N, .f = \(this_file){
#     
#     base <- tools::file_path_sans_ext(basename(this_file))  
#     
#      this_country <- if_else(grepl("EPC|NAT|OTH", base), true = str_extract(base, "EPC|NAT|OTH"), false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
#     
#     this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
#     
#     this_ingredient <- str_extract(base, "[^_]+")
#     
#     saveName <- sprintf(here("data/spatial/marine_diet/int/%s_N_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
#     
#     if(!file.exists(saveName)){
#       
#       message("Processing ", basename(saveName))
#       
#       this_rast <- rast(this_file)
#       
#       this_reprojected_rast <- project(this_rast, gall_peters)
#       
#       this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km") #multiply up the footprints from per km2 to total based on cellsize
#       names(this_adj_reproj_rast) <- "N_eq"
# 
#       writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
#       
#       return(this_adj_reproj_rast)
#     }
#   })
# 
# 
# 
# #Plant diets
# 
# this_file <- plant_diet_N[[1]]
# 
#   
# map(.x = plant_diet_N, .f = \(this_file){
#     
#     base <- tools::file_path_sans_ext(basename(this_file))  
#     
#    this_country <- if_else(grepl("EPC|NAT|OTH", base), true = str_extract(base, "EPC|NAT|OTH"), false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
#     
#     this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
#     this_ingredient <- str_extract(base, "[^_]+")
#     
#     saveName <- sprintf(here("data/spatial/plant_diet/int/%s_N_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
#     
#     if(!file.exists(saveName)){
#       
#       message("processing ", basename(saveName))
#       
#       this_rast <- rast(this_file)
#       
#       this_reprojected_rast <- project(this_rast, gall_peters)
#       
#       this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km") #multiply up the footprints from per km2 to total based on cellsize
#       
#       names(this_adj_reproj_rast) <- "N_eq"
#       
#       writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
#       
#       return(this_adj_reproj_rast)
#     }
#   })
# 
# 
# #RESCALE NITROGEN DATA BY PROP OF MAXIMUM GLOBAL NITROGEN ACROSS DIETS
# 
# 
# #NOW FIND THE MAXIMUM TOTAL SCORES FOR NITOGREN INCLUDING ALL INGREDIENTS FOR EACH DIET AND ALLOCATION METHOD - THIS WILL BE USED TO GET A GIVEN CELLS' PROPORTIONAL CONTRIBUTION
# 
# # First need to find maximum scores each ingredient (i.e. which origin yields the greatest)
# # Then need to rescale using the maximum plausible nitrogen for each diet as the denominator
# 
# 
# 
# ## Marine diets
# 
# 
# md_reprojected_rasters_N <- list.files(path = here("data/spatial/marine_diet/int"), pattern = "N_gp", full=TRUE)
# 
# md_ingredients_list <- 
#   readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet=="marine_diet" & total_ingredient_demand>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()
# 
# 
# #test function
# this_df <- md_ingredients_list[[1]]
# 
# 
# 
# maximum_N_scores_md <-  
#   
#   map_df(md_ingredients_list, .f = \(this_df){
#   
#   this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
#   
#   these_ingredient_rasters_ge <- md_reprojected_rasters_N[grepl(this_ingredient, md_reprojected_rasters_N) & grepl("ge_allocation", md_reprojected_rasters_N)]
#   
#   these_ingredient_rasters_mass <- md_reprojected_rasters_N[grepl(this_ingredient, md_reprojected_rasters_N) & grepl("mass_allocation", md_reprojected_rasters_N)]
#   
#   max_N_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   max_N_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   
#   this_adj_df <- bind_cols(this_df, max_N_ge = max_N_ge, max_N_mass= max_N_mass)
#   
#   
# })
# 
# #get maximum marine diet disturbance scores for each allocation method
# max_md_N_ge <- maximum_N_scores_md |> pull(max_N_ge) |> sum()
# max_md_N_mass <- maximum_N_scores_md |> pull(max_N_mass) |> sum()
# 
# 
# 
# 
# #Plant diets
# 
# pd_reprojected_rasters_N <- list.files(path = here("data/spatial/plant_diet/int"), pattern = "N_gp", full=TRUE)
# 
# 
# pd_ingredients_list <- 
#   readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet=="plant_diet" & total_ingredient_demand>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()
# 
# this_df <- pd_ingredients_list[[1]]
# 
# maximum_N_scores_pd <-  
#   
#   map_df(pd_ingredients_list, .f = \(this_df){
#   
#   this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
#   
#   these_ingredient_rasters_ge <- pd_reprojected_rasters_N[grepl(this_ingredient, pd_reprojected_rasters_N) & grepl("ge_allocation", pd_reprojected_rasters_N)]
#   
#   these_ingredient_rasters_mass <- pd_reprojected_rasters_N[grepl(this_ingredient, pd_reprojected_rasters_N) & grepl("mass_allocation", pd_reprojected_rasters_N)]
#   
#   max_N_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   max_N_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   
#   this_adj_df <- bind_cols(this_df, max_N_ge = max_N_ge, max_N_mass= max_N_mass)
#   
#   
# })
# 
# #get maximum N scores for each allocation method
# max_pd_N_ge <- maximum_N_scores_pd |> pull(max_N_ge) |> sum()
# max_pd_N_mass <- maximum_N_scores_pd |> pull(max_N_mass) |> sum()
# 
# 
# max_total_N_ge <- max(max_md_N_ge,max_pd_N_ge)
# max_total_N_mass <- max(max_md_N_mass,max_pd_N_mass)
# 
# 
# 
# 
# ##rescale all rasters by the proportion each cell represents of total possible nitrogen pollution for each diet globally (for both energy and mass allocation, this is marine_diet)
# 
# #marine diets
# 
# this_file <- md_reprojected_rasters_N[[1]]
# 
# map(.x= md_reprojected_rasters_N, .f = \(this_file){
#   
#   this_rast <- rast(this_file)
#   
#   this_basename <- str_replace(basename(this_file), pattern = "_gp", replacement = "")
#   
#   this_allocation_method <- str_extract(string = this_basename, pattern = "ge_allocation|mass_allocation")
#   
#   saveName <- sprintf(here("data/spatial/marine_diet/pressures/%s"), this_basename)
#   
#   if(!file.exists(saveName)){
#     
#      message("Processing", "'", this_basename, "'")
#     
#     if(this_allocation_method=="ge_allocation"){
#         
#         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_N_ge, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#         
#         
#       } else { 
#         
#          this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_N_mass, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#     }
#   }
#   
# })
# 
# 
# 
# #plant diets _ Nitrogen
# 
# map(.x= pd_reprojected_rasters_N, .f = \(this_file){
#   
#   this_rast <- rast(this_file)
#   
#   this_basename <- str_replace(basename(this_file), pattern = "_gp", replacement = "")
#   
#   this_allocation_method <- str_extract(string = this_basename, pattern = "ge_allocation|mass_allocation")
#   
#   saveName <- sprintf(here("data/spatial/plant_diet/pressures/%s"), this_basename)
#   
#   if(!file.exists(saveName)){
#     
#     message("Processing", "'", this_basename, "'")
#     
#    if(this_allocation_method=="ge_allocation"){
#         
#         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_N_ge, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#         
#         
#       } else { 
#         
#          this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_N_mass, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#       
#     }
#     
#   }
#   
# })
```


Repeat for phosphorus

```{r}
# Phosphorus
# marine_diet_P <- list.files(here("data/spatial/marine_diet/int"), pattern = "_P_", full.names = TRUE)
# marine_diet_P <- marine_diet_P[!grepl("forage_fish|_gp", marine_diet_P)]
# 
# plant_diet_P <- list.files(here("data/spatial/plant_diet/int"), pattern = "_P_", full.names = TRUE)
# plant_diet_P <- plant_diet_P[!grepl("forage_fish|_gp", plant_diet_P)]
# 
# 
# 
# #REPROJECT PER KM2 DATA TO GALL PETERS EQUAL AREA FOR TOTAL PRESSURES
# 
# this_file <- marine_diet_P[[1]]
# 
# 
# # Marine diets
#   
# map(.x = marine_diet_P, .f = \(this_file){
#     
#     base <- tools::file_path_sans_ext(basename(this_file))  
#     
#     this_country <- if_else(grepl("EPC|NAT|OTH", base), true = str_extract(base, "EPC|NAT|OTH"), false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
#     
#     this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
#     
#     this_ingredient <- str_extract(base, "[^_]+")
#     
#     saveName <- sprintf(here("data/spatial/marine_diet/int/%s_P_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
#     
#     if(!file.exists(saveName)){
#       
#       message("Processing ", basename(saveName))
#       
#       this_rast <- rast(this_file)
#       
#       this_reprojected_rast <- project(this_rast, gall_peters)
#       
#       this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km") #multiply up the footprints from per km2 to total based on cellsize
#       names(this_adj_reproj_rast) <- "P_eq"
# 
#       writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
#       
#       return(this_adj_reproj_rast)
#     }
#   })
# 
# 
# 
# #Plant diets
# 
# this_file <- plant_diet_P[[1]]
# 
#   
# map(.x = plant_diet_P, .f = \(this_file){
#     
#     base <- tools::file_path_sans_ext(basename(this_file))  
#     
#     this_country <- if_else(grepl("EPC|NAT|OTH", base), true = str_extract(base, "EPC|NAT|OTH"), false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
#     
#     this_allocation_method <- str_extract(base, "ge_allocation|mass_allocation")
#     
#     this_ingredient <- str_extract(base, "[^_]+")
#     
#     saveName <- sprintf(here("data/spatial/plant_diet/int/%s_P_gp_%s_%s.tif"), this_ingredient, this_allocation_method, this_country)
#     
#     if(!file.exists(saveName)){
#       
#       message("processing ", basename(saveName))
#       
#       this_rast <- rast(this_file)
#       
#       this_reprojected_rast <- project(this_rast, gall_peters)
#       
#       this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km") #multiply up the footprints from per km2 to total based on cellsize
#       
#       names(this_adj_reproj_rast) <- "P_eq"
#       
#       writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
#       
#       return(this_adj_reproj_rast)
#     }
#   })
# 
# 
# 
# 
# #RESCALE PHOSPHORUS DATA BY PROP OF MAXIMUM GLOBAL PHOSPHOROUS ACROSS DIETS
# 
# 
# #NOW FIND THE MAXIMUM TOTAL SCORES FOR PHOSPHOROUS INCLUDING ALL INGREDIENTS FOR EACH DIET AND ALLOCATION METHOD - THIS WILL BE USED TO GET A GIVEN CELLS' PROPORTIONAL CONTRIBUTION
# 
# # First need to find maximum scores each ingredient (i.e. which origin yields the greatest)
# # Then need to rescale using the maximum plausible disturbance for each diet as the denominator
# 
# 
# 
# ## Marine diets
# 
# 
# md_reprojected_rasters_P <- list.files(path = here("data/spatial/marine_diet/int"), pattern = "P_gp", full=TRUE)
# 
# md_ingredients_list <- 
#   readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet=="marine_diet" & total_ingredient_demand>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()
# 
# this_df <- md_ingredients_list[[1]]
# 
# 
# 
# maximum_P_scores_md <-  
#   
#   map_df(md_ingredients_list, .f = \(this_df){
#   
#   this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
#   
#   these_ingredient_rasters_ge <- md_reprojected_rasters_P[grepl(this_ingredient, md_reprojected_rasters_P) & grepl("ge_allocation", md_reprojected_rasters_P)]
#   
#   these_ingredient_rasters_mass <- md_reprojected_rasters_P[grepl(this_ingredient, md_reprojected_rasters_P) & grepl("mass_allocation", md_reprojected_rasters_P)]
#   
#   max_P_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   max_P_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   
#   this_adj_df <- bind_cols(this_df, max_P_ge = max_P_ge, max_P_mass= max_P_mass)
#   
#   
# })
# 
# #get maximum marine diet disturbance scores for each allocation method
# max_md_P_ge <- maximum_P_scores_md |> pull(max_P_ge) |> sum()
# max_md_P_mass <- maximum_P_scores_md |> pull(max_P_mass) |> sum()
# 
# 
# 
# 
# #Plant diets
# 
# pd_reprojected_rasters_P <- list.files(path = here("data/spatial/plant_diet/int"), pattern = "P_gp", full=TRUE)
# 
# 
# pd_ingredients_list <- 
#   readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet=="plant_diet" & total_ingredient_demand>0) |> mutate(ingredients = sub(pattern="-wild", replacement="", ingredients)) |> arrange(ingredients) |> group_by(ingredients) |> group_split()
# 
# this_df <- pd_ingredients_list[[1]]
# 
# maximum_P_scores_pd <-  
#   
#   map_df(pd_ingredients_list, .f = \(this_df){
#   
#   this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
#   
#   these_ingredient_rasters_ge <- pd_reprojected_rasters_P[grepl(this_ingredient, pd_reprojected_rasters_P) & grepl("ge_allocation", pd_reprojected_rasters_P)]
#   
#   these_ingredient_rasters_mass <- pd_reprojected_rasters_P[grepl(this_ingredient, pd_reprojected_rasters_P) & grepl("mass_allocation", pd_reprojected_rasters_P)]
#   
#   max_P_ge <- these_ingredient_rasters_ge |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   max_P_mass <- these_ingredient_rasters_mass |> map(rast) |> map(no_na_values) |> map(sum) |> unlist() |> max()
#   
#   this_adj_df <- bind_cols(this_df, max_P_ge = max_P_ge, max_P_mass= max_P_mass)
#   
#   
# })
# 
# #get maximum N scores for each allocation method
# max_pd_P_ge <- maximum_P_scores_pd |> pull(max_P_ge) |> sum()
# max_pd_P_mass <- maximum_P_scores_pd |> pull(max_P_mass) |> sum()
# 
# 
# max_total_P_ge <- max(max_md_P_ge,max_pd_P_ge)
# max_total_P_mass <- max(max_md_P_mass,max_pd_P_mass)
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# # Now rescale pressure by the total plausible phosphorus pollution value 
# 
# #marine diets - Phosphorus
# 
# #this_file <- marine_diet_P[[1]]
# 
# map(.x= md_reprojected_rasters_P, .f = \(this_file){
#   
#   this_rast <- rast(this_file)
#   
#   this_basename <- str_replace(basename(this_file), pattern = "_gp", replacement = "")
#   
#   this_allocation_method <- str_extract(this_basename, "ge_allocation|mass_allocation")
#   
#   saveName <- sprintf(here("data/spatial/marine_diet/pressures/%s"), this_basename)
#   
#   if(!file.exists(saveName)){
#     
#      message("Processing", "'", this_basename, "'")
#     
#     if(this_allocation_method=="ge_allocation"){
#         
#         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_P_ge, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#         
#         
#       } else { 
#         
#          this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_P_mass, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#       
#     }
#     
#     
#    
#   }
#   
# })
# 
# 
# 
# #plant diets - Phosphorus
# 
# map(.x= pd_reprojected_rasters_P, .f = \(this_file){
#   
#   this_rast <- rast(this_file)
#   
#   this_basename <- str_replace(basename(this_file), pattern = "_gp", replacement = "")
#   
#   this_allocation_method <- str_extract(this_basename, "ge_allocation|mass_allocation")
#   
#   saveName <- sprintf(here("data/spatial/plant_diet/pressures/%s"), this_basename)
#   
#   if(!file.exists(saveName)){
#     
#      message("Processing", "'", this_basename, "'")
#     
#     if(this_allocation_method=="ge_allocation"){
#         
#         this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_P_ge, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#         
#         
#       } else { 
#         
#          this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
#           if_else(this_cell>0, true = this_cell/max_total_P_mass, false = 0)})
#         
#         names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
#       
#         writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
#       
#         return(this_rescaled_rast)
#       
#     }
#     
#   }
#   
# })
# 




```


# Water extraction
```{r}
marine_diet_h2o <- list.files(here("data/spatial/marine_diet/int"), pattern = paste0("_water_km2_", this_allocation_method), full.names = TRUE)

plant_diet_h2o <- list.files(here("data/spatial/plant_diet/int"), pattern = paste0("_water_km2_", this_allocation_method), full.names = TRUE)



#REPROJECT PER KM2 DATA TO GALL PETERS EQUAL AREA FOR TOTAL PRESSURES

this_file <- marine_diet_h2o[[1]]


# Marine diets
  
map(.x = marine_diet_h2o, .f = \(this_file){
    
    base <- tools::file_path_sans_ext(basename(this_file))  
    
     this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
    
    this_ingredient <- str_extract(base, "[^_]+")
    
    saveName <- gsub("km2", "moll",this_file)
    
    if(!file.exists(saveName)){
      
      message("Processing ", basename(saveName))
      
      this_rast <- rast(this_file)
      
      this_reprojected_rast <- project(this_rast, moll_rast)
      
      this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km", transform=FALSE) #multiply up the footprints from per km2 to total based on cellsize
      names(this_adj_reproj_rast) <- "m3"

      writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reproj_rast)
    }
  })



#Plant diets

this_file <- plant_diet_h2o[[1]]

  
map(.x = plant_diet_h2o, .f = \(this_file){
    
    base <- tools::file_path_sans_ext(basename(this_file))  
    
     this_country <- if_else(grepl("fishmeal|fish oil", base), 
                            true = str_sub(base, start = nchar(base)-1, end = nchar(base)) , 
                            false = str_sub(base, start = nchar(base)-2, end = nchar(base)))
    
    
    this_ingredient <- str_extract(base, "[^_]+")
    
    saveName <- gsub("km2", "moll",this_file)
    
    if(!file.exists(saveName)){
      
      message("processing ", basename(saveName))
      
      this_rast <- rast(this_file)
      
      this_reprojected_rast <- project(this_rast, moll_rast)
      
      this_adj_reproj_rast <- this_reprojected_rast*cellSize(this_reprojected_rast, unit = "km", transform=FALSE) #multiply up the footprints from per km2 to total based on cellsize
      
      names(this_adj_reproj_rast) <- "m3"
      
      writeRaster(x = this_adj_reproj_rast, filename = saveName, overwrite = TRUE)
      
      return(this_adj_reproj_rast)
    }
  })




#RESCALE WATER DATA BY PROP OF MAXIMUM GLOBAL WATER USE ACROSS DIETS


#NOW FIND THE MAXIMUM TOTAL SCORES FOR WATER INCLUDING ALL INGREDIENTS FOR EACH DIET AND ALLOCATION METHOD - THIS WILL BE USED TO GET A GIVEN CELLS' PROPORTIONAL CONTRIBUTION

# First need to find maximum scores each ingredient (i.e. which origin yields the greatest)
# Then need to rescale using the maximum plausible water consumption for each diet as the denominator



## Marine diets


md_reprojected_rasters_h2o <- list.files(path = here("data/spatial/marine_diet/int"), pattern = paste0("water_moll_", this_allocation_method), full=TRUE)

md_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> filter(diet=="marine_diet" & ingredient_demand_tonnes>0)  |> arrange(ingredients) |> group_by(ingredients) |> group_split()

this_df <- md_ingredients_list[[3]]


maximum_h2o_scores_md <-  
  
  map_df(md_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
  these_ingredient_production_rasters <- md_reprojected_rasters_h2o[grepl(this_ingredient, md_reprojected_rasters_h2o) & grepl(this_allocation_method, md_reprojected_rasters_h2o) & !grepl("processing", md_reprojected_rasters_h2o)] |> map(rast)
  
  these_ingredient_processing_rasters <- md_reprojected_rasters_h2o[grepl(this_ingredient, md_reprojected_rasters_h2o) & grepl(this_allocation_method, md_reprojected_rasters_h2o) & grepl("processing", md_reprojected_rasters_h2o)] |> map(rast)
  
 these_combined_rasters <- map2(.x=these_ingredient_production_rasters, .y= these_ingredient_processing_rasters, .f= \(this_production_rast, this_processing_rast){
    return(app(c(this_production_rast, this_processing_rast), fun = sum, na.rm=TRUE))
  })

  max_h2o <- these_combined_rasters |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  this_adj_df <- bind_cols(this_df, max_h2o= max_h2o)
  
  
})

#get maximum marine diet disturbance scores for each allocation method
max_md_h2o <- maximum_h2o_scores_md |> pull(max_h2o) |> sum()




#Plant diets

pd_reprojected_rasters_h2o <- list.files(path = here("data/spatial/plant_diet/int"), pattern = paste0("water_moll_", this_allocation_method), full=TRUE)



pd_ingredients_list <- 
  readRDS(here("data/tidy_data/demand/total_ingredient_demand_by_diet.rds")) |> 
  filter(diet=="plant_diet" & ingredient_demand_tonnes>0) |> 
  arrange(ingredients) |>
  group_by(ingredients) |> 
  group_split()

this_df <- pd_ingredients_list[[5]]

maximum_h2o_scores_pd <-  
  
  map_df(pd_ingredients_list, .f = \(this_df){
  
  this_ingredient <- paste0(this_df$ingredients |> unique(),"_")
  
 these_ingredient_production_rasters <- pd_reprojected_rasters_h2o[grepl(this_ingredient, pd_reprojected_rasters_h2o) & grepl(this_allocation_method, pd_reprojected_rasters_h2o) & !grepl("processing", pd_reprojected_rasters_h2o)] |> map(rast)
  
  these_ingredient_processing_rasters <- pd_reprojected_rasters_h2o[grepl(this_ingredient, pd_reprojected_rasters_h2o) & grepl(this_allocation_method, pd_reprojected_rasters_h2o) & grepl("processing", pd_reprojected_rasters_h2o)] |> map(rast)
  
 these_combined_rasters <- map2(.x=these_ingredient_production_rasters, .y= these_ingredient_processing_rasters, .f= \(this_production_rast, this_processing_rast){
    return(app(c(this_production_rast, this_processing_rast), fun = sum, na.rm=TRUE))
  })

  max_h2o <- these_combined_rasters |> map(no_na_values) |> map(sum) |> unlist() |> max()
  
  this_adj_df <- bind_cols(this_df, max_h2o = max_h2o)
  
  
})

#get maximum N scores for each allocation method
max_pd_h2o <- maximum_h2o_scores_pd |> pull(max_h2o) |> sum()

max_total_h2o <- max(max_md_h2o,max_pd_h2o)


#Compare water consumption by diet
maximum_h2o_scores_md
maximum_h2o_scores_pd



#Rescale each diet by the max disturbance denomintor for each diet and allocation method

#marine diets - Water

map(.x= md_reprojected_rasters_h2o, .f = \(this_file){
  
  this_rast <- rast(this_file)
  
  this_basename <- str_replace(basename(this_file), pattern = "_moll", replacement = "")
  
  saveName <- sprintf(here("data/spatial/marine_diet/pressures/%s"), this_basename)
  
  if(!file.exists(saveName)){
    
     message("Processing", "'", this_basename, "'")
        
        this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_h2o, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
  }
  
})



#plant diets - water

map(.x= pd_reprojected_rasters_h2o, .f = \(this_file){
  
  this_rast <- rast(this_file)
  
  this_basename <- str_replace(basename(this_file), pattern = "_moll", replacement = "")
  
  saveName <- sprintf(here("data/spatial/plant_diet/pressures/%s"), this_basename)
  
  if(!file.exists(saveName)){
    
     message("Processing", "'", this_basename, "'")
    
        this_rescaled_rast <-  app(x = this_rast, fun = \(this_cell){
          if_else(this_cell>0, true = this_cell/max_total_h2o, false = 0)})
        
        names(this_rescaled_rast) <- tools::file_path_sans_ext(this_basename)
      
        writeRaster(x = this_rescaled_rast, filename = saveName, overwrite=TRUE)
      
        return(this_rescaled_rast)
  }
  
})
```


```{r}
#clear the environment
rm(list = ls(all.names = TRUE))
```
