---
title: "15_combine_pressures"
author: "Rich Cottrell"
date: "31/07/2022"
output: html_document
---


```{r}
library(tidyverse)
library(here)
library(terra)

select <- dplyr::select
values <- terra::values

```


Bring in all ingredients by diet

```{r}
marine_diet <- readRDS(here("data/tidy_data/diet-scenarios/marine_diet.rds")) |> filter(prop>0) |> mutate(ingredients = gsub(pattern = "-wild", replacement = "", x = ingredients)) |> pull(ingredients)

plant_diet <- readRDS(here("data/tidy_data/diet-scenarios/plant_diet.rds")) |> filter(prop>0) |> mutate(ingredients = gsub(pattern = "-wild", replacement = "", x = ingredients)) |> pull(ingredients)


```

Now iterate different versions of diet based on sources - the gtools::permutations function doesn't handle this well so have iterated based on random draws from three soures for each ingredient

Doing this first for marine diets

```{r}
md_ingredient_source_list <- crossing(ingredient = marine_diet, source = 1:5) |> group_by(ingredient) |> group_split()

set.seed(7)

md_df_list <- vector(mode="list", length = 1000)

#iterate a random pull from each ingredient source 2000 times and isolate non duplicates
for(i in 1:1000){
  
  message(i, "/1000")
  
  this_diet_df <- 
    map_df(md_ingredient_source_list, \(this_df){
      
    this_ingredient <- this_df |> pull(ingredient) |> unique()
      
    source_select <- sample(x=1:5, replace = FALSE, size=1)
      
    this_ingredient_source <- this_df |> slice(source_select) |> mutate(rep = i)
    
    })
    
  md_df_list[[i]] <- this_diet_df
}


source_combinations_md <- bind_rows(md_df_list) |> pivot_wider(names_from = rep, values_from = source)
source_combinations_md <- source_combinations_md[!duplicated(as.list(source_combinations_md))]


```

Now repeat the iteration for plant diets
```{r}
pd_ingredient_source_list <- crossing(ingredient = plant_diet, source = 1:5) |> group_split(ingredient)

set.seed(7)

pd_df_list <- vector(mode="list", length = 1000)

#iterate a random pull from each ingredient source 1000 times and isolate non duplicates
for(i in 1:1000){
  
  message("iteration ", i, "/1000")
  
  this_diet_df <- 
    map_df(pd_ingredient_source_list, \(this_df){
      
    this_ingredient <- this_df |> pull(ingredient) |> unique()
      
    source_select <- sample(x=1:5, replace = FALSE, size=1)
      
      this_ingredient_source <- this_df |> slice(source_select) |> mutate(rep = i)
    })
    
  pd_df_list[[i]] <- this_diet_df
}

#pivot wider and remove duplicate columns
source_combinations_pd <- bind_rows(pd_df_list) |> pivot_wider(names_from = rep, values_from = source)
source_combinations_pd <- source_combinations_pd[!duplicated(as.list(source_combinations_pd))]


```

Given the number of unique combinations varies from marine to plant diet - here I take the first 500 diets generated from each and save them for future use.

```{r}

source_combinations_md <- source_combinations_md |> select(1:501)

source_combinations_pd <-  source_combinations_pd |> select(1:501)


#save marine and plant diet combinations for reuse during calculations
saveRDS(object = source_combinations_md, file = here("data/tidy_data/diets/ingredient_source_combinations_marine_diet.rds"))

saveRDS(object = source_combinations_pd, file = here("data/tidy_data/diets/ingredient_source_combinations_plant_diet.rds"))


#clear the environment
rm(list = ls(all.names = TRUE))

```

