---
title: "Explore Halpern layers"
output: html_document
---

Libraries
```{r}
library(tidyverse)
library(here)
library(janitor)
library(raster)
library(countrycode)
library(ggpubr)
library(RColorBrewer)
library(rnaturalearth)



source(here("src/directories.R"))
```

Shapefiles
```{r}

countries_shp <- ne_countries(scale = 110, returnclass = "sf")
coasts <- ne_coastline(scale = 110, returnclass = "sf")

```




Explore production, feed composition and feed conversion ratios

```{r}

production <- read_csv(file.path(food_systems_dir,"aquaculture/marine/STEP1_species_groups/int/tonnes_per_country_group.csv")) %>%
  filter(!(aq_group %in% "bivalves")) %>%
  mutate(aq_group = ifelse(aq_group=="marine_fish_general", "marine-fish-general", aq_group),
         aq_group = ifelse(aq_group=="salmonids", "salmon", aq_group),
         aq_group = ifelse(aq_group=="shrimps_prawns", "shrimp", aq_group)) %>%
  dplyr::select(country, species = aq_group, tonnes=total_tonnes)

salmon_production <- production %>% filter(species=="salmon")

diets <- read_csv(file.path(food_systems_feed_dir, "data", "aquaculture_diet_composition.csv")) %>% mutate(country=countrycode(iso3c, origin = "iso3c", destination = "country.name", warn=TRUE))

salmon_diets <- diets %>% filter(species == "salmon" & country %in% salmon_production$country) #inclusion are the same for all ingredients for all regions

fcr <- read_csv(file.path(food_systems_feed_dir, "data/feed_conversion_aquaculture.csv")) %>% filter(species=="salmon") %>% dplyr::select(species, value)



```

Countries of interest for salmon production
```{r}

salmon_countries <- tibble(unique(salmon_production$country)) %>% 
  rename(country = 1) %>% 
  mutate(iso3c = countrycode(country, origin = "country.name", destination = "iso3c", warn=TRUE))


```


Calculated feed demand from salmon per ingredient


```{r}

salmon_feed_demand <- read_csv(file.path(food_systems_feed_dir, "data/total_aquaculture_feedstuff_consumption.csv")) %>% 
  filter(animal=="salmon")

salmon_feed_demand_agg <- salmon_feed_demand %>% 
  group_by(feedstuff) %>% 
  summarise(value_tonnes = sum(tonnes_feedstuff)) %>% 
  arrange(-value_tonnes)


ggplot(salmon_feed_demand_agg, 
       aes(x=value_tonnes, y=reorder(feedstuff, value_tonnes)))+
  geom_bar(stat = "identity", aes(fill=feedstuff))+
  guides(fill=FALSE)+
  theme(text=element_text(size=10))+
  labs(caption = "Tonnes of feed stuff", y="", x="")+
  theme_pubr()

ggsave(here("explore", "salmon_feed_demand_global.jpg"), device = "jpg", dpi=300, width = 9, height=5)

```

Going to concentrate on soybeans for the next part of the analysis to exemplify one ingredient first. 


Next work out where the soybeans have come from - i.e. traded or domestically produced

```{r}

trade_data <- read_csv(file.path(food_systems_feed_dir, "data/FAO_MAPSPAMcrop_trade_data.csv"))

consumption <- read_csv(file.path(food_systems_feed_dir, "data/system_country_mapspam_tonnes_consumption.csv")) %>%
  rename(iso3c_consuming = iso3c, consuming_crop_consumed_tonnes =tonnes_product)


production_location <- left_join(trade_data, consumption, by=c("iso3c_consuming", "SPAM_super")) %>%
  filter(!is.na(animal_system)) %>%  #crops not fed to animal system in given country
  mutate(consuming_crop_consumed_tonnes = ifelse(is.na(consuming_crop_consumed_tonnes), 0, consuming_crop_consumed_tonnes)) %>%
  mutate(producing_crop_consumed_tonnes = consuming_crop_consumed_tonnes * prop_of_supply)


#filter for just salmon and soybeans
salmon_soy_location <- production_location %>% filter(grepl("salmon",animal_system) & SPAM_super=="soyb")


#summarise by the producing country to get totals coming from 

feed_source <- salmon_soy_location %>%
  group_by(iso3c_producing, animal_system, SPAM_super) %>%
  summarize(consumed_tonnes = sum(producing_crop_consumed_tonnes)) 



```
Next calculate the proportion of crop produced for soybean in salmon feed for each country
```{r}

soy_production_df <- read_csv(file.path(food_systems_feed_dir, "data/MAPSPAMcrop_production.csv")) %>% filter(SPAM_super == "soyb")

soy_production_global <- soy_production_df %>%
  group_by(SPAM_super) %>%
  summarize(tonnes_global = sum(tonnes_producing_crop, na.rm=TRUE))

proportion <- left_join(feed_source, soy_production_df, by=c("iso3c_producing", "SPAM_super")) %>%
  mutate(tonnes_producing_crop = ifelse(is.na(tonnes_producing_crop), 0, tonnes_producing_crop)) %>%
  mutate(prop_produced_for_system = ifelse(consumed_tonnes == 0, 0, consumed_tonnes/tonnes_producing_crop))


```

Identify countries where apparent consumption exceeds production and correct the proportions. Halpern and all have constrauned consumption from a country to not exceed 10% of what is produced. Not sure why 10%.
```{r}

consumption_exceed <- proportion %>%
  group_by(iso3c_producing, SPAM_super) %>%
  mutate(tonnes_consumed_country = sum(consumed_tonnes)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(extra_crop_tonnes = tonnes_producing_crop - tonnes_consumed_country) %>%
  mutate(percent_exceed = ifelse(tonnes_producing_crop == 0, 0, tonnes_consumed_country/tonnes_producing_crop)) %>%
    data.frame()



corrected_prop <- consumption_exceed %>%
  mutate(prop_of_consumption_within_iso = ifelse(tonnes_consumed_country == 0, 0, consumed_tonnes/tonnes_consumed_country)) %>%  # calculate proportion of crop each system eats within a country
  mutate(prop_produced_for_system_adjust = ifelse(percent_exceed >1 , 
                                        ifelse(percent_exceed > 1.1, prop_of_consumption_within_iso*1.1, prop_of_consumption_within_iso*percent_exceed),
                                        prop_produced_for_system))

summary(corrected_prop)

corrected_prop %>%
  group_by(iso3c_producing, SPAM_super) %>%
  summarize(total_adj_prop = sum(prop_produced_for_system_adjust),
            total_prop = sum(prop_produced_for_system)) %>% 
  mutate(check_diff = total_adj_prop == total_prop)

```

For each animal system isolate the consumption included for one country versus that needs to be redistributed or "excluded".

```{r}

tonnes_included_system <- corrected_prop %>%
  mutate(included_tonnes = prop_produced_for_system_adjust * tonnes_producing_crop) %>%
  mutate(lost_crop_tonnes = consumed_tonnes - included_tonnes) %>%
  mutate(included_prop = ifelse(tonnes_producing_crop == 0 , 0, included_tonnes/tonnes_producing_crop))


# check, sum of included and excluded should equal sum of all consumption
tonnes_included_system %>%
  group_by(SPAM_super) %>%
  summarize(included_crop = sum(included_tonnes, na.rm=TRUE),
            excluded_crop = sum(lost_crop_tonnes, na.rm=TRUE)) %>% 
  mutate(prop_redistributed = excluded_crop/(excluded_crop + included_crop)) %>%
           data.frame() 


tonnes_lost_system <- tonnes_included_system %>%
  group_by(SPAM_super, animal_system) %>%
  summarize(lost_crop_tonnes = sum(lost_crop_tonnes, na.rm=TRUE))


#calculate extra crop tonnes in each country, and determine what proportion of extra each country accounts for
tonnes_extra_crop <- corrected_prop %>%
  dplyr::select(iso3c_producing, SPAM_super, extra_crop_tonnes, tonnes_producing_crop) %>%
  unique() %>%
  mutate(extra_crop_tonnes = ifelse(extra_crop_tonnes<0, 0, extra_crop_tonnes)) %>%
  group_by(SPAM_super) %>%
  mutate(excess_crop_global = sum(extra_crop_tonnes)) %>%
  rowwise() %>%
  mutate(prop_excess_crop = ifelse(excess_crop_global == 0, 0, extra_crop_tonnes/excess_crop_global))


# distribute the tonnes lost for each crop/system based on each countries extra production
excess_prop <- merge(tonnes_extra_crop, tonnes_lost_system) %>%
  mutate(tonnes_dispersed = prop_excess_crop * lost_crop_tonnes) %>%
  mutate(prop_dispersed = ifelse(tonnes_producing_crop == 0, 0, tonnes_dispersed/tonnes_producing_crop))


### NOTE: you can get higher total props for countries with very small proportions. This happens only for soybeans and for small production areas.
## this explore this:
excess_excess_prop <- excess_prop %>%
  mutate(extra_dispersed_tonnes = tonnes_dispersed - extra_crop_tonnes) %>%
  mutate(extra_dispersed_tonnes = ifelse(prop_dispersed > 1, extra_dispersed_tonnes, 0)) %>%
  left_join(soy_production_global, by = "SPAM_super") %>%
  mutate(prop_global_crop_production = tonnes_producing_crop/tonnes_global) %>%
  group_by(SPAM_super, animal_system) %>%
  mutate(extra_dispersed_tonnes_by_system_crop = sum(extra_dispersed_tonnes))
  

filter(excess_excess_prop, prop_dispersed>1) %>% data.frame() #nothing for salmon soybean



```


Combine the included and lost proportions for each system and country for plotting below

```{r}

assigned_prop <- tonnes_included_system  %>%
  dplyr::select(iso3c_producing, animal_system, SPAM_super, prop_produced_for_system = included_prop)
excess_prop <- excess_prop %>%
  dplyr::select(iso3c_producing, animal_system, SPAM_super, prop_extra_for_system = prop_dispersed)

total_prop <- left_join(assigned_prop, excess_prop, by=c("iso3c_producing", "animal_system", "SPAM_super")) %>%
  rowwise() %>%
  mutate(total_prop = prop_produced_for_system + prop_extra_for_system)
summary(total_prop)
filter(total_prop, iso3c_producing=="AFG" & SPAM_super=="soyb") %>% data.frame()


check <- left_join(total_prop, soy_production_df) %>%
  mutate(total_tonnes = total_prop*tonnes_producing_crop) %>%
  group_by(animal_system, SPAM_super) %>%
  summarize(est_tonnes = sum(total_tonnes, na.rm=TRUE)) %>%
  left_join(consumption_global) %>%
  rowwise() %>%
  mutate(difference = obs_tonnes - est_tonnes )

# these should be the same
check %>%
  group_by(SPAM_super) %>%
  summarize(est_tonnes =sum(est_tonnes),
            obs_tonnes = sum(obs_tonnes)) %>% data.frame()


```

Summarised version of the proportion of soy for salmon coming from which countries and generate the proportional raster for each country
```{r}


soy_producers <- read_csv(file.path(food_systems_feed_dir, "data/proportion_feed_per_country_system.csv")) %>% filter(SPAM_super=="soyb" & grepl("salmon", animal_system)) %>% arrange(iso3c_producing) %>% 
  mutate(total_prop = if_else(total_prop == 0, NA_real_, total_prop))


rgns <- read_csv(here("data/spatial/food_rgns_xy.csv"), col_types = "ddddcc")



####################
## loop is superfluous for one system and one ingredient but will be useful later for multiple ingredients
#################3

todo <- unique(dplyr::select(soy_producers, SPAM_super, animal_system))


for(i in 1:dim(todo)[1]){
#i=182
crop <- todo$SPAM_super[i]
system <- todo$animal_system[i]

# total country feed production
feed_crop <- filter(soy_producers, SPAM_super==crop, animal_system==system) %>%
  dplyr::select(iso3c = iso3c_producing, prop_feed = total_prop)

# proportion crop grown for feed per country
prop_feed_crop <- left_join(rgns, feed_crop, by="iso3c") %>%
  dplyr::select(x, y, prop_feed)

prop_feed_crop_raster <- rasterFromXYZ(prop_feed_crop)
# plot(prop_feed_crop_raster)
crs(prop_feed_crop_raster) <- "+proj=longlat +datum=WGS84"
writeRaster(prop_feed_crop_raster, filename= sprintf("/mnt/rdsi/feed_pressure_mapping/data/spatial/marine_%s_x_land_%s_crop_produce.tif", system, crop), overwrite=TRUE)


}

#check it plots OK. Produces a raster of the proportion of that country's production is assigned to salmon soy-based feed
plot(log(prop_feed_crop_raster+1))


```


Apply proportional quantity to growing regions' production

```{r}

crop_list <- "soyb"

total_feed_by_crop <- data.frame(SPAM_super = crop_list, calc_tonnes_from_rast = NA)

for(crop in crop_list){ #crop="soyb"
  all_props <- list.files(here("data/spatial"), pattern="_x_", full=TRUE)
  all_props <- grep(crop, all_props, value=TRUE)
  #all_props <- grep("human_human_human", all_props, invert=TRUE, value=TRUE)
  #all_crop_stack <- raster::stack(all_props)
  #all_crop_sum <- raster::calc(all_crop_stack, sum, na.rm=TRUE)
  all_props <- raster(all_props)
  
  
  #plot(all_crop_sum)
  #click(all_crop_sum)
  prod <- raster(sprintf(here("data/spatial/crop_%s_A_scaled.tif"), crop))
  prod[prod==0] <- NA
  
  tonnes_consumed <- prod*all_props
  tonnes_consumed[tonnes_consumed==0] <- NA
  total_tonnes_consumed <- cellStats(tonnes_consumed, "sum", na.rm=TRUE)
  total_feed_by_crop$calc_tonnes_from_rast[total_feed_by_crop$SPAM_super==crop] <- total_tonnes_consumed
}


soy_production_raster <- prod %>% as("SpatialPixelsDataFrame") %>% data.frame()


soy_production <- ggplot()+
  geom_sf(data = countries_shp %>% filter(iso_a3!="ATA"), fill="grey90", colour=NA)+
  geom_tile(data =soy_production_raster, aes(x=x, y=y, fill=(crop_soyb_A_scaled+1)))+
  scale_fill_gradientn(colours = brewer.pal(9, name = "YlGn"),
                       trans="log10")+
  theme_bw()+
   theme(panel.grid = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal",
        legend.box.spacing = unit(0.1, "cm"),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(colour = "grey70", size=0.5),
        plot.margin = unit(c(0,r=0.10,-0.1,0), "cm"),
         legend.text = element_text(size=6),
        legend.title = element_text(size = 6))+
  labs(fill="Soy production (tonnes)")+
  guides(fill = guide_colorbar(barheight = unit(0.1, "cm"), title.position = "top", title.hjust = 0.5))

ggsave(here("explore/soy_production_map.jpg"), dpi = 300, device = "jpg", width = 7, height=4)



tonnes_consumed_spatial_df <- tonnes_consumed %>% as("SpatialPixelsDataFrame") %>% data.frame()

soy_demand <- ggplot()+
  geom_sf(data = countries_shp %>% filter(iso_a3!="ATA"), fill="grey90", colour=NA)+
  geom_tile(data =tonnes_consumed_spatial_df, aes(x=x, y=y, fill=layer+1))+
  scale_fill_gradientn(colours = brewer.pal(9, name = "PuBu"),
                       trans = "log10", 
                       labels = c(0.001, 10, 100, 1000),
                       breaks = c(1+0.001, 10, 100, 1000)
                       )+
  theme_bw()+
   theme(panel.grid = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal",
        legend.box.spacing = unit(0.1, "cm"),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(colour = "grey70", size=0.5),
        plot.margin = unit(c(0,r=0.10,-0.1,0), "cm"),
         legend.text = element_text(size=6),
        legend.title = element_text(size = 6))+
  labs(fill="Soy demand (tonnes)")+
 guides(fill = guide_colorbar(barheight = unit(0.1, "cm"), title.position = "top", title.hjust = 0.5))


ggsave(here("explore/salmon_soy_demand_map.jpg"), dpi = 300, device = "jpg", width = 7, height=4)



```

Import tif from pressure per tonne data layers

```{r}

soy_dist_raster <- raster(here("data/spatial/land_soyb_crop_produce_disturbance_per_tonne_production.tif"))

soy_ghg_raster <- raster(here("data/spatial/land_soyb_crop_produce_ghg_per_tonne_production.tif"))

soy_h2o_raster <- raster(here("data/spatial/land_soyb_crop_produce_water_per_tonne_production.tif"))

soy_nutr_raster <- raster(here("data/spatial/land_soyb_crop_produce_nutrient_per_tonne_production.tif"))



crop_soyb <- raster(here("data/spatial/crop_soyb.tiff"))


```

Create new rasters from press x tonnage calculations
```{r}

crs(tonnes_consumed) <- crs(soy_dist_raster) #ensure the soy demand is in the same crs as the pressures

soy_dist_product <- soy_dist_raster*tonnes_consumed

soy_ghg_product <- soy_ghg_raster*tonnes_consumed

soy_h2o_product <- soy_h2o_raster*tonnes_consumed

soy_nutr_product <- soy_nutr_raster*tonnes_consumed



```

```{r}
#disturbance
soy_dist_df <- soy_dist_product %>% as("SpatialPixelsDataFrame") %>% data.frame()


soy_dist <- ggplot()+
  geom_sf(data = countries_shp %>% filter(iso_a3!="ATA"), fill="grey85", colour=NA)+
  geom_tile(data = soy_dist_df, aes(x=x, y=y, fill=layer))+
  theme_bw()+
  scale_fill_gradientn(colours = brewer.pal(9, name = "YlOrBr"))+
  theme(panel.grid = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal",
        legend.box.spacing = unit(0.1, "cm"),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(colour = "grey70", size=0.5),
        plot.margin = unit(c(0,r=0.10,-0.1,0), "cm"),
        plot.subtitle = element_text(size=6, vjust=-1),
         legend.text = element_text(size=6),
        legend.title = element_text(size = 6))+
  labs(fill="km2 eq.", subtitle = "Disturbance")+
  guides(fill = guide_colorbar(barheight = unit(0.1, "cm"), title.position = "top", title.hjust = 0.5))


#ghgs 

soy_ghg_df <- soy_ghg_product %>% as("SpatialPixelsDataFrame") %>% data.frame()

soy_ghg <- ggplot()+
  geom_sf(data = countries_shp %>% filter(iso_a3!="ATA"), fill="grey85", colour=NA)+
  geom_tile(data = soy_ghg_df, aes(x=x, y=y, fill=layer))+
  theme_bw()+
  scale_fill_gradientn(colours = brewer.pal(9, name = "RdPu"))+
  theme(panel.grid = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal",
        legend.box.spacing = unit(0.1, "cm"),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(colour = "grey70", size=0.5),
        plot.margin = unit(c(0,0,-0.1,0.1), "cm"),
          plot.subtitle = element_text(size=6, vjust=-1),
         legend.text = element_text(size=6),
        legend.title = element_text(size =6))+
  labs(fill="kg CO2 eq.", subtitle =  "GHGs")+
  guides(fill = guide_colorbar(barheight = unit(0.1, "cm"), title.position = "top", title.hjust = 0.5))



#water 

soy_h2o_df <- soy_h2o_product %>% as("SpatialPixelsDataFrame") %>% data.frame()

soy_h2o <- ggplot()+
  geom_sf(data = countries_shp %>% filter(iso_a3!="ATA"), fill="grey85", colour=NA)+
  geom_tile(data = soy_h2o_df, aes(x=x, y=y, fill=layer/1000))+
  theme_bw()+
  scale_fill_gradientn(colours = brewer.pal(9, name = "Purples"))+
  theme(panel.grid = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal",
        legend.box.spacing = unit(0.1, "cm"),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(colour = "grey70", size=0.5),
        plot.margin = unit(c(t=0,r= 0.1, b=-0.1,l=0.1), "cm"),
         plot.subtitle = element_text(size=6, vjust=-1),
         legend.text = element_text(size=6),
        legend.title = element_text(size = 6))+
  labs(fill="1000s m3", subtitle = "Water")+
  guides(fill = guide_colorbar(barheight = unit(0.1, "cm"), title.position = "top", title.hjust = 0.5))

#nutrients

soy_nutr_df <- soy_nutr_product %>% as("SpatialPixelsDataFrame") %>% data.frame()

soy_nutr <- ggplot()+
  geom_sf(data = countries_shp %>% filter(iso_a3!="ATA"), fill="grey85", colour=NA)+
  geom_tile(data = soy_nutr_df, aes(x=x, y=y, fill=layer))+
  theme_bw()+
  scale_fill_gradientn(colours = brewer.pal(9, name = "Oranges"))+
  theme(panel.grid = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal",
        legend.box.spacing = unit(0, "cm"),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(colour = "grey70", size=0.5),
        plot.margin = unit(c(0,r=0.1,-0.1,0), "cm"),
         plot.subtitle = element_text(size=6, vjust=-1),
         legend.text = element_text(size=6),
        legend.title = element_text(size = 6))+
  labs(fill="po4 eq.", subtitle = "Nutrients")+
  guides(fill = guide_colorbar(barheight = unit(0.1, "cm"), title.position = "top", title.hjust = 0.5))


```
Combine stressor plots
```{r}
(soy_production+soy_demand)/
  (soy_dist+soy_h2o)/
  (soy_nutr+soy_ghg)


ggsave(filename = here("explore/composite_stressor_plot.jpg"), device="jpg", dpi=600, height = 18, width = 18, units="cm")

ggsave(filename = here("explore/composite_stressor_plot.pdf"), device="pdf", dpi=600, height = 16, width = 18, units="cm")
```

