---
title: "19 CPI by ingredient (energetic allocation)"
author: "Rich Cottrell"
date: "14/02/2023"
output: html_document
---


Set up
```{r}
library(tidyverse)
library(ggpubr)
library(patchwork)
library(here)
library(sf)
library(terra)
library(rnaturalearth)
#library(rnaturalearthdata)
library(RColorBrewer)
library(gameofthrones)
library(parallel)
library(ggdist)
library(data.table)
library(dtplyr)


# taking cumulative impoact palette from Halpern et al 
red <- "#B90000"
d_brown <- "#515256"
m_brown <- "#B27B54"
l_brown <- "#BC995F"
green <- "#A8B072"
yellow<- "#EFCE71"
l_yellow <- "#F7F6C1"
light_gray <- "#F8F9FA"
discrete_pal <- c(l_yellow, yellow, green, l_brown, m_brown, d_brown, "firebrick")
continuous_pal <-  colorRampPalette(discrete_pal, space="Lab", bias = 3.5)(10000)
final_palette <- c(light_gray, continuous_pal, red)

#source files
source(here("src/fxns.R"))

#spatial products
moll_crs <- "ESRI:53009"
gp_crs <- "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"


countries <-    ne_countries(scale = "medium", returnclass = "sf") |> st_transform(crs = moll_crs)
bbox <- ne_download(scale = 50, type = "wgs84_bounding_box", category = "physical", returnclass = "sf") |> st_transform(crs = moll_crs)
coastline <- ne_coastline(scale = 50, returnclass = "sf") |> st_transform(crs = moll_crs)
select <- dplyr::select
values <- terra::values
```


Bring in the origin-ingredient combinations within each diet.  

```{r}
marine_diets <- readRDS(here("data/tidy_data/diets/ingredient_source_combinations_marine_diet.rds")) 
plant_diets <- readRDS(here("data/tidy_data/diets/ingredient_source_combinations_plant_diet.rds")) 
```


#Disaggregation by pressure and ingredient


Disaggregate by ingredients and pressures -  Marine diets

```{r}
marine_source_ingredient <- 
  marine_diets |>
  select(1:501) |> 
  rename_at(vars(-ingredient), ~ paste("source", .x)) |>
  mutate(row = row_number()) |> 
  mutate_at(vars(-ingredient, -row), ~ case_when(ingredient == "guar meal"~ as.integer(1),
                                           TRUE ~ .)) |> 
  group_split(row)


sources <- bind_rows(marine_source_ingredient) |> select(-c("ingredient", "row")) |>  names()  

marine_source_file_list <- list()
# Test data for loop
# this_df <- marine_source_ingredient[[1]]
# source <- sources[[1]]
#Now list the files for all ingredients and sources
for(source in sources){
    
    marine_diet_files <- 
      map(.x = marine_source_ingredient, .f = \(this_df){
        
        this_ingredient <- this_df$ingredient
        
        message("processing...", this_ingredient, "...", source)
        
        these_ingredient_files <- list.files(here("data/spatial/marine_diet/pressures"), pattern = paste0(this_ingredient, "_"), full.names = TRUE)
        
         these_ingredient_files <- these_ingredient_files[grepl("ge_allocation", these_ingredient_files)]
        #sources
        this_source_ingredient <- this_df |> pull(source)
        
        #disturbance
        this_disturbance_file <- these_ingredient_files[grepl("disturbance", these_ingredient_files)][[this_source_ingredient]]
        
        #ghg
        this_ghg_file <- these_ingredient_files[grepl("ghg", these_ingredient_files)][[this_source_ingredient]]
        
        #nitrogen
        this_N_file <- these_ingredient_files[grepl("_N_", these_ingredient_files)][[this_source_ingredient]]
        
        #phosphorus
        this_P_file <- these_ingredient_files[grepl("_P_", these_ingredient_files)][[this_source_ingredient]]
        
        #water
        this_h2o_file <- these_ingredient_files[grepl("water", these_ingredient_files)][[this_source_ingredient]]
        
        
        these_diet_pressures <- c(this_disturbance_file, this_ghg_file, this_N_file, this_P_file, this_h2o_file)
        return(these_diet_pressures)
        
      }) |> 
      unlist()
    
    marine_source_file_list[[source]] <- marine_diet_files
    
  }
  
#stack and sum by individual pressures for each ingredient source combination
marine_ingredients <- readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> 
  filter(diet=="marine_diet" & total_ingredient_demand>0) |> 
  mutate(ingredients = gsub(pattern = "-wild", replacement = "", x = ingredients)) |> pull(ingredients)

this_ingredient_pressure_df_list <- list()

all_ingredients_pressures_df_list <- list()

for(source in sources){
  
  message("processing.....", source)
  
  this_combination_all_pressures_list <- marine_source_file_list[[source]]
  
  for(ingredient in marine_ingredients){
    
    this_ingredient_pressure_list <- this_combination_all_pressures_list[grepl(ingredient, this_combination_all_pressures_list)]
    
    #########Separate GHG files, read rasters into a list, stack all GHG layers
  
  message("stacking ghg layers for...", source, " ", ingredient)
  
  this_combination_ghg_list <- this_ingredient_pressure_list[grep("ghg", this_ingredient_pressure_list)]
  
  these_ghg_rasters <- map(.x = this_combination_ghg_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_ghg_df <- global(app(rast(these_ghg_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "GHG", source = source, ingredient = gsub(pattern = "_", replacement = "", ingredient), .before = sum)
  
  
  
  ##########Separate water files, read rasters into a list, stack all water layers
  
  message("stacking water layers for...", source, " ", ingredient)
  
  this_combination_water_list <- this_ingredient_pressure_list[grep("water", this_ingredient_pressure_list)]
  
  these_water_rasters <- map(.x = this_combination_water_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_water_df <- global(app(rast(these_water_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Water", source = source, ingredient = gsub(pattern = "_", replacement = "", ingredient), .before = sum)
  
  
  #########Separate N files, read rasters into a list, stack all N layers
  
  message("stacking N layers for...", source, " ", ingredient)
  
  this_combination_N_list <-  this_ingredient_pressure_list[grep("_N_", this_ingredient_pressure_list)]
  
  these_N_rasters <- map(.x = this_combination_N_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_N_df <- global(app(rast(these_N_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Nitrogen", source = source, ingredient = gsub(pattern = "_", replacement = "", ingredient), .before = sum)
  
  
  #########Separate P files, read rasters into a list, stack all P layers
  
  message("stacking P layers for...", source, " ", ingredient)
  
  this_combination_P_list <-  this_ingredient_pressure_list[grep("_P_", this_ingredient_pressure_list)]
  
  these_P_rasters <- map(.x = this_combination_P_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_P_df <- global(app(rast(these_P_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Phosphorus", source = source, ingredient = gsub(pattern = "_", replacement = "", ingredient), .before = sum)
  
  
  #########Separate disturbance files, read rasters into a list, stack all disturbance layers
  
  message("stacking disturbance layers for...", source, " ", ingredient)
  
  this_combination_disturbance_list <-  this_ingredient_pressure_list[grep("disturbance", this_ingredient_pressure_list)]
  
  these_disturbance_rasters <- map(.x = this_combination_disturbance_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_disturbance_df <- global(app(rast(these_disturbance_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Disturbance", source = source, ingredient = gsub(pattern = "_", replacement = "", ingredient), .before = sum)
  
  
  this_ingredient_pressure_df_list[[ingredient]] <- bind_rows(this_ghg_df, this_water_df, this_N_df, this_P_df, this_disturbance_df)
  
    
    
  }
  
  
  all_ingredients_pressures_df_list[[source]] <- bind_rows(this_ingredient_pressure_df_list)
  
}
marine_diet_pressures_by_ingredient_df <- bind_rows(all_ingredients_pressures_df_list) |> mutate(diet = "marine diet")
saveRDS(marine_diet_pressures_by_ingredient_df, file = here("data/tidy_data/pressures/marine_disaggregated_pressures_by_ingredient_df_ge_allocation.rds"))
```

Disaggreagte by ingredients and pressures -  Plant diets

```{r}
plant_source_ingredient <- 
  plant_diets |>
  select(1:501) |> 
  rename_at(vars(-ingredient), ~ paste("source", .x)) |>
  mutate(row = row_number()) |> 
  mutate_at(vars(-ingredient, -row), ~ case_when(ingredient == "linseed oil"~ as.integer(1),
                                           TRUE ~ .)) |> 
  group_split(row)

sources <- bind_rows(plant_source_ingredient) |> select(-c("ingredient", "row")) |>  names()   

plant_source_file_list <- list()
#test data for loop
# this_df <- plant_source_ingredient[[1]]
# source <- sources[[1]]
#Now list the files for all ingredients and sources
for(source in sources){
    
    plant_diet_files <- 
      map(.x = plant_source_ingredient, .f = \(this_df){
        
        this_ingredient <- this_df$ingredient
        
        message("processing...", this_ingredient, "...", source)
        
        these_ingredient_files <- list.files(here("data/spatial/plant_diet/pressures"), pattern = paste0(this_ingredient, "_"), full.names = TRUE)
        
        these_ingredient_files <- these_ingredient_files[grepl("ge_allocation", these_ingredient_files)]
        
        #sources
        this_source_ingredient <- this_df |> pull(source)
        
        #disturbance
        this_disturbance_file <- these_ingredient_files[grepl("disturbance", these_ingredient_files)][[this_source_ingredient]]
        
        #ghg
        this_ghg_file <- these_ingredient_files[grepl("ghg", these_ingredient_files)][[this_source_ingredient]]
        
        #nitrogen
        this_N_file <- these_ingredient_files[grepl("_N_", these_ingredient_files)][[this_source_ingredient]]
        
        #phosphorus
        this_P_file <- these_ingredient_files[grepl("_P_", these_ingredient_files)][[this_source_ingredient]]
        
        #water
        this_h2o_file <- these_ingredient_files[grepl("water", these_ingredient_files)][[this_source_ingredient]]
        
        
        these_diet_pressures <- c(this_disturbance_file, this_ghg_file, this_N_file, this_P_file, this_h2o_file)
        return(these_diet_pressures)
        
      }) |> 
      unlist()
    
    plant_source_file_list[[source]] <- plant_diet_files
    
  }
  
plant_ingredients <- readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> 
  filter(diet=="plant_diet" & total_ingredient_demand>0) |> 
  mutate(ingredients = gsub(pattern = "-wild", replacement = "", x = ingredients)) |> pull(ingredients)
this_ingredient_pressure_df_list <- list()
all_ingredients_pressures_df_list <- list()
for(source in sources){
  
  message("processing.....", source)
  
  this_combination_all_pressures_list <- plant_source_file_list[[source]]
  
  for(ingredient in plant_ingredients){
    
    this_ingredient_pressure_list <- this_combination_all_pressures_list[grepl(ingredient, this_combination_all_pressures_list)]
    
    #########Separate GHG files, read rasters into a list, stack all GHG layers
  
  message("stacking ghg layers for...", source, " ", ingredient)
  
  this_combination_ghg_list <- this_ingredient_pressure_list[grep("ghg", this_ingredient_pressure_list)]
  
  these_ghg_rasters <- map(.x = this_combination_ghg_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_ghg_df <- global(app(rast(these_ghg_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "GHG", source = source, ingredient = gsub(pattern = "_", replacement = "", ingredient), .before = sum)
  
  
  
  ##########Separate water files, read rasters into a list, stack all water layers
  
  message("stacking water layers for...", source, " ", ingredient)
  
  this_combination_water_list <- this_ingredient_pressure_list[grep("water", this_ingredient_pressure_list)]
  
  these_water_rasters <- map(.x = this_combination_water_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_water_df <- global(app(rast(these_water_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Water", source = source, ingredient = gsub(pattern = "_", replacement = "", ingredient), .before = sum)
  
  
  #########Separate N files, read rasters into a list, stack all N layers
  
  message("stacking N layers for...", source, " ", ingredient)
  
  this_combination_N_list <-  this_ingredient_pressure_list[grep("_N_", this_ingredient_pressure_list)]
  
  these_N_rasters <- map(.x = this_combination_N_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_N_df <- global(app(rast(these_N_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Nitrogen", source = source, ingredient = gsub(pattern = "_", replacement = "", ingredient), .before = sum)
  
  
  #########Separate P files, read rasters into a list, stack all P layers
  
  message("stacking P layers for...", source, " ", ingredient)
  
  this_combination_P_list <-  this_ingredient_pressure_list[grep("_P_", this_ingredient_pressure_list)]
  
  these_P_rasters <- map(.x = this_combination_P_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_P_df <- global(app(rast(these_P_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Phosphorus", source = source, ingredient = gsub(pattern = "_", replacement = "", ingredient), .before = sum)
  
  
  #########Separate disturbance files, read rasters into a list, stack all disturbance layers
  
  message("stacking disturbance layers for...", source, " ", ingredient)
  
  this_combination_disturbance_list <-  this_ingredient_pressure_list[grep("disturbance", this_ingredient_pressure_list)]
  
  these_disturbance_rasters <- map(.x = this_combination_disturbance_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_disturbance_df <- global(app(rast(these_disturbance_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Disturbance", source = source, ingredient = gsub(pattern = "_", replacement = "", ingredient), .before = sum)
  
  
  this_ingredient_pressure_df_list[[ingredient]] <- bind_rows(this_ghg_df, this_water_df, this_N_df, this_P_df, this_disturbance_df)
  
    
    
  }
  
  
  all_ingredients_pressures_df_list[[source]] <- bind_rows(this_ingredient_pressure_df_list)
  
}
plant_diet_pressures_by_ingredient_df <- bind_rows(all_ingredients_pressures_df_list) |> mutate(diet = "plant diet")
saveRDS(plant_diet_pressures_by_ingredient_df, file = here("data/tidy_data/pressures/plant_disaggregated_pressures_by_ingredient_df_ge_allocation.rds"))
```

Add marine and plant-dominant diet disaggregation by ingredient and pressure

```{r}
marine_diet_pressures_by_ingredient_df <- readRDS(here("data/tidy_data/pressures/marine_disaggregated_pressures_by_ingredient_df_ge_allocation.rds"))

plant_diet_pressures_by_ingredient_df <- readRDS(here("data/tidy_data/pressures/plant_disaggregated_pressures_by_ingredient_df_ge_allocation.rds"))

marine_plant_pressure_ingredient_disagg_combined_df <- bind_rows(marine_diet_pressures_by_ingredient_df, plant_diet_pressures_by_ingredient_df)

saveRDS(object = marine_plant_pressure_ingredient_disagg_combined_df, file = here("data/tidy_data/pressures/marine_plant_disaggregated_pressures_by_ingredient_combined_df_ge_allocation.rds"))
```


Visualise the disaggregation by pressure and ingredient

```{r}
diets <- bind_rows(readRDS(here("data/tidy_data/diet-scenarios/marine_diet.rds")) |> 
                     filter(prop >0) |> 
                     add_row(groups = "Microingredients", ingredients = "microingredients", diet = "marine_diet", prop = 0.02) |> 
                     mutate(ingredients = case_when(ingredients == "canola/camelina oil" ~ "Canola oil",
                                                    TRUE ~ str_to_sentence(ingredients)),
                            ingredients = gsub(pattern = "-wild", replacement = "", ingredients)) |> 
                     mutate(ingredients = factor(ingredients, levels = c("Fishmeal", "Fish oil", "Faba beans", "Soybean meal", "Corn gluten meal", "Wheat gluten", "Wheat", "Microingredients"))),
                   readRDS(here("data/tidy_data/diet-scenarios/plant_diet.rds")) |> 
                     filter(prop >0) |> 
                     add_row(groups = "Microingredients", ingredients = "microingredients", diet = "plant_diet", prop = 0.04) |> 
                     mutate(ingredients = case_when(ingredients == "canola/camelina oil" ~ "Canola oil",
                                                    TRUE ~ str_to_sentence(ingredients)),
                            ingredients = gsub(pattern = "-wild", replacement = "", ingredients)) |> 
                     mutate(ingredients = factor(ingredients, levels = c("Fishmeal", "Fish oil", "Soy protein concentrate", "Corn gluten meal","Faba beans", "Pea protein concentrate",  "Sunflower meal", "Wheat gluten", "Canola oil", "Linseed oil", "Wheat", "Pea starch", "Corn starch", "Microingredients")))) |> 
  mutate(diet = gsub("_", " ", diet)) |> 
  rename(diet_prop = prop)


marine_plant_pressure_ingredient_disagg_combined_df <- readRDS(here("data/tidy_data/pressures/marine_plant_disaggregated_pressures_by_ingredient_combined_df_ge_allocation.rds"))

marine_plant_pressure_ingredient_disagg_combined_df <- 
  marine_plant_pressure_ingredient_disagg_combined_df |> 
  group_by(diet, source) |> 
  nest() |> 
  mutate(total = map(data, ~(sum(.$sum)))) |> 
  unnest(c(data, total)) |> 
  mutate(prop = sum/total) |> 
  mutate(ingredient = str_to_sentence(ingredient)) |> 
  left_join(y = diets, by = c("diet", "ingredient" = "ingredients")) |> 
  mutate(weighted_prop = prop/diet_prop)


#Weighted contributions

summarised_ingredient_pressure_df <- 
  marine_plant_pressure_ingredient_disagg_combined_df |> 
  group_by(diet, pressure, ingredient) |> 
  summarise(mean = mean(weighted_prop, na.rm=TRUE),
            sd = sd(weighted_prop, na.rm=TRUE),
            se = sd/sqrt(500),
            median = median(weighted_prop, na.rm = TRUE),
            lq = quantile(weighted_prop, 0.25),
            uq = quantile(weighted_prop, 0.75), .groups = "drop") |> 
  mutate(diet = if_else(diet == "marine diet", true = "Fish-dominant feed", false = "Plant-dominant feed")) |> 
  mutate(ingredient = if_else(ingredient == "foragefish", true = "Forage fish", false = ingredient)) |> 
  mutate(pressure = factor(pressure, levels = c("Water", "Phosphorus", "Nitrogen", "GHG", "Disturbance")))


(marine_diet_ingredients_p <- 
  ggplot(data = summarised_ingredient_pressure_df |> 
           filter(diet == "Fish-dominant feed"),
                aes(x = reorder(ingredient, mean), y = mean, fill=pressure))+
  coord_flip(clip="off")+
  geom_col(aes(fill=pressure), position = "dodge", width = 0.75, alpha = 0.8)+
  geom_errorbar(aes(x = ingredient, ymin = if_else(mean-sd<0, true = 0, false = mean-sd), ymax = mean+sd), position = position_dodge(width=0.75), width = 0, linewidth = 0.2)+
  scale_fill_manual(values = rev(rcartocolor::carto_pal(n=5, name = "Safe")))+
  scale_y_continuous(expand = c(0,0), limits = c(0,2))+
  theme_bw()+
  theme(
        #panel.border = element_blank(),
        panel.grid.minor = element_blank(),
        # axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        # axis.text.y = element_blank(),
        # axis.ticks = element_blank(),
        plot.title = element_text(size=8, hjust=0.5),
        legend.position = c(0.8,0.25),
        # legend.title = element_text(size=7, colour = "grey20"),
        # legend.text = element_text(size=7, colour = "grey20"),
        legend.background = element_blank(),
        # legend.key.size = unit(0.3, "cm"),
        #legend.box.spacing = unit(-0.2, "cm"),
        text = element_text(size=8))+
  labs(y = "Wgtd. contribution to Cumulative Pressure", fill = "Pressure", title = "Fish-dominant feed")+
  guides(fill = "none"))

ggsave(here("figures/marine_disaggreation_by_ingredient_pressure_ge_allocation.jpg"), dpi = 600, height=12, width=8.9, units="cm")



(plant_diet_ingredients_p <- 
  ggplot(data = summarised_ingredient_pressure_df |> 
           filter(diet == "Plant-dominant feed"),
         aes(x = reorder(ingredient, mean), y = mean, fill=pressure))+
  coord_flip(clip="off")+
  geom_col(aes(fill=pressure), position = "dodge", width = 0.85, alpha = 0.8)+
  geom_errorbar(aes(x = ingredient, ymin = if_else(mean-sd<0, true = 0, false = mean-sd), ymax = mean+sd), position = position_dodge(width=0.85), width = 0, linewidth = 0.2)+
  scale_fill_manual(values = rev(rcartocolor::carto_pal(n=5, name = "Safe")))+
  #scale_y_continuous(expand = c(0,0), limits = c(0,2))+
  theme_bw()+
  theme(
        #panel.border = element_blank(),
        panel.grid.minor = element_blank(),
        # axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        # axis.text.y = element_blank(),
        # axis.ticks = element_blank(),
        plot.title = element_text(size=8, hjust=0.5),
        legend.position = c(0.8,0.13),
        legend.title = element_text(size=7, colour = "grey20"),
        legend.text = element_text(size=7, colour = "grey20"),
        legend.background = element_rect(fill="transparent"),
        legend.key.size = unit(0.3, "cm"),
        #legend.box.spacing = unit(-0.2, "cm"),
        text = element_text(size=8))+
  labs(y = "Wgtd. contribution to Cumulative Pressure", fill = "Pressure", title = "Plant-dominant feed"))+
  guides(fill = guide_legend(reverse = TRUE))

ggsave(here("figures/plant_disaggreation_by_ingredient_pressure_ge_allocation.jpg"), dpi = 600, height=12, width=8.9, units="cm")
 
(marine_diet_ingredients_p|plant_diet_ingredients_p) +
  plot_layout(guides = "keep")+
  guides(fill = guide_legend(reverse = TRUE))+
  plot_annotation(tag_levels = "a")&
  theme(plot.tag = element_text(size=8))
ggsave(here("figures/disaggregation_of_pressure_by_ingredient_ge_allocation.jpg"), dpi = 600, width = 18, height =13, units="cm")



#Proportional contributions

summarised_ingredient_pressure_df2 <- 
  marine_plant_pressure_ingredient_disagg_combined_df |> 
  group_by(diet, pressure, ingredient) |> 
  summarise(mean = mean(prop, na.rm=TRUE),
            sd = sd(prop, na.rm=TRUE),
            se = sd/sqrt(500),
            median = median(prop, na.rm = TRUE),
            lq = quantile(prop, 0.25),
            uq = quantile(prop, 0.75), .groups = "drop") |> 
  mutate(diet = if_else(diet == "marine diet", true = "Fish-dominant feed", false = "Plant-dominant feed")) |> 
  mutate(ingredient = if_else(ingredient == "foragefish", true = "Forage fish", false = ingredient)) |> 
   mutate(pressure = factor(pressure, levels = c("Water", "Phosphorus", "Nitrogen", "GHG", "Disturbance")))


(marine_diet_ingredients_p2 <- 
  ggplot(data = summarised_ingredient_pressure_df2 |> 
           filter(diet == "Fish-dominant feed"),
                aes(x = reorder(ingredient, mean), y = mean, fill=pressure))+
  coord_flip(clip="off")+
  geom_col(aes(fill=pressure), position = "dodge", width = 0.75, alpha = 0.8)+
  geom_errorbar(aes(x = ingredient, ymin = if_else(mean-sd<0, true = 0, false = mean-sd), ymax = mean+sd), position = position_dodge(width=0.75), width = 0, size = 0.2)+
  scale_fill_manual(values = rev(rcartocolor::carto_pal(n=5, name = "Safe")))+
    scale_y_continuous(expand = c(0,0), limits = c(0,0.302))+
  theme_bw()+
  theme(
        #panel.border = element_blank(),
        panel.grid.minor = element_blank(),
        # axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        # axis.text.y = element_blank(),
        # axis.ticks = element_blank(),
        plot.title = element_text(size=8, hjust=0.5),
        legend.position = c(0.8,0.25),
        legend.title = element_text(size=7, colour = "grey20"),
        legend.text = element_text(size=7, colour = "grey20"),
        legend.background = element_rect(fill="transparent"),
        legend.key.size = unit(0.3, "cm"),
        #legend.box.spacing = unit(-0.2, "cm"),
        text = element_text(size=8))+
  labs(y = "Prop. of Cumulative Pressure Index", fill = "Pressure", title = "Fish-dominant feed")+
  guides(fill = "none"))
ggsave(here("figures/supplementary/marine_disaggreation_by_ingredient_pressure_prop_ge_allocation.jpg"), dpi = 600, height=12, width=8.9, units="cm")



(plant_diet_ingredients_p2 <- 
  ggplot(data = summarised_ingredient_pressure_df2 |> 
           filter(diet == "Plant-dominant feed"),
         aes(x = reorder(ingredient, mean), y = mean, fill=pressure))+
  coord_flip(clip="off")+
  geom_col(aes(fill=pressure), position = "dodge", width = 0.85, alpha = 0.8)+
  geom_errorbar(aes(x = ingredient, ymin = if_else(mean-sd<0, true = 0, false = mean-sd), ymax = mean+sd), position = position_dodge(width=0.85), width = 0, size = 0.2)+
  scale_fill_manual(values = rev(rcartocolor::carto_pal(n=5, name = "Safe")))+
  scale_y_continuous(expand = c(0,0), limits = c(0,0.302))+
  theme_bw()+
  theme(
        #panel.border = element_blank(),
        panel.grid.minor = element_blank(),
        # axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        # axis.text.y = element_blank(),
        # axis.ticks = element_blank(),
        plot.title = element_text(size=8, hjust=0.5),
        legend.position = c(0.8,0.13),
        legend.title = element_text(size=7, colour = "grey20"),
        legend.text = element_text(size=7, colour = "grey20"),
        legend.background = element_rect(fill="transparent"),
        legend.key.size = unit(0.3, "cm"),
        #legend.box.spacing = unit(-0.2, "cm"),
        text = element_text(size=8))+
  labs(y = "Prop. of Cumulative Pressure Index", fill = "Pressure", title = "Plant-dominant feed"))+
  guides(fill = guide_legend(reverse = TRUE))
ggsave(here("figures/supplementary/plant_disaggreation_by_ingredient_pressure_prop_ge_allocation.jpg"), dpi = 600, height=12, width=8.9, units="cm")
 
(marine_diet_ingredients_p2|plant_diet_ingredients_p2) +
 plot_layout(guides = "keep")+
  guides(fill = guide_legend(reverse = TRUE))+
  plot_annotation(tag_levels = "a")&
  theme(plot.tag = element_text(size=8))

ggsave(here("figures/supplementary/disaggregation_of_pressure_by_ingredient_prop_ge_allocation.jpg"), dpi = 600, width = 18, height =13, units="cm")
```


#clear the environment
```{r}
rm(list = ls(all.names = TRUE))
```

