---
title: "CPI permutation"
author: "Rich Cottrell"
date: "14/02/2023"
output: html_document
---

Set up
```{r}
library(tidyverse)
library(ggpubr)
library(patchwork)
library(here)
library(sf)
library(terra)
library(rnaturalearth)
#library(rnaturalearthdata)
library(RColorBrewer)
#library(gameofthrones)
library(parallel)
library(ggdist)
library(data.table)
library(dtplyr)



select <- dplyr::select
values <- terra::values


#ALLOCATION METHOD (un-comment preference)
this_allocation_method <- "econ_allocation"
# this_allocation_method <- "ge_allocation"
# this_allocation_method <- "mass_allocation"
```


Bring in the origin-ingredient combinations within each diet.  

```{r}
marine_diets <- readRDS(here("data/tidy_data/diets/ingredient_source_combinations_marine_diet.rds")) 
plant_diets <- readRDS(here("data/tidy_data/diets/ingredient_source_combinations_plant_diet.rds")) 
```


#Permutation analysis - Energetic allocation

For each of the 500 combinations for marine-dominant and plant-dominant diets, identify the source of each ingredient and for each pressure. So perform a similar function for each above but summarise the rasters. We also need to repeat this for each allocation method.

Marine diets - Energetic Allocation

```{r}
marine_source_ingredient <- 
  marine_diets |>
  select(1:501) |> 
  rename_at(vars(-ingredient), ~ paste("source", .x)) |>
  mutate(row = row_number()) |> 
  mutate_at(vars(-ingredient, -row), ~ case_when(ingredient == "guar meal"~ as.integer(1),
                                           TRUE ~ .)) |> 
  group_split(row)

sources <- bind_rows(marine_source_ingredient) |> select(-c("ingredient", "row")) |>  names()   
marine_source_file_list <- list()
this_df <- marine_source_ingredient[[1]]
source <- sources[[1]]

#Now list the files for all ingredients and sources
for(source in sources){
    
    marine_diet_files <- 
      map(.x = marine_source_ingredient, .f = \(this_df){
        
        this_allocation_method <- "ge_allocation"
        
        this_ingredient <- this_df$ingredient
        
        message("processing...", this_ingredient, "...", source)
        
        these_ingredient_files <- list.files(here("data/spatial/marine_diet/pressures"), pattern = paste0(this_ingredient, "_"), full.names = TRUE)
        these_ingredient_files <- these_ingredient_files[grepl("ge_allocation", these_ingredient_files)]
        
        
        #sources
        this_source_ingredient <- this_df |> pull(source)
        
        #disturbance
        this_disturbance_file <- these_ingredient_files[grepl("disturbance", these_ingredient_files)][[this_source_ingredient]]
        
        #ghg
        this_ghg_file <- these_ingredient_files[grepl("ghg", these_ingredient_files)][[this_source_ingredient]]
        
        #nutrients
        this_nutrient_file <- these_ingredient_files[grepl("_nutrient_", these_ingredient_files)][[this_source_ingredient]]
        
        # 
        # #nitrogen
        # this_N_file <- these_ingredient_files[grepl("_N_", these_ingredient_files)][[this_source_ingredient]]
        # 
        # #phosphorus
        # this_P_file <- these_ingredient_files[grepl("_P_", these_ingredient_files)][[this_source_ingredient]]
        
        #water
        this_h2o_file <- these_ingredient_files[grepl("water", these_ingredient_files)][[this_source_ingredient]]
        
        
        these_diet_pressures <- c(this_disturbance_file, this_ghg_file, this_nutrient_file, this_h2o_file)
        return(these_diet_pressures)
        
      }) |> 
      unlist()
    
    marine_source_file_list[[source]] <- marine_diet_files
    
  }
  
#bring in rasters to list
marine_source_rast_list <- list()
for(source in sources){
  
  message("processing....", source)
  
  marine_rast_list <- 
  map(.x = marine_source_file_list[[source]], .f = \(this_file){
  
  
  this_rast <- rast(this_file)
  this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
  
  # this_rast[is.na(this_rast)] <- 0
  # names(this_rast) <- this_source
  return(this_rast)
})
  
  marine_source_rast_list[[source]] <- marine_rast_list
  
}
#stack and sum the pressures for all ingredients and source combinations
marine_diet_raster <- list()
for(source in sources){
  
  message("processing stack for combination....", source)
  
  #stack rasters for each sourcing combination
    marine_diet_raster[[source]] <- global(app(rast(marine_source_rast_list[[source]]), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(source = source, .before = sum)
  
  
}
marine_diet_raster_df <- bind_rows(marine_diet_raster) |> mutate(diet = "marine diet")
saveRDS(marine_diet_raster_df, file = here("data/tidy_data/pressures/marine_cumulative_pressure_df_ge_allocation.rds"))
```

Plant diets - Energetic allocation

```{r}
plant_source_ingredient <- 
  plant_diets |>
  select(1:501) |> 
  rename_at(vars(-ingredient), ~ paste("source", .x)) |>
  mutate(row = row_number()) |> 
  mutate_at(vars(-ingredient, -row), ~ case_when(ingredient == "linseed oil"~ as.integer(1),
                                           TRUE ~ .)) |> 
  group_split(row)
sources <- bind_rows(plant_source_ingredient) |> select(-c("ingredient", "row")) |>  names()   
plant_source_file_list <- list()
#test data for loop
# this_df <- plant_source_ingredient[[1]]
# source <- sources[[1]]
#Now list the files for all ingredients and sources
for(source in sources){
    
    plant_diet_files <- 
      map(.x = plant_source_ingredient, .f = \(this_df){
        
        this_allocation_method <- "ge_allocation"
        
        this_ingredient <- this_df$ingredient
        
        message("processing...", this_ingredient, "...", source)
        
        these_ingredient_files <- list.files(here("data/spatial/plant_diet/pressures"), pattern = paste0(this_ingredient, "_"), full.names = TRUE)
        these_ingredient_files <- these_ingredient_files[grepl("ge_allocation", these_ingredient_files)]
        
        #sources
        this_source_ingredient <- this_df |> pull(source)
        
        #disturbance
        this_disturbance_file <- these_ingredient_files[grepl("disturbance", these_ingredient_files)][[this_source_ingredient]]
        
        #ghg
        this_ghg_file <- these_ingredient_files[grepl("ghg", these_ingredient_files)][[this_source_ingredient]]
        
         #nutrients
        this_nutrient_file <- these_ingredient_files[grepl("_nutrient_", these_ingredient_files)][[this_source_ingredient]]
        
        # 
        # #nitrogen
        # this_N_file <- these_ingredient_files[grepl("_N_", these_ingredient_files)][[this_source_ingredient]]
        # 
        # #phosphorus
        # this_P_file <- these_ingredient_files[grepl("_P_", these_ingredient_files)][[this_source_ingredient]]
        
        #water
        this_h2o_file <- these_ingredient_files[grepl("water", these_ingredient_files)][[this_source_ingredient]]
        
        
        these_diet_pressures <- c(this_disturbance_file, this_ghg_file, this_nutrient_file, this_h2o_file)
        return(these_diet_pressures)
        
      }) |> 
      unlist()
    
    plant_source_file_list[[source]] <- plant_diet_files
    
  }
  
#bring in rasters to list
plant_source_rast_list <- list()
  
for(source in sources){
  
  message("processing....", source)
  
  plant_rast_list <- 
  map(.x = plant_source_file_list[[source]], .f = \(this_file){
  
  
  this_rast <- rast(this_file)
  this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
  # 
  # this_rast[is.na(this_rast)] <- 0
  # names(this_rast) <- this_source
  return(this_rast)
})
  
  plant_source_rast_list[[source]] <- plant_rast_list
}
#stack and sum the pressures for all ingredients and source combinations
plant_diet_raster <- list()
for(source in sources){
  
  message("processing stack for combination....", source)
  
  #stack rasters for each sourcing combination
  plant_diet_raster[[source]] <- global(app(rast(plant_source_rast_list[[source]]), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(source = source, .before = sum)
  
}
plant_diet_raster_df <- bind_rows(plant_diet_raster) |> mutate(diet = "plant diet")
saveRDS(plant_diet_raster_df, file = here("data/tidy_data/pressures/plant_cumulative_pressure_df_ge_allocation.rds"))
```

Energetic allocation - add marine and plant diet data together and save 

```{r}
marine_diet_raster_df <- readRDS(file = here("data/tidy_data/pressures/marine_cumulative_pressure_df_ge_allocation.rds"))
plant_diet_raster_df <- readRDS(file = here("data/tidy_data/pressures/plant_cumulative_pressure_df_ge_allocation.rds"))
marine_plant_df <- bind_rows(marine_diet_raster_df, plant_diet_raster_df)
saveRDS(marine_plant_df, file = here("data/tidy_data/pressures/marine_plant_pressures_combined_ge_allocation.rds"))
```


Energetic allocation - Visualise the differences between the two diets for cumulative pressures
```{r}
marine_plant_df <- readRDS(here("data/tidy_data/pressures/marine_plant_pressures_combined_ge_allocation.rds"))

my_pal <- rcartocolor::carto_pal(n = 8, name = "Bold")[c(1, 3, 7, 2)]

both_diet_means <- tibble(diet = c("Fish-\ndominant\nfeed", "Plant-\ndominant\nfeed"), 
                          mean_CPI = c(marine_plant_df |> filter(diet == "marine diet") |> pull(sum) |> mean(), 
                                       marine_plant_df |> filter(diet == "plant diet") |> pull(sum) |> mean()))


(both_diets_agg_p <- ggplot(data = marine_plant_df |> mutate(diet = case_when(diet == "plant diet" ~ "Plant-\ndominant\nfeed",
                                                         diet == "marine diet" ~ "Fish-\ndominant\nfeed"),
                                        diet = factor(diet, levels = c("Plant-\ndominant\nfeed", "Fish-\ndominant\nfeed"))))+
  aes(x = sum, y = diet, colour = diet)+
    geom_jitter(shape =20, size = 1.8, alpha = 0.1, height = 0.4
  )+
  # ggdist::stat_gradientinterval(
  #   width = .3, color = "black", fill_type = "gradient"
  # )+
  scale_colour_manual(values = my_pal[c(4,2)], guide = "none")+
  scale_x_continuous(limits = c(0,3))+
  theme_pubr()+
  geom_boxplot(notch = TRUE, colour = "grey40", width = 0.2, alpha = 0, linewidth=0.5)+
  geom_point(data = both_diet_means, aes(y = diet, x=mean_CPI), colour = "red", size = 1, alpha = 0.7, shape = 0)+
  labs(y = "Diet", x = "Cumulative pressure index (CPI)")+
  theme(text = element_text(size=8),
        axis.text = element_text(size=8),
        axis.title = element_text(size=8),
        axis.title.y = element_blank()))
ggsave(filename = here("figures/diet_permutation_comparisons_ge_allocation.jpg"), device = "jpg", dpi = 600, width = 8.9, height = 5, units="cm")
  
  
```

Energetic allocation - Some statistics for marine-plant comparisons
```{r}
marine_vals <-  marine_plant_df |> filter(diet == "marine diet") |> pull(sum)
plant_vals <- marine_plant_df |> filter(diet == "plant diet") |> pull(sum)
max_marine <- max(marine_vals)
min_marine <- min(marine_vals)
max_plant <- max(plant_vals)
min_plant <- min(plant_vals)
#How much variation in scores within a feed
max_marine/min_marine
max_plant/min_plant
#which marine vals are higher than the plant minimum?
marine_vals[marine_vals>min_plant] |> length()/ length(marine_vals) #93.2% of marine values
#which plant vals are lower than the marine maximum?
plant_vals[plant_vals<max_marine] |> length()/ length(plant_vals) #98.8%
#median comparisons
median(marine_vals)
median(plant_vals)
((median(plant_vals)/median(marine_vals))-1)*100 #10% higher median CPI for plants
#maximum comparisons
max(marine_vals)
max(plant_vals)
max(plant_vals)/max(marine_vals) #2.7% higher maximum CPI for plants
min(plant_vals)/min(marine_vals) #min plant diets have CPI 32.4% > marine diets
#minimum of one, maximum of anotehr
min(plant_vals)/max(marine_vals) #plant diets can have CPI 69% lower than marine diets
min(marine_vals)/max(plant_vals) #marine diets can have CPI 77% lower than plant diets
summary(marine_vals)
summary(plant_vals)
#manual calculation of CI
```



# Disaggregation of permutation by pressure - Energetic allocation

Once we have compared the two diets with different sourcing combinations, it's important to know how they disaggregate into the pressures that are the major drivers.
Pressure Disaggregation - Marine diets

```{r}
marine_source_ingredient <- 
  marine_diets |>
  select(1:501) |> 
  rename_at(vars(-ingredient), ~ paste("source", .x)) |>
  mutate(row = row_number()) |> 
  mutate_at(vars(-ingredient, -row), ~ case_when(ingredient == "guar meal"~ as.integer(1),
                                           TRUE ~ .)) |> 
  group_split(row)

sources <- bind_rows(marine_source_ingredient) |> select(-c("ingredient", "row")) |>  names()   

marine_source_file_list <- list()
this_df <- marine_source_ingredient[[1]]
source <- sources[[1]]
#Now list the files for all ingredients and sources

for(source in sources){
    
    marine_diet_files <- 
      map(.x = marine_source_ingredient, .f = \(this_df){
        
        this_ingredient <- this_df$ingredient
        
        message("processing...", this_ingredient, "...", source)
        
        these_ingredient_files <- list.files(here("data/spatial/marine_diet/pressures"), pattern = paste0(this_ingredient, "_"), full.names = TRUE)
        
        these_ingredient_files <- these_ingredient_files[grepl("ge_allocation", these_ingredient_files)]
        
        #sources
        this_source_ingredient <- this_df |> pull(source)
        
        #disturbance
        this_disturbance_file <- these_ingredient_files[grepl("disturbance", these_ingredient_files)][[this_source_ingredient]]
        
        #ghg
        this_ghg_file <- these_ingredient_files[grepl("ghg", these_ingredient_files)][[this_source_ingredient]]
        
         #nutrients
        this_nutrient_file <- these_ingredient_files[grepl("_nutrient_", these_ingredient_files)][[this_source_ingredient]]
        
        # 
        # #nitrogen
        # this_N_file <- these_ingredient_files[grepl("_N_", these_ingredient_files)][[this_source_ingredient]]
        # 
        # #phosphorus
        # this_P_file <- these_ingredient_files[grepl("_P_", these_ingredient_files)][[this_source_ingredient]]
        
        #water
        this_h2o_file <- these_ingredient_files[grepl("water", these_ingredient_files)][[this_source_ingredient]]
        
        
        these_diet_pressures <- c(this_disturbance_file, this_ghg_file, this_nutrient_file, this_h2o_file)
        return(these_diet_pressures)
        
      }) |> 
      unlist()
    
    marine_source_file_list[[source]] <- marine_diet_files
    
  }
  
#stack and sum by individual pressures for each ingredient source combination
md_disaggregated_pressures_df_list <- list()
for(source in sources){
  
  message("processing.....", source)
  
  this_combination_all_pressures_list <- marine_source_file_list[[source]]
  
  
  
  #########Separate GHG files, read rasters into a list, stack all GHG layers
  
  message("stacking ghg layers for...", source)
  
  this_combination_ghg_list <- this_combination_all_pressures_list[grep("ghg", this_combination_all_pressures_list)]
  
  these_ghg_rasters <- map(.x = this_combination_ghg_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_ghg_df <- global(app(rast(these_ghg_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "GHG", source = source, .before = sum)
  
  
  
  ##########Separate water files, read rasters into a list, stack all water layers
  
  message("stacking water layers for...", source)
  
  this_combination_water_list <- this_combination_all_pressures_list[grep("water", this_combination_all_pressures_list)]
  
  these_water_rasters <- map(.x = this_combination_water_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_water_df <- global(app(rast(these_water_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Water", source = source, .before = sum)
  
  
   #########Separate nutrient files, read rasters into a list, stack all nutrient layers
  
  message("stacking nutrient layers for...", source)
  
  this_combination_nutrient_list <-  this_combination_all_pressures_list[grep("_nutrient_", this_combination_all_pressures_list)]
  
  these_nutrient_rasters <- map(.x = this_combination_nutrient_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_nutrient_df <- global(app(rast(these_nutrient_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Nutrients", source = source, .before = sum)
  
# 
#   
#   #########Separate N files, read rasters into a list, stack all N layers
#   
#   message("stacking N layers for...", source)
#   
#   this_combination_N_list <-  this_combination_all_pressures_list[grep("_N_", this_combination_all_pressures_list)]
#   
#   these_N_rasters <- map(.x = this_combination_N_list, .f = \(this_file){
#     
#     this_rast <- rast(this_file)
#     this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
#     
#     return(this_rast)
#   })
#   
#   this_N_df <- global(app(rast(these_N_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Nitrogen", source = source, .before = sum)
#   
#   
#   #########Separate P files, read rasters into a list, stack all P layers
#   
#   message("stacking P layers for...", source)
#   
#   this_combination_P_list <-  this_combination_all_pressures_list[grep("_P_", this_combination_all_pressures_list)]
#   
#   these_P_rasters <- map(.x = this_combination_P_list, .f = \(this_file){
#     
#     this_rast <- rast(this_file)
#     this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
#     
#     return(this_rast)
#   })
#   
#   this_P_df <- global(app(rast(these_P_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Phosphorus", source = source, .before = sum)
#   
  
  
  #########Separate disturbance files, read rasters into a list, stack all disturbance layers
  
  message("stacking disturbance layers for...", source)
  
  this_combination_disturbance_list <-  this_combination_all_pressures_list[grep("disturbance", this_combination_all_pressures_list)]
  
  these_disturbance_rasters <- map(.x = this_combination_disturbance_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_disturbance_df <- global(app(rast(these_disturbance_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Disturbance", source = source, .before = sum)
  
  
  md_disaggregated_pressures_df_list[[source]] <- bind_rows(this_ghg_df, this_water_df, this_nutrient_df, this_disturbance_df)
  
}
marine_diet_pressures_df <- bind_rows(md_disaggregated_pressures_df_list) |> mutate(diet = "marine diet")
saveRDS(marine_diet_pressures_df, file = here("data/tidy_data/pressures/marine_disaggregated_pressures_df_ge_allocation.rds"))
```

Pressure disaggregation - Plant diets 

```{r}
plant_source_ingredient <- 
  plant_diets |>
  select(1:501) |> 
  rename_at(vars(-ingredient), ~ paste("source", .x)) |>
  mutate(row = row_number()) |> 
  mutate_at(vars(-ingredient, -row), ~ case_when(ingredient == "linseed oil"~ as.integer(1),
                                           TRUE ~ .)) |> 
  group_split(row)
sources <- bind_rows(plant_source_ingredient) |> select(-c("ingredient", "row")) |>  names()   
plant_source_file_list <- list()
#test data for loop
# this_df <- plant_source_ingredient[[1]]
# source <- sources[[1]]
#Now list the files for all ingredients and sources
for(source in sources){
    
    plant_diet_files <- 
      map(.x = plant_source_ingredient, .f = \(this_df){
        
        this_ingredient <- this_df$ingredient
        
        message("processing...", this_ingredient, "...", source)
        
        these_ingredient_files <- list.files(here("data/spatial/plant_diet/pressures"), pattern = paste0(this_ingredient, "_"), full.names = TRUE)
        
        these_ingredient_files <- these_ingredient_files[grepl("ge_allocation", these_ingredient_files)]
        
        #sources
        this_source_ingredient <- this_df |> pull(source)
        
        #disturbance
        this_disturbance_file <- these_ingredient_files[grepl("disturbance", these_ingredient_files)][[this_source_ingredient]]
        
        #ghg
        this_ghg_file <- these_ingredient_files[grepl("ghg", these_ingredient_files)][[this_source_ingredient]]
        
        #nutrients
        this_nutrient_file <- these_ingredient_files[grepl("_nutrient_", these_ingredient_files)][[this_source_ingredient]]
        
        # 
        # #nitrogen
        # this_N_file <- these_ingredient_files[grepl("_N_", these_ingredient_files)][[this_source_ingredient]]
        # 
        # #phosphorus
        # this_P_file <- these_ingredient_files[grepl("_P_", these_ingredient_files)][[this_source_ingredient]]
        
        #water
        this_h2o_file <- these_ingredient_files[grepl("water", these_ingredient_files)][[this_source_ingredient]]
        
        
        these_diet_pressures <- c(this_disturbance_file, this_ghg_file, this_nutrient_file, this_h2o_file)
        return(these_diet_pressures)
        
        #water
        this_h2o_file <- these_ingredient_files[grepl("water", these_ingredient_files)][[this_source_ingredient]]
        
        
        these_diet_pressures <- c(this_disturbance_file, this_ghg_file, this_nutrient_file, this_h2o_file)
        return(these_diet_pressures)
        
      }) |> 
      unlist()
    
    plant_source_file_list[[source]] <- plant_diet_files
    
  }
  
  
#stack and sum by individual pressures for each ingredient source combination
pd_disaggregated_pressures_df_list <- list()
for(source in sources){
  
  message("processing.....", source)
  
  this_combination_all_pressures_list <- plant_source_file_list[[source]]
  
  
  
  #########Separate GHG files, read rasters into a list, stack all GHG layers
  
  message("stacking ghg layers for...", source)
  
  this_combination_ghg_list <- this_combination_all_pressures_list[grep("ghg", this_combination_all_pressures_list)]
  
  these_ghg_rasters <- map(.x = this_combination_ghg_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_ghg_df <- global(app(rast(these_ghg_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "GHG", source = source, .before = sum)
  
  
  
  ##########Separate water files, read rasters into a list, stack all water layers
  
  message("stacking water layers for...", source)
  
  this_combination_water_list <- this_combination_all_pressures_list[grep("water", this_combination_all_pressures_list)]
  
  these_water_rasters <- map(.x = this_combination_water_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_water_df <- global(app(rast(these_water_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Water", source = source, .before = sum)
  
  
  
  #########Separate nutrient files, read rasters into a list, stack all nutrient layers
  
  message("stacking nutrient layers for...", source)
  
  this_combination_nutrient_list <-  this_combination_all_pressures_list[grep("_nutrient_", this_combination_all_pressures_list)]
  
  these_nutrient_rasters <- map(.x = this_combination_nutrient_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_nutrient_df <- global(app(rast(these_nutrient_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Nutrients", source = source, .before = sum)
  
  # 
  #   
  #   #########Separate N files, read rasters into a list, stack all N layers
  #   
  #   message("stacking N layers for...", source)
  #   
  #   this_combination_N_list <-  this_combination_all_pressures_list[grep("_N_", this_combination_all_pressures_list)]
  #   
  #   these_N_rasters <- map(.x = this_combination_N_list, .f = \(this_file){
  #     
  #     this_rast <- rast(this_file)
  #     this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
  #     
  #     return(this_rast)
  #   })
  #   
  #   this_N_df <- global(app(rast(these_N_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Nitrogen", source = source, .before = sum)
  #   
  #   
  #   #########Separate P files, read rasters into a list, stack all P layers
  #   
  #   message("stacking P layers for...", source)
  #   
  #   this_combination_P_list <-  this_combination_all_pressures_list[grep("_P_", this_combination_all_pressures_list)]
  #   
  #   these_P_rasters <- map(.x = this_combination_P_list, .f = \(this_file){
  #     
  #     this_rast <- rast(this_file)
  #     this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
  #     
  #     return(this_rast)
  #   })
  #   
  #   this_P_df <- global(app(rast(these_P_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Phosphorus", source = source, .before = sum)
  #   
  
  
  
  #########Separate disturbance files, read rasters into a list, stack all disturbance layers
  
  message("stacking disturbance layers for...", source)
  
  this_combination_disturbance_list <-  this_combination_all_pressures_list[grep("disturbance", this_combination_all_pressures_list)]
  
  these_disturbance_rasters <- map(.x = this_combination_disturbance_list, .f = \(this_file){
    
    this_rast <- rast(this_file)
    this_source <- tools::file_path_sans_ext(basename(sources(this_rast)))
    
    return(this_rast)
  })
  
  this_disturbance_df <- global(app(rast(these_disturbance_rasters), sum, na.rm=TRUE), sum, na.rm=TRUE) |> mutate(pressure = "Disturbance", source = source, .before = sum)
  
  
  pd_disaggregated_pressures_df_list[[source]] <- bind_rows(this_ghg_df, this_water_df, this_nutrient_df, this_disturbance_df)
  
}
plant_diet_pressures_df <- bind_rows(pd_disaggregated_pressures_df_list) |> mutate(diet = "plant diet")
saveRDS(plant_diet_pressures_df, file = here("data/tidy_data/pressures/plant_disaggregated_pressures_df_ge_allocation.rds"))
```

Add the marine and plant pressure disaggregation together and save

```{r}
marine_pressure_disagg <- readRDS(here("data/tidy_data/pressures/marine_disaggregated_pressures_df_ge_allocation.rds"))
plant_pressure_disagg <- readRDS(here("data/tidy_data/pressures/plant_disaggregated_pressures_df_ge_allocation.rds"))
marine_plant_pressure_disagg_df <- bind_rows(marine_pressure_disagg, plant_pressure_disagg)
saveRDS(object = marine_plant_pressure_disagg_df, file = here("data/tidy_data/pressures/marine_plant_disaggregated_pressures_combined_df_ge_allocation.rds"))
marine_pressure_disagg
plant_pressure_disagg
```


Visualise the pressure disaggregation

```{r}
marine_plant_pressure_disagg_df <- 
  readRDS(here("data/tidy_data/pressures/marine_plant_disaggregated_pressures_combined_df_ge_allocation.rds")) |> 
  group_by(diet, source) |> 
  nest() |> 
  mutate(total = map(data, ~(sum(.$sum)))) |> 
  unnest(c(data, total)) |> 
  mutate(prop = sum/total)

my_pal <- rcartocolor::carto_pal(n = 8, name = "Bold")[c(1, 3, 7, 2)]


ggplot(data = marine_plant_pressure_disagg_df |> 
                              mutate(diet = case_when(diet == "plant diet" ~ "Plant-dominant\nfeed",
                                                      diet == "marine diet" ~ "Fish-dominant\nfeed"),
                                                      diet = factor(diet, levels = c("Plant-dominant\nfeed", "Fish-dominant\nfeed"))))+
  aes(x = prop, y = reorder(pressure, prop), fill = diet)+
  geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.5), alpha = 0.1)+
  geom_boxplot(notch = TRUE, colour = "grey35", alpha = 0.4, linewidth=0.5)+
  scale_fill_manual(values = my_pal[c(4,2)])+
  scale_x_continuous(limits = c(0,1))+
  theme_pubr()+
  theme(legend.position = c(0.85, 0.15),
        legend.background = element_rect(fill="transparent"),
        legend.title = element_blank(),
        text = element_text(size=8),
        axis.text = element_text(size=8),
        axis.title = element_text(size=8),
        axis.title.y = element_blank())+
  labs(x = "Proportion of CPI")



ggsave(filename = here("figures/disaggregation_of_pressures_ge_allocation.jpg"), width = 12, height=9, units = "cm", device = "jpg", dpi = 600)
```

Combine aggregated and disaggregated figures
```{r}
both_diets_agg_p+both_diets_disagg_p+
  plot_annotation(tag_levels = "a")+
  theme(plot.tag = element_text(size=8))
ggsave(here("figures/both_diet_agg_disagg_combine_ge_allocation.jpg"), dpi = 600, width = 18.9, height = 8, units="cm")
```

Some calculation regarding the decomposition of CPI by individual pressures
```{r}
#Disturbance (marine)
marine_plant_pressure_disagg_df |> filter(pressure == "Disturbance" & diet == "marine diet") |> pull(prop) |> max()
marine_plant_pressure_disagg_df |> filter(pressure == "Disturbance" & diet == "marine diet") |> pull(prop) |> min()
marine_plant_pressure_disagg_df |> filter(pressure == "Disturbance" & diet == "marine diet") |> pull(prop) |> mean()
marine_plant_pressure_disagg_df |> filter(pressure == "Disturbance" & diet == "marine diet") |> pull(prop) |> sd()
#Disturbance (plant)
marine_plant_pressure_disagg_df |> filter(pressure == "Disturbance" & diet == "plant diet") |> pull(prop) |> max()
marine_plant_pressure_disagg_df |> filter(pressure == "Disturbance" & diet == "plant diet") |> pull(prop) |> min()
marine_plant_pressure_disagg_df |> filter(pressure == "Disturbance" & diet == "plant diet") |> pull(prop) |> mean()
marine_plant_pressure_disagg_df |> filter(pressure == "Disturbance" & diet == "plant diet") |> pull(prop) |> sd()
```


#clear the environment
```{r}
rm(list = ls(all.names = TRUE))
```

