---
title: "20_ingredient_case_studies"
author: "Rich Cottrell"
date: "14/02/2023"
output: html_document
---


Set up
```{r}
library(tidyverse)
library(ggpubr)
library(patchwork)
library(here)
library(sf)
library(terra)
library(rnaturalearth)
#library(rnaturalearthdata)
library(RColorBrewer)
library(gameofthrones)
library(parallel)
library(ggdist)
library(data.table)
library(dtplyr)


# taking cumulative impoact palette from Halpern et al 
red <- "#B90000"
d_brown <- "#515256"
m_brown <- "#B27B54"
l_brown <- "#BC995F"
green <- "#A8B072"
yellow<- "#EFCE71"
l_yellow <- "#F7F6C1"
light_gray <- "#F8F9FA"
discrete_pal <- c(l_yellow, yellow, green, l_brown, m_brown, d_brown, "firebrick")
continuous_pal <-  colorRampPalette(discrete_pal, space="Lab", bias = 3.5)(10000)
final_palette <- c(light_gray, continuous_pal, red)

#source files
source(here("src/fxns.R"))

#spatial products
moll_crs <- "ESRI:53009"
gp_crs <- "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
gp_rast <- rast() |> project(y = gp_crs, res  = 10000)

gp_crs_pacific <- "+proj=cea +lon_0=160 +x_0=0 +y_0=0 +lat_ts=45 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

countries <-   ne_countries(scale = "large", returnclass = "sf") |> st_transform(crs = gp_crs)

countries_pac <- ne_countries(scale = "large", returnclass = "sf") 

maptools::nowrapRecenter(countries, offset = 180-150, avoidGEOS=TRUE)
countries_pacific <- st_transform(st_segmentize(countries_pac$geometry,5000), gp_crs_pacific)


bbox <- ne_download(scale = 50, type = "wgs84_bounding_box", category = "physical", returnclass = "sf") |> st_transform(crs = moll_crs)
coastline <- ne_coastline(scale = 50, returnclass = "sf") |> st_transform(crs = moll_crs)
select <- dplyr::select
values <- terra::values

ggplot()+geom_sf(data = countries_pacific)
```

Pull in the dietary data to compre percentage composition of different ingredients in each
```{r}

total_demand <- readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds"))
marine_composition <- readRDS(here("data/tidy_data/diet-scenarios/marine_diet.rds"))
plant_composition <- readRDS(here("data/tidy_data/diet-scenarios/plant_diet.rds"))
marine_composition
plant_composition |> filter(diet == "plant_diet")

```


Now some case studies of ingredients 


Fish oil - disturbance

```{r}

fish_oil_files_md <- list.files(here("data/spatial/marine_diet/int"), pattern = "fish oil_production_disturbance_moll", full=TRUE)
fish_oil_files_pd <- list.files(here("data/spatial/plant_diet/int"), pattern = "fish oil_disturbance_gp_mass_allocation", full=TRUE) 

fish_oil_regions <- str_sub(tools::file_path_sans_ext(fish_oil_files_md), start = nchar(tools::file_path_sans_ext(fish_oil_files_md))-2, end = nchar(tools::file_path_sans_ext(fish_oil_files_md)))

fish_oil_marine <-  fish_oil_files_md |> 
  rast() |> 
  global(fun = sum, na.rm=TRUE) |> 
  mutate(origins = fish_oil_regions) |> 
  rename(total_disturbance = sum) |> 
  mutate(feed_scenario = "Fish-dominant")

fish_oil_plant <- fish_oil_files_pd |> 
  rast() |> global(fun = sum, na.rm=TRUE) |> 
   mutate(origins = fish_oil_regions) |> 
  rename(total_disturbance = sum)|> 
  mutate(feed_scenario = "Plant-dominant")

fish_oil_all <- bind_rows(fish_oil_marine, fish_oil_plant) |> 
  mutate(origins = case_when(origins == "EPC" ~ "East Pacific",
                             origins == "NAT" ~ "North Atlantic",
                             origins == "OTH" ~ "All other regions")) |> as_tibble() |> 
  mutate(origins = factor(origins, levels = c("")))

#pull in the forage fish png for the plot
ff_png <- png::readPNG(here("data/tidy_data/png/engraulis-australis-australian-anchovy.png"))
ff_png <- grid::rasterGrob(ff_png)

(fish_oil_distrubance_p <- 
  ggplot()+
  geom_col(data = fish_oil_all, 
           aes(x = feed_scenario, y = total_disturbance, fill = reorder(origins, -total_disturbance)), position = "dodge", width = 0.6)+
  geom_hline(yintercept = fish_oil_all |> filter(feed_scenario == "Fish-dominant") |> pull(total_disturbance) |> min(), linetype = 2, colour = "grey30", linewidth = 0.4)+
  labs(fill = "Origins", y = bquote(Total~habitat~disturbance~("1000"*s~km^2~eq.)), x= "Feed scenario")+
  ggpubr::theme_pubr()+
  scale_y_continuous(limits = c(0,200000), breaks = c(0,50000, 100000, 150000, 200000), labels = c(0, 50,100,150,200))+
  theme(axis.text = element_text(size=8),
        axis.title = element_text(size = 8),
        axis.title.x = element_blank(),
        legend.title = element_text(size=6),
        legend.text = element_text(size=6),
        legend.position = c(0.85,0.85),
        legend.key.size = unit(0.4, "cm"))+
  scale_fill_manual(values = rcartocolor::carto_pal(n=3, name = "TealGrn"))+
  annotate("text", x = 0.5, y = 200000, label =  "Fish oil", size=2.8, hjust=0, fontface = "bold")+
  annotation_custom(ff_png, xmin = 0.5, xmax = 0.8, ymin = 175000, ymax = 190000))
  
ggsave(here("explore/fish_oil_disturbance.jpg"), device = "jpeg", dpi=600, width = 9.8, height = 7, units = "cm")

```




Wheat - nitrogen pollution

```{r}

wheat_files_md <- list.files(here("data/spatial/marine_diet/int"), pattern = "wheat_N_gp_mass_allocation", full=TRUE)
wheat_files_pd <- list.files(here("data/spatial/plant_diet/int"), pattern = "wheat_N_gp_mass_allocation", full=TRUE) 

wheat_countries <- str_sub(tools::file_path_sans_ext(wheat_files_md), start = nchar(tools::file_path_sans_ext(wheat_files_md))-2, end = nchar(tools::file_path_sans_ext(wheat_files_md)))

wheat_marine <-  wheat_files_md |> 
  rast() |> 
  global(fun = sum, na.rm=TRUE) |> 
  mutate(origins = wheat_countries) |> 
  rename(total_N = sum) |> 
  mutate(feed_scenario = "Fish-dominant")

wheat_plant <- wheat_files_pd |> 
  rast() |> global(fun = sum, na.rm=TRUE) |> 
   mutate(origins = wheat_countries) |> 
  rename(total_N = sum)|> 
  mutate(feed_scenario = "Plant-dominant")

wheat_all <- bind_rows(wheat_marine, wheat_plant) |> 
  mutate(origins_new = countrycode::countrycode(sourcevar = origins, origin = "iso3c", destination = "country.name")) |> as_tibble()

#pull in wheat png for plot
wheat_png <- png::readPNG(here("data/tidy_data/png/wheat-sheath.png"))
wheat_png <- grid::rasterGrob(wheat_png)

(wheat_n_p <- ggplot()+
  geom_col(data = wheat_all, 
           aes(x = feed_scenario, y = total_N, fill = reorder(origins_new, -total_N)), position = "dodge")+
  geom_hline(yintercept = wheat_all |> filter(feed_scenario == "Plant-dominant") |> pull(total_N) |> min(), linetype = 2,colour = "grey30", linewidth = 0.4)+
  labs(fill = "Origins", y = bquote(Total~nitrogen~pollution~(tonnes~N~eq.)), x= "Feed scenario")+
  ggpubr::theme_pubr()+
  scale_y_continuous(limits = c(0,500000), breaks = c(0,100000,200000,300000,400000,500000), labels = c(0,100,200,300,400,500))+
  theme(axis.text = element_text(size=8),
        axis.title = element_text(size = 8),
        legend.title = element_text(size=6),
        legend.text = element_text(size=6),
        legend.position = c(0.9,0.8),
        legend.key.size = unit(0.4, "cm"))+
  scale_fill_manual(values = rcartocolor::carto_pal(n=5, name = "BrwnYl"))+
  annotate("text", x = 0.5, y = 500000, label = "Wheat", size=2.8, hjust=0, fontface = "bold")+
  annotation_custom(wheat_png, xmin = 0.4, xmax = 0.8, ymin = 420000, ymax = 480000))
  
ggsave(here("explore/wheat_nitrogen.jpg"), device = "jpeg", dpi=600, width = 9.8, height = 7, units = "cm")


```


Put the fish oil and wheat plots together
```{r}

fish_oil_distrubance_p/wheat_n_p+
  # plot_layout(guides = "collect")+
  plot_annotation(tag_levels = "a")&
  theme(plot.tag = element_text(size=9))



ggsave(here("figures/case_studies_comparison_column.jpg"), device = "jpeg", dpi = 600, width = 8.9, height =14, units="cm")

```



Some statistics about demand versus total pressure for both fish oil and wheat
```{r}

#Fish oil demand

#marine diet
(mar_fo_demand <- readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet == "marine_diet" & ingredients == "fish oil-wild") |> pull(total_ingredient_demand))

#plant diet
(plt_fo_demand <- readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet == "plant_diet" & ingredients == "fish oil-wild") |> pull(total_ingredient_demand))


#Embodied biomass and total disturbance pressures

#EPC - Marine diet
(mar_epc_fish_biomass <- readRDS(here("data/tidy_data/demand/embodied_fish_demand_by_source_diet.rds")) |> 
  filter(demand_from == "Fish oil" & allocation == "Mass allocation" & diet == "marine_diet", source_code == "EPC") |> pull(embodied_fish))

(mar_total_disturbance_epc <-list.files(here("data/spatial/marine_diet/int"), pattern = "fish oil_disturbance_gp_mass_allocation", full=TRUE) |> pluck(1) |> rast() |> global(fun = sum, na.rm=TRUE) |> pull(sum)
)

#NAT - Marine diet
(mar_nat_fish_biomass <- readRDS(here("data/tidy_data/demand/embodied_fish_demand_by_source_diet.rds")) |> 
  filter(demand_from == "Fish oil" & allocation == "Mass allocation" & diet == "marine_diet", source_code == "NAT") |> pull(embodied_fish))

(mar_total_disturbance_nat <- list.files(here("data/spatial/marine_diet/int"), pattern = "fish oil_disturbance_gp_mass_allocation", full=TRUE) |> pluck(2) |> rast() |> global(fun = sum, na.rm=TRUE) |> pull(sum))


#OTH - Marine diet
(mar_oth_fish_biomass <- readRDS(here("data/tidy_data/demand/embodied_fish_demand_by_source_diet.rds")) |> 
  filter(demand_from == "Fish oil" & allocation == "Mass allocation" & diet == "marine_diet", source_code == "NAT") |> pull(embodied_fish))

(mar_total_disturbance_oth <- list.files(here("data/spatial/marine_diet/int"), pattern = "fish oil_disturbance_gp_mass_allocation", full=TRUE) |> pluck(3) |> rast() |> global(fun = sum, na.rm=TRUE) |> pull(sum))



#EPC - plant diet
(plt_epc_fish_biomass <- readRDS(here("data/tidy_data/demand/embodied_fish_demand_by_source_diet.rds")) |> 
  filter(demand_from == "Fish oil" & allocation == "Mass allocation" & diet == "plant_diet", source_code == "EPC") |> pull(embodied_fish))

(plt_total_disturbance_epc <-list.files(here("data/spatial/plant_diet/int"), pattern = "fish oil_disturbance_gp_mass_allocation", full=TRUE) |> pluck(1) |> rast() |> global(fun = sum, na.rm=TRUE) |> pull(sum))


#NAT - plant diet
(plt_nat_fish_biomass <- readRDS(here("data/tidy_data/demand/embodied_fish_demand_by_source_diet.rds")) |> 
  filter(demand_from == "Fish oil" & allocation == "Mass allocation" & diet == "plant_diet", source_code == "NAT") |> pull(embodied_fish))

(plt_total_disturbance_nat <- list.files(here("data/spatial/plant_diet/int"), pattern = "fish oil_disturbance_gp_mass_allocation", full=TRUE) |> pluck(2) |> rast() |> global(fun = sum, na.rm=TRUE) |> pull(sum))


#OTH - Plant diet
(plt_oth_fish_biomass <- readRDS(here("data/tidy_data/demand/embodied_fish_demand_by_source_diet.rds")) |> 
  filter(demand_from == "Fish oil" & allocation == "Mass allocation" & diet == "plant_diet", source_code == "OTH") |> pull(embodied_fish))

(plt_total_disturbance_oth <- list.files(here("data/spatial/plant_diet/int"), pattern = "fish oil_disturbance_gp_mass_allocation", full=TRUE) |> pluck(3) |> rast() |> global(fun = sum, na.rm=TRUE) |> pull(sum))







#Wheat

#marine diet
(mar_wheat_demand <- readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet == "marine_diet" & ingredients == "wheat") |> pull(total_ingredient_demand))

#plant diet
(plt_wheat_demand <- readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet == "plant_diet" & ingredients == "wheat") |> pull(total_ingredient_demand))


#how much higher can marine-dominant feed nitrogen from wheat be despite less demand for wheat?
(mar_total_N_aus <- list.files(here("data/spatial/marine_diet/int"), pattern = "wheat_N_gp_mass_allocation", full=TRUE) |> pluck(3) |> rast() |> global(fun = sum, na.rm=TRUE) |> pull(sum))/ (plt_total_N_aus <- list.files(here("data/spatial/plant_diet/int"), pattern = "wheat_N_gp_mass_allocation", full=TRUE) |> pluck(1) |> rast() |> global(fun = sum, na.rm=TRUE) |> pull(sum))


 
total_crop_demand

```


```{r}

wheat_fra <- rast("/mnt/rdsi/github/feed_pressure_mapping/data/spatial/marine_diet/production/wheat_AllTech_mass_allocation_FRA.tif")
wheat_aus <- rast("/mnt/rdsi/github/feed_pressure_mapping/data/spatial/marine_diet/production/wheat_AllTech_mass_allocation_AUS.tif")

plot(wheat_fra)
wheat_fra |> global(sum, na.rm=TRUE)
wheat_aus |> global(sum, na.rm=TRUE)





```














Mapping approaches - couldn't get them to look good with unequal plotting regions in dimension and geographical coverage
```{r}
#Create dataframes for plotting for high and low pressure origins for fish oil 

#Example of disturbance associated from fish oil from area with low pressure for the marine diet (EPC)
mar_fo_low_epc <- list.files(here("data/spatial/marine_diet/int"), pattern = "fish oil_disturbance_gp_mass_allocation", full=TRUE) |> pluck(1) |> rast() |> terra::as.data.frame(xy = TRUE) |> mutate(source = "Eastern Pacific", diet = "Fish-dominant feed", ingredient_pressure = "Fish oil - low pressure origins")
 

#Example of disturbance associated from fish oil from area with higher pressure for the marine diet (NAT)
mar_fo_hi_nat <- list.files(here("data/spatial/marine_diet/int"), pattern = "fish oil_disturbance_gp_mass_allocation", full=TRUE) |> pluck(2) |> rast() |> terra::as.data.frame(xy = TRUE) |> mutate(source = "North Atlantic", diet = "Fish-dominant feed", ingredient_pressure = "Fish oil - high pressure origins")


#Example of disturbance associated from fish oil from area with low pressure for the plant diet (EPC)
plt_fo_low_epc <- list.files(here("data/spatial/plant_diet/int"), pattern = "fish oil_disturbance_gp_mass_allocation", full=TRUE) |> pluck(1) |> rast() |> terra::as.data.frame(xy = TRUE) |> mutate(source = "Eastern Pacific", diet = "Plant-dominant feed", ingredient_pressure = "Fish oil - low pressure origins")

#Example of disturbance associated from fish oil from area with higher pressure for the plant diet (EPC)
plt_fo_hi_nat <- list.files(here("data/spatial/plant_diet/int"), pattern = "fish oil_disturbance_gp_mass_allocation", full=TRUE) |> pluck(2) |> rast() |> terra::as.data.frame(xy = TRUE) |> mutate(source = "North Atlantic", diet = "Plant-dominant feed", ingredient_pressure = "Fish oil - high pressure origins")

all_df <- bind_rows(mar_fo_low_epc, mar_fo_hi_nat, plt_fo_low_epc, plt_fo_hi_nat)


#get disturbance values for each location relative to embodied biomass and fish oil demand

#EPC - Marine diet
mar_fo_demand <- readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet == "marine_diet" & ingredients == "fish oil-wild") |> pull(total_ingredient_demand)

mar_epc_fish_biomass <- readRDS(here("data/tidy_data/demand/embodied_fish_demand_by_source_diet.rds")) |> 
  filter(demand_from == "Fish oil" & allocation == "Mass allocation" & diet == "marine_diet", source_code == "EPC") |> pull(embodied_fish)
mar_total_disturbance_epc <-list.files(here("data/spatial/marine_diet/int"), pattern = "fish oil_disturbance_gp_mass_allocation", full=TRUE) |> pluck(1) |> rast() |> global(fun = sum, na.rm=TRUE) |> pull(sum)

#NAT - Marine diet
mar_fo_demand <- readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet == "marine_diet" & ingredients == "fish oil-wild") |> pull(total_ingredient_demand)
mar_nat_fish_biomass <- readRDS(here("data/tidy_data/demand/embodied_fish_demand_by_source_diet.rds")) |> 
  filter(demand_from == "Fish oil" & allocation == "Mass allocation" & diet == "marine_diet", source_code == "NAT") |> pull(embodied_fish)
mar_total_disturbance_nat <- list.files(here("data/spatial/marine_diet/int"), pattern = "fish oil_disturbance_gp_mass_allocation", full=TRUE) |> pluck(2) |> rast() |> global(fun = sum, na.rm=TRUE) |> pull(sum)


#EPC - plant diet
plt_fo_demand <- readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet == "plant_diet" & ingredients == "fish oil-wild") |> pull(total_ingredient_demand)
plt_epc_fish_biomass <- readRDS(here("data/tidy_data/demand/embodied_fish_demand_by_source_diet.rds")) |> 
  filter(demand_from == "Fish oil" & allocation == "Mass allocation" & diet == "plant_diet", source_code == "EPC") |> pull(embodied_fish)
plt_total_disturbance_epc <-list.files(here("data/spatial/plant_diet/int"), pattern = "fish oil_disturbance_gp_mass_allocation", full=TRUE) |> pluck(1) |> rast() |> global(fun = sum, na.rm=TRUE) |> pull(sum)


#NAT - Marine diet
mar_fo_demand <- readRDS(here("data/tidy_data/demand/total_ingredient_demand.rds")) |> filter(diet == "plant_diet" & ingredients == "fish oil-wild") |> pull(total_ingredient_demand)
plt_nat_fish_biomass <- readRDS(here("data/tidy_data/demand/embodied_fish_demand_by_source_diet.rds")) |> 
  filter(demand_from == "Fish oil" & allocation == "Mass allocation" & diet == "plant_diet", source_code == "NAT") |> pull(embodied_fish)
plt_total_disturbance_nat <- list.files(here("data/spatial/plant_diet/int"), pattern = "fish oil_disturbance_gp_mass_allocation", full=TRUE) |> pluck(2) |> rast() |> global(fun = sum, na.rm=TRUE) |> pull(sum)



#Fish dominant low pressure fish oil disturbance

fo_lo_crop_box <- st_bbox(mar_fo_low_epc |> select(x,y,lyr.1) |>  rast(type="xyz"))
fo_lo_crop_box[["xmax"]] <- 500000

(mar_lo_fo_p <-  ggplot()+
     geom_sf(data = countries |> 
            st_crop(y = fo_lo_crop_box) , 
           linewidth = 0.1, fill="grey70",colour="grey10")+
  geom_tile(data = mar_fo_low_epc, aes(x = x, y = y, fill = lyr.1))+
  scale_fill_gradientn(colours = rev(rcartocolor::carto_pal(n = 7, name = "ag_Sunset")), limits = c(0,450))+
  theme_bw()+
    labs(fill = bquote(atop(Disturbance, (D*","~km^2~eq.))), y = "Fish oil - low pressure origins")+
    theme(axis.title.x = element_blank(),
          axis.title.y = element_text(size = 10),
          legend.text = element_text(size = 10),
          legend.title = element_text(size=10),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.grid = element_blank())+
    annotate("text", hjust=0, x=fo_lo_crop_box[["xmin"]], y=fo_lo_crop_box[["ymin"]]*0.8, label = bquote(FO~demand~"="~.(mar_fo_demand)~MT),size=2.5)+
    annotate("text", hjust=0, x=fo_lo_crop_box[["xmin"]], y=fo_lo_crop_box[["ymin"]]*0.88, label = bquote(Embodied~fish~"="~.(mar_epc_fish_biomass)~MT),size=2.5)+
    annotate("text", hjust=0, x=fo_lo_crop_box[["xmin"]], y=fo_lo_crop_box[["ymin"]]*0.96, label = bquote(Sigma~D~"="~.(mar_total_disturbance_epc)~km^2~eq.),size=2.5)
  )

#Plant-dominant low pressure fish oil disturbance
(plt_lo_fo_p <- ggplot()+
     geom_sf(data = countries |> 
            st_crop(y = fo_lo_crop_box) , 
           linewidth = 0.1, fill="grey70",colour="grey10")+
  geom_tile(data = plt_fo_low_epc, aes(x = x, y = y, fill = lyr.1))+
 
  scale_fill_gradientn(colours = rev(rcartocolor::carto_pal(n = 7, name = "ag_Sunset")), limits = c(0,450))+
  theme_bw()+
    labs(fill = bquote(atop(Disturbance, (D*","~km^2~eq.))))+
    theme(axis.title.x = element_blank(),
          axis.title.y = element_text(size = 10),
          legend.text = element_text(size = 10),
          legend.title = element_text(size=10),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.grid = element_blank())+
    annotate("text", hjust=0, x=fo_lo_crop_box[["xmin"]], y=fo_lo_crop_box[["ymin"]]*0.8, label = bquote(FO~demand~"="~.(plt_fo_demand)~MT),size=2.5)+
    annotate("text", hjust=0, x=fo_lo_crop_box[["xmin"]], y=fo_lo_crop_box[["ymin"]]*0.88, label = bquote(Embodied~fish~"="~.(plt_epc_fish_biomass)~MT),size=2.5)+
    annotate("text", hjust=0, x=fo_lo_crop_box[["xmin"]], y=fo_lo_crop_box[["ymin"]]*0.96, label = bquote(Sigma~D~"="~.(plt_total_disturbance_epc)~km^2~eq.),size=2.5)
)    


#Fish dominant high pressure

fo_hi_crop_box <- st_bbox(mar_fo_hi_nat |> select(x,y,lyr.1) |>  rast(type="xyz"))
fo_hi_crop_box[["ymin"]] <- 500000

(mar_hi_fo_p <- ggplot()+
      geom_tile(data = mar_fo_hi_nat, aes(x = x, y = y, fill = lyr.1))+
     geom_sf(data = countries |> 
            st_crop(y = fo_hi_crop_box) , 
          linewidth = 0.1, fill="grey70",colour="grey10")+
  scale_fill_gradientn(colours = rev(rcartocolor::carto_pal(n = 7, name = "ag_Sunset")), limits = c(0,450))+
  theme_bw()+
  labs(fill = bquote(atop(Disturbance, (D*","~km^2~eq.))), y = "Fish oil - high pressure origins")+
    theme(axis.title.x = element_blank(),
          axis.title.y = element_text(size = 10),
          legend.text = element_text(size = 10),
          legend.title = element_text(size=10),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.grid = element_blank())+
    annotate("text", hjust=0, x=fo_hi_crop_box[["xmin"]], y=fo_hi_crop_box[["ymin"]]*2, label = bquote(FO~demand~"="~.(mar_fo_demand)~MT),size=2.5)+
    annotate("text", hjust=0, x=fo_hi_crop_box[["xmin"]], y=fo_hi_crop_box[["ymin"]]*1.5, label = bquote(Embodied~fish~"="~.(mar_nat_fish_biomass)~MT),size=2.5)+
    annotate("text", hjust=0, x=fo_hi_crop_box[["xmin"]], y=fo_hi_crop_box[["ymin"]]*1, label = bquote(Sigma~D~"="~.(mar_total_disturbance_nat)~km^2~eq.),size=2.5)+
    guides(fill = "none")
)    


#plant dominant high pressure

(plt_hi_fo_p <- ggplot()+
      geom_tile(data = plt_fo_hi_nat, aes(x = x, y = y, fill = lyr.1))+
     geom_sf(data = countries |> 
            st_crop(y = fo_hi_crop_box) , 
          linewidth = 0.1, fill="grey70",colour="grey10")+

  scale_fill_gradientn(colours = rev(rcartocolor::carto_pal(n = 7, name = "ag_Sunset")), limits = c(0,450))+
  theme_bw()+
  labs(fill = bquote(atop(Disturbance, (D*","~km^2~eq.))))+
   theme(axis.title.x = element_blank(),
          axis.title.y = element_text(size = 10),
          legend.text = element_text(size = 10),
          legend.title = element_text(size=10),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.grid = element_blank())+
    annotate("text", hjust=0, x=fo_hi_crop_box[["xmin"]], y=fo_hi_crop_box[["ymin"]]*2, label = bquote(FO~demand~"="~.(plt_fo_demand)~MT),size=2.5)+
    annotate("text", hjust=0, x=fo_hi_crop_box[["xmin"]], y=fo_hi_crop_box[["ymin"]]*1.5, label = bquote(Embodied~fish~"="~.(plt_nat_fish_biomass)~MT),size=2.5)+
    annotate("text", hjust=0, x=fo_hi_crop_box[["xmin"]], y=fo_hi_crop_box[["ymin"]]*1, label = bquote(Sigma~D~"="~.(plt_total_disturbance_nat)~km^2~eq.),size=2.5)
)    


#facet grid approach

#get the two bounding boxes for low (EPC) and higher (NAT) pressure origins
fo_lo_crop_box <- st_bbox(mar_fo_low_epc_r |> select(x,y,lyr.1) |>  rast(type="xyz"))
fo_lo_crop_box[["xmax"]] <- 500000

fo_hi_crop_box <- st_bbox(mar_fo_hi_nat |> select(x,y,lyr.1) |>  rast(type="xyz"))
fo_hi_crop_box[["ymin"]] <- 500000

#combine and create one bounding box based on minima and maxima of both regions
crop_box_all <- bind_rows(fo_lo_crop_box, fo_hi_crop_box)
fo_bbox <- c(apply(crop_box_all[,c("xmin", "ymin")], MARGIN = 2, FUN = min), apply(crop_box_all[,c("xmax", "ymax")], MARGIN = 2, FUN = max))


(all_fo_p <- ggplot()+
      geom_tile(data = all_df, aes(x = x, y = y, fill = lyr.1))+
     geom_sf(data = countries |>
            st_crop(y = fo_bbox) ,
          linewidth = 0.1, fill="grey50",colour="grey50")+
  scale_fill_gradientn(colours = rev(rcartocolor::carto_pal(n = 7, name = "ag_Sunset")), limits = c(0,450))+
  labs(fill = bquote(atop(Disturbance, (D*","~km^2~eq.))))+
  theme_bw()+
    theme(axis.title = element_blank(),
          legend.text = element_text(size = 10),
          legend.title = element_text(size=10),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.grid = element_blank())+
  facet_grid(rows = vars(ingredient_pressure), cols = vars(diet))+
  guides(fill = guide_colourbar(barwidth = 0.5, unit = "cm"))
)


# (mar_lo_fo_p | plt_lo_fo_p)/
# (mar_hi_fo_p | plt_hi_fo_p)+
#   plot_layout(guides = "collect")+
#   plot_annotation(tag_levels = "a",
#                   title = "                      Fish-dominant feed                                Plant-dominant feed") & 
#   theme(plot.tag = element_text(size=9), 
#         plot.tag.position = c(0.1, 1),
#         legend.position = "right")
#   


ggsave(here("explore/fo_test_plot.jpeg"), device = "jpg", dpi = 600, width = 18, height =12, units = "cm")



```


Wheat - GHG

```{r}
list.files(here("data/spatial/marine_diet/int"), pattern = "wheat_N_gp_mass_allocation", full=TRUE) |> rast() |> global(fun = sum, na.rm=TRUE)
list.files(here("data/spatial/plant_diet/int"), pattern = "wheat_N_gp_mass_allocation", full=TRUE) |> rast() |> global(fun = sum, na.rm=TRUE)


#Example of nitrogen pollution associated with wheat from area with low pressure for the marine diet (ukraine)
mar_wheat_low_ukr <- list.files(here("data/spatial/marine_diet/int"), pattern = "wheat_N_gp_mass_allocation", full=TRUE) |> pluck(4) |> rast()  |> terra::as.data.frame(xy = TRUE) |> mutate(source = "Russia", diet = "Fish-dominant feed", ingredient_pressure = "Wheat - high pressure origins")

#Example of nitrogen pollution associated with wheat from area with high pressure for the marine diet (France)
mar_wheat_hi_fra <- list.files(here("data/spatial/marine_diet/int"), pattern = "wheat_N_gp_mass_allocation", full=TRUE) |> pluck(1) |> rast() |> terra::as.data.frame(xy = TRUE) |> mutate(source = "Australia", diet = "Fish-dominant feed", ingredient_pressure = "Wheat - low pressure origins")

#Example of nitrogen pollution associated with wheat from area with low pressure for the plant diet (ukraine)
plt_wheat_low_ukr <- list.files(here("data/spatial/plant_diet/int"), pattern = "wheat_N_gp_mass_allocation", full=TRUE) |> pluck(4) |> rast() |> terra::as.data.frame(xy = TRUE) |> mutate(source = "Russia", diet = "Plant-dominant feed", ingredient_pressure = "Wheat - high pressure origins")

#Example of nitrogen pollution associated with wheat from area with high pressure for the plant diet (france)
plt_wheat_hi_fra <- list.files(here("data/spatial/plant_diet/int"), pattern = "wheat_N_gp_mass_allocation", full=TRUE) |> pluck(1) |> rast() |> terra::as.data.frame(xy = TRUE) |> mutate(source = "Australia", diet = "Plant-dominant feed", ingredient_pressure = "Wheat - low pressure origins")



wheat_all_df <- bind_rows(mar_wheat_low_ukr, mar_wheat_hi_fra, plt_wheat_low_ukr, plt_wheat_hi_fra)

#get bounding box of both low and high regions combined
wheat_lo_bbox <- st_bbox(mar_wheat_low_ukr |> select(x,y, N_eq) |> rast())
wheat_hi_bbox <- st_bbox(mar_wheat_hi_fra |> select(x,y,N_eq) |> rast())

#get the bouding box to crop the country shapefiles to
crop_box_wheat_all <- bind_rows(wheat_lo_bbox, wheat_hi_bbox) 
wheat_bbox <- c(apply(crop_box_wheat_all[,c("xmin", "ymin")], MARGIN = 2, FUN = min), apply(crop_box_wheat_all[,c("xmax", "ymax")], MARGIN = 2, FUN = max))

#now customise bounding box to not look so rubbish
wheat_bbox[["ymin"]] <- -6100000
wheat_bbox[["ymax"]]
wheat_bbox[["xmin"]] <- 90000
wheat_bbox[["xmax"]] <- 15000000

(all_wheat_p <- ggplot()+
    geom_sf(data = countries |>
            st_crop(y = wheat_bbox) ,
          linewidth = 0.1, fill="grey50",colour="grey50")+
      geom_tile(data = wheat_all_df, aes(x = x, y = y, fill = N_eq))+
     
  scale_fill_gradientn(colours = (rcartocolor::carto_pal(n = 7, name = "ag_GrnYl")), limits = c(0,450))+
  labs(fill = "Nitrogen\npollution\n(kg N eq.)" )+
  theme_bw()+
    theme(axis.title = element_blank(),
          legend.text = element_text(size = 10),
          legend.title = element_text(size=10),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.grid = element_blank())+
  facet_grid(rows = vars(ingredient_pressure), cols = vars(diet))+ 
    guides(fill = guide_colourbar(barwidth = 0.5, unit = "cm")))


ggsave(here("explore/wheat_test_plot.jpeg"), device = "jpg", dpi = 600, width = 18, height =12, units = "cm")


```

Combine plots

```{r}
(fo_dist_p/wheat_N_p)+
  plot_annotation(tag_levels = "a")+
  plot_layout(heights = c(1.2,1))
ggsave(filename = here("figures/ingredient_casestudies.jpg"), device = "jpg", dpi = 600, width = 18, height = 26, units = "cm")
```